---
title: "Environmental taxes and the path to net-zero in the European Union"
author: "Daphné Perry and Justine Catta"
format:
  html: 
    toc: true
  wordcount-html:
    toc: true
  pdf:
    toc: true
    sansfont: "Times New Roman"
    mainfont: "Times New Roman"
    geometry:
      - bottom=2.2cm
      - top=2.5cm
      - left=2.5cm
      - right=2.5cm
    include-in-header:
      text: |
        \usepackage{caption}
        \captionsetup[table]{labelformat=empty}
execute:
  echo: false
  warning: false
  message: false
---

{{< pagebreak >}}

# Introduction

## 1. Description of our research question

Our research question is the following: “To what extent do environmental taxes reduce Greenhouse Gas Emissions (GHG) in EU countries, and what are their long-term socioeconomic impacts across economy?”. As the European Union commits to achieving net-zero emissions by 2050 under the European Green Deal, we aim to better understand the effectiveness and consequences of environmental taxation, both on emissions and socially.

To answer this question, we will use panel data for all EU member states. Our data set includes GHG emissions, GDP, environmental tax revenues, and several indicators of social well-being. These include the Gender Inequality Index (GII), a coefficient of human inequality, and a measure of energy poverty that is the inability to adequately heat one’s home. Together, these variables allow us to evaluate the distributive consequences of taxes.

After cleaning, harmonizing, and describing the data sets, we aim to estimate several relationships. First, we will measure the link between environmental taxation and GHG emissions. Do higher taxes lead to statistically significant reductions in emissions at EU level? 

We will then investigate the association between environmental taxes and social inequality indicators. Our core question here can be summarized as: “Is there a social cost associated with environmental taxation, and if so, how large is it?” Specifically, we will analyse whether environmental taxes correlate with increases or decreases in gender inequality, income inequality, and energy poverty. This step is essential because policies that are environmentally effective but socially regressive may have a reduced long-term viability.

Building on these empirical relationships, we will construct forecasts for the period 2026–2050 to estimate the level of taxation required to meet the European Green Deal’s emissions targets. We consider a policy scenario relying solely on environmental taxation applied to relevant sectors. Our analysis will assess whether such an approach is socially and economically sustainable: How would different tax levels affect inequalities and energy poverty? How much revenue would such taxes generate, and would the required tax burden be compatible with each country’s GDP levels?

By combining environmental effectiveness, economic feasibility, and social impact, we can then address our overarching question: Is it socially and economically optimal for the EU to rely exclusively on environmental taxation to achieve net-zero emissions by 2050? If taxation alone proves insufficient or too socially costly, our results may highlight the need for complementary policies such as subsidies for green technologies or targeted social transfers to ensure a fair and effective ecological transition.

## 2. Link to Github

Our project is available on [Github](https://github.com/cattajustine-spec/project).

```{r echo=FALSE}
#| message: false
#| warning: false
# Setup
library(vroom)
library(ggplot2)
library(knitr)
library(dplyr)
library(here)
library(countrycode)
library(plotly)
library(tidyr)
library(tidyverse)
library(forecast)
library(stargazer)
```

```{r, echo=FALSE}
#| message: false
#| warning: false
here::i_am("project.Rproj")
```

## 3. Our data:
### a. Environmental tax revenues

This data set reports annual environmentally related tax revenues for EU countries from the ([OECD](https://data-explorer.oecd.org/vis?lc=en&tm=DF_ERTR&pg=0&snb=1&vw=tb&df%5Bds%5D=dsDisseminateFinalDMZ&df%5Bid%5D=DSD_ERTR%40DF_ERTR&df%5Bag%5D=OECD.ENV.EPI&df%5Bvs%5D=&pd=1994%2C2023&dq=A.HRV%2BLVA%2BLTU%2BLUX%2BMLT%2BNLD%2BCYP%2BAUT%2BBEL%2BCZE%2BDNK%2BEST%2BFIN%2BFRA%2BDEU%2BGRC%2BHUN%2BIRL%2BITA%2BPOL%2BPRT%2BSVK%2BSVN%2BESP%2BSWE%2BBGR%2BROU.TAXREV.CAT_ENE%2BCAT_TRA%2BCAT_POL%2BCAT_RES%2B_T._T.USD.&ly%5Brw%5D=REF_AREA&ly%5Bcl%5D=TIME_PERIOD&to%5BTIME_PERIOD%5D=false)), with values in millions of USD covering from 1994 to 2023.

```{r, echo=FALSE}
#| message: false
#| warning: false
Envt_Tax_Revenues <- vroom(here("data", "OECD_envt_tax.csv"))
```

::: {columns}
|                   |                             |
|-------------------|-----------------------------|
| Name              | `r "Envt_Tax_Revenues"`     |
| Number of rows    | `r nrow(Envt_Tax_Revenues)` |
| Number of columns | `r ncol(Envt_Tax_Revenues)` |
:::

### b. GDP

This data set provides annual GDP figures for EU countries from the ([OECD](https://data-explorer.oecd.org/vis?df%5Bds%5D=dsDisseminateFinalDMZ&df%5Bid%5D=DSD_NAMAIN10%40DF_TABLE1_OUTPUT&df%5Bag%5D=OECD.SDD.NAD&df%5Bvs%5D=&dq=A.BEL%2BAUT%2BCZE%2BDNK%2BEST%2BFIN%2BFRA%2BDEU%2BGRC%2BHUN%2BIRL%2BITA%2BLVA%2BLTU%2BLUX%2BNLD%2BPOL%2BPRT%2BSVK%2BSVN%2BESP%2BSWE%2BBGR%2BHRV%2BCYP%2BMLT%2BROU...B1G%2BB1GQ.._Z%2B_T%2BC%2BA%2BBTE%2BF%2BGTI%2BJ%2BK%2BL%2BM_N%2BOTQ%2BRTU..USD_EXC.V..&to%5BTIME_PERIOD%5D=false&pd=1990%2C2024&vw=tb)), measured in millions of USD, using exchange rate and covering from 1990 to 2024.

```{r, echo=FALSE}
#| message: false
#| warning: false
GDP_and_AV <- vroom(here("data", "OECD_GDP_and_AV.csv"))
```

::: {columns}
|                   |                      |
|-------------------|----------------------|
| Name              | `r "GDP_and_AV"`     |
| Number of rows    | `r nrow(GDP_and_AV)` |
| Number of columns | `r ncol(GDP_and_AV)` |
:::

### c. Greenhouse gas emissions

This data set reports total greenhouse gas emissions for the 27 EU countries, in thousand tonnes of CO$₂$-equivalent. The data set is from ([Eurostat](https://ec.europa.eu/eurostat/databrowser/view/env_air_gge__custom_18942108/default/table)) and covers from 1990 to 2023.

```{r, echo=FALSE}
#| message: false
#| warning: false
GHG_Emissions <- vroom(here("data", "Eurostat_ghg.csv"))
```

::: {columns}
|                   |                         |
|-------------------|-------------------------|
| Name              | `r "GHG_Emissions"`     |
| Number of rows    | `r nrow(GHG_Emissions)` |
| Number of columns | `r ncol(GHG_Emissions)` |
:::

### d. Inability to keep home adequately warm

This data set collects the percentage of households reporting they are unable to keep their home adequately warm. We chose 3 households type: total, single female and single male; and 2 income situations: total and below 60% of the median (in situation of poverty). The data set is from  ([Eurostat](https://ec.europa.eu/eurostat/databrowser/view/ILC_MDES01__custom_18945942/default/table)) and covers 2003-2024.

```{r, echo=FALSE}
#| message: false
#| warning: false
Warm_Households <- vroom(here("data","Eurostat_warm.csv"))
```

::: {columns}
|                   |                           |
|-------------------|---------------------------|
| Name              | `r "Warm_Households"`     |
| Number of rows    | `r nrow(Warm_Households)` |
| Number of columns | `r ncol(Warm_Households)` |
:::

### e. Human indexes

We focus on the Coefficient of Human Inequality and Gender Inequality Index (GII), from this ([UNDP HDR](https://hdr.undp.org/sites/default/files/2025_HDR/HDR25_Composite_indices_complete_time_series.csv)) data set providing annual values for lots of Indexes, from 1990 to 2023.

```{r, echo=FALSE}
#| message: false
#| warning: false
Soc_Indexes <- vroom(here("data", "HDR_indexes.csv"))
```

::: {columns}
|                   |                       |
|-------------------|-----------------------|
| Name              | `r "Soc_Indexes"`     |
| Number of rows    | `r nrow(Soc_Indexes)` |
| Number of columns | `r ncol(Soc_Indexes)` |
:::

{{< pagebreak >}}

# I. Modifying the data sets ^[For extensive description, please refer to the *"dm-project phase 2.qmd"* document.]

The data sets being quite large, we first remove variables we will not use and add aggregates that we think are important, before characterizing the variables we kept. Finally, to simplify our analysis, we merge all our sets in one data base. 

## 1. Country keys {#countrykeys}

We construct a key that filtering the EU-27 countries and harmonizing the different identifiers, to ensure consistency.

```{r, echo=FALSE}
#| message: false
#| warning: false
eu_countries <- unique(GHG_Emissions$geo)

eu_country_key <- data.frame(name_en = eu_countries) |>
  mutate(
    iso3 = countrycode(name_en, origin = "country.name", destination = "iso3c"),
    iso2 = countrycode(name_en, origin = "country.name", destination = "iso2c"),
    name_fr = countrycode(name_en, origin = "country.name", destination = "country.name.fr"))  |>
  bind_rows(tibble(name_en = "EU27", name_fr = "EU27", iso2 = "EU27", iso3 = "EU27"))
```

## 2. Removing unwanted variables

For each dataset, we remove units, metadata, etc and systematically retain four essential components:

-   Geographical identifier (ISO2, ISO3, or name).
-   Reference year.
-   Observed value (e.g., revenues, emissions, ...).

For the HDR index, we  create two independent data sets, one for the Coefficient of Human inequality and another for the GII.

```{r, echo=FALSE}
#| message: false
#| warning: false

### 1. Environmental tax revenues
Envt_Tax_Revenues <- Envt_Tax_Revenues |>
  filter(REF_AREA %in% eu_country_key$iso3,
         CATEGORY == "_T") |>
  select(REF_AREA, TIME_PERIOD, OBS_VALUE)

### 2. GDP (in million of USD)
GDP <- GDP_and_AV |>
  filter(Transaction == "Produit intérieur brut", 
         REF_AREA %in% eu_country_key$iso3) |>
  select(REF_AREA, TIME_PERIOD, OBS_VALUE)

### 3. Greenhouse gas emissions (in million of ton)
GHG_Emissions <- GHG_Emissions |>
  filter(geo %in% eu_country_key$name_en, 
         src_crf == "Total (excluding memo items)")|>
    select(geo, TIME_PERIOD, OBS_VALUE)
    
### 4. Inability to keep home adequately warm
Warm_Households <- Warm_Households |>
  select(geo, TIME_PERIOD, hhtyp, incgrp, OBS_VALUE) |>
  filter(geo %in% eu_country_key$iso2, 
         hhtyp %in% c("A1F", "A1M", "TOTAL"))
         
### 5. HDR index
# described per year as different variables (meaning we have a gii_2020, gii_2021, ..., coef_ineq_2003, coef_ineq_2004, etc.). Therefore, we need to pivot our database, and then create two independent data sets, one for each variable of interest.
Soc_Indexes <- Soc_Indexes |>
  pivot_longer(                           
    cols = matches("_[0-9]{4}$"),   
    names_to = c("Indicator", "TIME_PERIOD"),
    names_pattern = "^(.*)_(\\d{4})$",
    values_to = "OBS_VALUE"
  ) |>
  mutate(
    OBS_VALUE = as.numeric(OBS_VALUE),
    TIME_PERIOD = as.integer(TIME_PERIOD)
  )

#### Gender inequality Index
GII <- Soc_Indexes|>
  filter(iso3 %in% eu_country_key$iso3,
         Indicator == "gii") |>
  select(iso3, TIME_PERIOD, OBS_VALUE)

#### Coefficient of Human inequality 
coef_ineq <- Soc_Indexes|>
  filter(iso3 %in% eu_country_key$iso3,
         Indicator == "coef_ineq")|>
  select(iso3, TIME_PERIOD, OBS_VALUE)
```

## 3. Adding aggregates

First, for the data sets that describe inequality indexes, we compute the European average across countries.

```{r, echo=FALSE}
#| message: false
#| warning: false
EU27_GII <- GII |>  #computing the average European gender inequality index. 
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(iso3 = "EU27") |>
  relocate(iso3)
GII <- bind_rows(GII, EU27_GII)

EU27_Coefineq <- coef_ineq |> #computing the average European coefficient of human inequality. 
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(iso3 = "EU27") |>
  relocate(iso3)
coef_ineq <- bind_rows(coef_ineq, EU27_Coefineq)

EU27_Warming <- Warm_Households |>  #computing the average European inability to keep home adequately warm, depending on the household composition and income.  
  group_by(TIME_PERIOD, hhtyp, incgrp) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(geo = "EU27") |>
  relocate(geo)
#adding in Warm households, using country keys to have by iso3
Warm_Households <- Warm_Households |> 
  rename(iso2 = geo) |>
  left_join(eu_country_key |> select(iso2, iso3), by = "iso2")
EU27_Warming <- EU27_Warming |> 
  rename(iso2 = geo) |> 
  left_join(eu_country_key |> select(iso2, iso3), by = "iso2")
Warm_Households <- bind_rows(Warm_Households, EU27_Warming) 
Warm_Households <- Warm_Households |> 
  select (iso3, TIME_PERIOD, hhtyp, incgrp, OBS_VALUE)
```

Secondly, for other economic data sets (GDP, Taxes, GHG emissions), we compute the European total across countries.

```{r, echo=FALSE}
#| message: false
#| warning: false
EU27_GDP <- GDP |>                #computing the total European GDP 
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) |>
  mutate(REF_AREA = "EU27") |>
  relocate(REF_AREA)
GDP <- bind_rows(GDP, EU27_GDP)

EU27_Tax <- Envt_Tax_Revenues |>   #computing the total European environmental tax revenue
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) |>
  mutate(REF_AREA = "EU27") |>
  relocate(REF_AREA)
Envt_Tax_Revenues <- bind_rows(Envt_Tax_Revenues, EU27_Tax)

EU27_GHG <- GHG_Emissions |> #computing the total European greenhouse gas emissions
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) |>
  mutate(geo = "EU27") |>
  relocate(geo)
#adding in GHG_Emissions, using country keys to have by iso3
GHG_Emissions <- GHG_Emissions |> 
  rename(name_en = geo) |>
  left_join(eu_country_key |> select(name_en, iso3), by = "name_en")
EU27_GHG <- EU27_GHG  |> 
  rename(name_en = geo) |> 
  left_join(eu_country_key |> select(name_en, iso3), by = "name_en")
GHG_Emissions <- bind_rows(GHG_Emissions, EU27_GHG) 
GHG_Emissions <- GHG_Emissions  |> select (iso3, TIME_PERIOD, OBS_VALUE)
```

## 4. Modifiying the characteristics of the variables

For the inability to keep housing adequately warm variable, we consider the household type variable and the income group variable as *factor*.

```{r, echo=FALSE}
#| message: false
#| warning: false
Warm_Households$hhtyp <- factor(Warm_Households$hhtyp)
EU27_Warming$hhtyp <- factor(EU27_Warming$hhtyp)
Warm_Households$incgrp <- factor(Warm_Households$incgrp)
EU27_Warming$incgrp <- factor(EU27_Warming$incgrp)
```

We rename variables to be able to merge our data sets; from "**OBS_VALUE**" appearing in several data sets to "**name-of-dataset_value**".

```{r, echo=FALSE}
#| message: false
#| warning: false
names(Warm_Households)[names(Warm_Households) == "OBS_VALUE"] <- "Warm_Households_value"
names(EU27_Warming)[names(EU27_Warming) == "OBS_VALUE"] <- "Warm_Households_value"
names(Envt_Tax_Revenues)[names(Envt_Tax_Revenues) == "OBS_VALUE"] <- "Envt_Tax_Revenues_value"
names(EU27_Tax)[names(EU27_Tax) == "OBS_VALUE"] <- "Envt_Tax_Revenues_value"
names(GDP)[names(GDP) == "OBS_VALUE"] <- "GDP_value"
names(EU27_GDP)[names(EU27_GDP) == "OBS_VALUE"] <- "GDP_value"
names(GHG_Emissions)[names(GHG_Emissions) == "OBS_VALUE"] <- "GHG_Emissions_value"
names(EU27_GHG)[names(EU27_GHG) == "OBS_VALUE"] <- "GHG_Emissions_value"
names(GII)[names(GII) == "OBS_VALUE"] <- "GII_value"
names(EU27_GII)[names(EU27_GII) == "OBS_VALUE"] <- "GII_value"
names(coef_ineq)[names(coef_ineq) == "OBS_VALUE"] <- "coef_ineq_value"
names(EU27_Coefineq)[names(EU27_Coefineq) == "OBS_VALUE"] <- "coef_ineq_value"
```

We also rename the country identifiers as the ones in the country key we created.

```{r, echo=FALSE}
#| message: false
#| warning: false
Envt_Tax_Revenues <- Envt_Tax_Revenues |>
  rename(iso3 = REF_AREA)
EU27_Tax <- EU27_Tax |>
  rename(iso3 = REF_AREA)

GDP <- GDP |> 
  rename(iso3 = REF_AREA)
EU27_GDP <- EU27_GDP |> 
  rename(iso3 = REF_AREA)
```

## 5. Merging

We can now merge all datasets into a single analytical framework on a country-year pair basis : we match on iso3 code, reference year, with *left_join()*. 

```{r, echo=FALSE}
#| message: false
#| warning: false
all_merged <- Envt_Tax_Revenues |>
  left_join(GDP, by = c("iso3", "TIME_PERIOD")) |>
  left_join(GHG_Emissions, by = c("iso3", "TIME_PERIOD")) |>
  left_join(Warm_Households, by = c("iso3", "TIME_PERIOD")) |>
  left_join(GII, by = c("iso3", "TIME_PERIOD")) |>
  left_join(coef_ineq, by = c("iso3", "TIME_PERIOD")) |>
  select(iso3, TIME_PERIOD, hhtyp, incgrp, Warm_Households_value, GII_value, coef_ineq_value, Envt_Tax_Revenues_value, GDP_value, GHG_Emissions_value) |>
  filter(iso3 != "EU27")
```

{{< pagebreak >}}

# II. Descriptive analyses ^[For extensive analyses, please refer to the *"dm-project phase 2.qmd"* document.]

## 1. European environmental tax revenues

```{r, echo=FALSE}
#| message: false
#| warning: false
ggplot(EU27_Tax, aes(x = TIME_PERIOD, y = Envt_Tax_Revenues_value)) +
  geom_line(linewidth = 1, color = "darkgreen") +
  labs(
    title = "Total European environmental tax revenues",
    x = "Year",
    y = "Environmental tax revenues (millions of USD)"
  ) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 410000))
```

```{r, echo=FALSE}
#| message: false
#| warning: false
summary_stats_taxrevenues <- EU27_Tax |>
  summarise(
    Min = min(Envt_Tax_Revenues_value, na.rm = TRUE),
    Max = max(Envt_Tax_Revenues_value, na.rm = TRUE),
    Mean = mean(Envt_Tax_Revenues_value, na.rm = TRUE),
    Median = median(Envt_Tax_Revenues_value, na.rm = TRUE),
    SD = sd(Envt_Tax_Revenues_value, na.rm = TRUE)
    )

kable(summary_stats_taxrevenues, 
      caption = "Descriptive statistics of the total European environmental tax revenues (all years combined)")
```

The total tax revenues increased from 190 billion USD in the late 1990s to 350 billion USD by 2010. Since then, the trend seems broadly stable, with noticeable fluctuations from year to year.

## 2. European GDP

```{r, echo=FALSE}
#| message: false
#| warning: false
ggplot(EU27_GDP, aes(x = TIME_PERIOD, y = GDP_value)) +
  geom_line(linewidth = 1, color = "darkgreen") +
  labs(
    title = "European GDP (in millions of USD)",
    x = "Year",
    y = "GDP (million USD)"
  ) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 2e07))
```

```{r, echo=FALSE}
#| message: false
#| warning: false
summary_stats_GDP <- EU27_GDP |>
  summarise(
    Min = min(GDP_value, na.rm = TRUE),
    Max = max(GDP_value, na.rm = TRUE),
    Mean = mean(GDP_value, na.rm = TRUE),
    Median = median(GDP_value, na.rm = TRUE),
    SD = sd(GDP_value, na.rm = TRUE)
    )

kable(summary_stats_GDP, 
      caption = "Descriptive statistics of the European GDP (all years combined)")
```

European GDP has increased significantly from 6.3 trillion USD in 1990 to 19.4 trillion USD in 2024.

## 3. Greenhouse gas emissions 

```{r, echo=FALSE}
#| message: false
#| warning: false
ggplot(EU27_GHG, aes(x = TIME_PERIOD, y = GHG_Emissions_value)) +
  geom_line(linewidth = 1, color = "darkgreen") +
  labs(
    title = "Total European greenhouse gas emissions",
    subtitle = "EU-27, in thousand tonnes of CO2",
    y = "Emissions"
    ) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 4600000))
```


```{r, echo=FALSE}
#| message: false
#| warning: false
summary_stats_GHG <- EU27_GHG |>
  summarise(
    Min = min(GHG_Emissions_value, na.rm = TRUE),
    Max = max(GHG_Emissions_value, na.rm = TRUE),
    Mean = mean(GHG_Emissions_value, na.rm = TRUE),
    Median = median(GHG_Emissions_value, na.rm = TRUE),
    SD = sd(GHG_Emissions_value, na.rm = TRUE)
    )

kable(summary_stats_GHG, caption = "Descriptive statistics of EU total greenhouse gases emissions (all years combined)")
```

The data shows a clear a downward trend for the GHG: their maximum in 1990 is around 4.6 billion tonnes of CO$₂$-equivalent, and their minimum one in 2023 is around 2.9 billion tonnes of CO$₂$-equivalent.

## 4. Inability to keep home adequately warm per household type

```{r, echo=FALSE}
#| message: false
#| warning: false
ggplot(EU27_Warming, 
       aes(x = TIME_PERIOD, y = Warm_Households_value,
           color = interaction(hhtyp, incgrp),
           group = interaction(hhtyp, incgrp))) +
  geom_line(linewidth = 1) +
  labs(
    title = "EU27 – Inability to Keep the House Warm",
    subtitle = "By sex and income group",
    x = "Year",
    y = "Percentage of population",
    color = "Sex × Income"
  ) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 30))
```

```{r, echo=FALSE}
#| message: false
#| warning: false
summary_stats_Warm <- EU27_Warming |>
  group_by(hhtyp, incgrp) |>                    
  summarise(
    Min = min(Warm_Households_value, na.rm = TRUE),
    Max = max(Warm_Households_value, na.rm = TRUE),
    Mean = mean(Warm_Households_value, na.rm = TRUE),
    Median = median(Warm_Households_value, na.rm = TRUE),
    SD = sd(Warm_Households_value, na.rm = TRUE), 
    .groups = "drop"
  )

summary_stats_Warm |>
  mutate(across(where(is.numeric), ~ formatC(.x, format = "f", digits = 2))) |>
  kable(
    caption = "Descriptive statistics of the inability to keep house adequately warm, by household type (all years combined)"
  )
```

The six curves, defined by sex and income, follow a similar pattern: a sharp rise around 2003, a gradual decline, and notable increases around 2009 and 2021.

Overall, the total population shows the lowest percentages, while households below 60% of the national median income show the highest. Among low-income single households, women report slightly higher rates than men (about 23.1% versus 21.1%).

## 5. GII per country ^[For a more interactive and readable view of the graph in this section, please refer to the *"dm-project phase 3.html"* document.]

```{r, echo=FALSE}
#| message: false
#| warning: false
p_GII <- ggplot(GII, aes(x = TIME_PERIOD, y = GII_value, color = iso3, group = iso3)) +
  geom_line(linewidth = 1) +
  labs(
    title = "GII per Country (EU27)",
    x = "Year",
    y = "Gender inequality index",
    color = "Country"
  ) +
  scale_y_continuous(limits = c(0, 1)) +   
  theme_minimal()+
  theme(
    legend.position = "right",
    legend.text = element_text(size = 6), 
    legend.title = element_text(face = "bold"),
    legend.key.height = unit(0.3, "cm"),
    legend.key.width = unit(0.4, "cm"),
    legend.spacing.y = unit(0.05, "cm"), 
    plot.margin = margin(10, 10, 10, 10) 
  )

if (knitr::is_html_output()) {plotly::ggplotly(p_GII)} else {p_GII}
```

```{r, echo=FALSE}
summary_stats_GII <- EU27_GII |>
  summarise(
    Min = min(GII_value, na.rm = TRUE),
    Max = max(GII_value, na.rm = TRUE),
    Mean = mean(GII_value, na.rm = TRUE),
    Median = median(GII_value, na.rm = TRUE),
    SD = sd(GII_value, na.rm = TRUE)
    )

summary_stats_GII |>
  mutate(across(where(is.numeric), ~ formatC(.x, format = "f", digits = 2))) |>
  kable(
    caption = "Descriptive statistics of the GII average"
  )
```

The Gender Inequality Index (GII) shows a clear downward trend across all European countries between 1990 and 2024. The EU27 average decreased from 25,7% to 0,08%, with a mean of 16,6% overall. 

## 6. Coefficient of human inequality per country ^[For a more interactive and readable view of the graph in this section, please refer to the *"dm-project phase 3.html"* document.]

```{r, echo=FALSE}
#| message: false
#| warning: false
p_coef <- ggplot(coef_ineq, aes(x = TIME_PERIOD, y = coef_ineq_value, color = iso3, group = iso3)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Coefficient of human inequality per Country (EU27)",
    x = "Year",
    y = "Coefficient of human inequality",
    color = "Country"
  ) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 15))+
  theme(
    legend.position = "right",
    legend.text = element_text(size = 6), 
    legend.title = element_text(face = "bold"),
    legend.key.height = unit(0.3, "cm"),
    legend.key.width = unit(0.4, "cm"),
    legend.spacing.y = unit(0.05, "cm"), 
    plot.margin = margin(10, 10, 10, 10) 
  )

if (knitr::is_html_output()) {plotly::ggplotly(p_coef)} else {p_coef}
```

```{r, echo=FALSE}
summary_stats_coefineq <- EU27_Coefineq |>
  summarise(
    Min = min(coef_ineq_value, na.rm = TRUE),
    Max = max(coef_ineq_value, na.rm = TRUE),
    Mean = mean(coef_ineq_value, na.rm = TRUE),
    Median = median(coef_ineq_value, na.rm = TRUE),
    SD = sd(coef_ineq_value, na.rm = TRUE)
    )

summary_stats_coefineq |>
  mutate(across(where(is.numeric), ~ formatC(.x, format = "f", digits = 2))) |>
  kable(
    caption = "Descriptive statistics of the coefficient of human inequality average")
```

The coefficient of human inequality varies widely across Europe. The EU27 average sits at $≈$ 8.64%.

{{< pagebreak >}}

# III. Linear Regressions

## 1. Between taxes and GHG emissions

We answer one of our fundamental question: "Do taxes reduce GHG emissions ?".

First we evaluate the coefficient between an increase in environmental taxes in the year n and the impact on GHG emissions in the same year n.

```{r, echo=FALSE}
#| message: false
#| warning: false
model <- lm(GHG_Emissions_value ~ Envt_Tax_Revenues_value + GDP_value + TIME_PERIOD, data = all_merged)

stargazer(model, type = "text", 
          title = "Regression Results",
          digits = 3, 
          single.row = TRUE)
```

We estimated a linear regression to examine the relationship between environmental tax revenues and greenhouse gas emissions, while controlling for GDP and year. The model fits very well (Adjusted $R^2$ = 0.8456). Environmental tax revenues have a negative coefficient (≈ −1.03), indicating that higher environmental taxes are associated with lower emissions. The effect is statistically significant at the 1% level  (p \< 0.001). GDP enters positively,  consistently with scale effects: larger economies produce more, and thus emit more; time has a negative coefficient, both highly significant.
We do not calculate a log - level regression, because our following analysis and tax proposition will be made accordingly to levels of GHG.


Secondly, we try to evaluate the coefficient between an increase in taxes in the year n and the impact on ghg emissions in the following years n+1, to n+5.

{{< pagebreak >}}

```{r, echo=FALSE}
#| message: false
#| warning: false
# Create leads 1 to 5
for (k in 1:5) {
  GHG_Emissions[[paste0("GHG_Emissions_value.lead", k)]] <- ave(
    GHG_Emissions$GHG_Emissions_value, 
    GHG_Emissions$iso3, 
    FUN = function(x) c(x[(1+k):length(x)], rep(NA, k)))}

# add them in all merged
all_merged_plus_leads <- merge(all_merged, GHG_Emissions,
  by = c("iso3", "TIME_PERIOD", "GHG_Emissions_value"),
  all.x = TRUE)
all_merged_plus_leads <- all_merged_plus_leads |>
    select(iso3, TIME_PERIOD, GDP_value, Envt_Tax_Revenues_value, GHG_Emissions_value, GHG_Emissions_value.lead1, GHG_Emissions_value.lead2, GHG_Emissions_value.lead3, GHG_Emissions_value.lead4, GHG_Emissions_value.lead5 )

# Create the list of formulas for leads 1 to 5
formulas <- lapply(1:5, function(k) {
  as.formula(
    paste0("GHG_Emissions_value.lead", k, 
           " ~ Envt_Tax_Revenues_value + GDP_value + TIME_PERIOD")
  )
})

# Run all regressions
reg_results <- lapply(formulas, function(f) lm(f, data = all_merged_plus_leads))

# Display all 5 regressions in a single stargazer table
add_stars <- function(p) {
  case_when(
    p < 0.01 ~ "***",
    p < 0.05 ~ "**",
    p < 0.1  ~ "*",
    TRUE     ~ ""
  )
}

# Create tidy tables for all 5 lead regressions
tidy_list <- lapply(1:5, function(k) {
  broom::tidy(reg_results[[k]]) |>
    mutate(
      Lead = paste0("Lead ", k),
      Estimate = formatC(estimate, format = "e", digits = 3),
      `Std. Error` = formatC(std.error, format = "e", digits = 3),
      `t value` = formatC(statistic, format = "e", digits = 3),
      `p value` = formatC(p.value, format = "e", digits = 3),
      Signif = add_stars(p.value)
    ) |>
    select(Lead, term, Estimate, `Std. Error`, `t value`, `p value`, Signif)
})

# Combine all leads into one data frame
tidy_df <- bind_rows(tidy_list)

# Rename variable names for readability
tidy_df$term <- gsub("_value", "", tidy_df$term)

# Print table
kable(
  tidy_df,
  caption = "Regression Results for GHG Emissions Leads 1-5\n*p<0.1; **p<0.05; ***p<0.01",
  align = "lcccccc"
)
```

These lead regressions use GDP and a time trend as controls and all show a very strong fit, with adjusted R² remaining high (around 0.84) across all horizons.

Across the five specifications, the coefficient on environmental tax revenues is negative, statistically significant, and stable. It goes from –1.516 at horizon n+1 to approximately -2.0 at n+2 to n+5. This pattern suggests that higher environmental tax revenues are systematically associated with lower future GHG emissions.

All other coefficients are close to the ones seen before (GDP around 0.25, and time has coefficients -4.5, all highly significant).

## 2. Between taxes and social inequalities

We estimate three regressions to evaluate the association between environmental tax revenues and three social outcomes, in percentage: energy poverty, human inequality, and gender inequality. In all models, we also add GDP and time (years). We chose to not include the Gender Inequality index and the Income inequality coefficient in one each others regression because of collinearity problems. Also, the coefficient of energy poverty is rather a consequence of both than a cause, so we do not include it in the other two as well.

{{< pagebreak >}}

```{r, echo=FALSE}
#| message: false
#| warning: false
# taxes on inability to warm house (economic inequalities)
model_warm_tax <- lm(Warm_Households_value ~ Envt_Tax_Revenues_value +  GII_value + coef_ineq_value + hhtyp + incgrp + log(GDP_value) + TIME_PERIOD, data = all_merged)

stargazer(model_warm_tax, type = "text", 
          title = "Regression Results",
          digits = 3, 
          single.row = TRUE)
```

In the model explaining energy poverty, we include the income-inequality coefficient. The resulting model has strong explanatory power (Adjusted $R^2 ≈$ 0.60). The coefficient on environmental tax revenues remains small, positive, stable (+3.7e-05) and significant. Also, the income-inequality coefficient shows strong positive and significant associations with energy poverty.

{{< pagebreak >}}

```{r, echo=FALSE}
#| message: false
#| warning: false
# taxes on the coefficient of inequalities
model_ineq_tax <- lm(coef_ineq_value ~ Envt_Tax_Revenues_value + log(GDP_value) + TIME_PERIOD, data = all_merged)

stargazer(model_ineq_tax, type = "text", 
          title = "Regression Results",
          digits = 3, 
          single.row = TRUE)
```

For human inequality, the model remains statistically valid but with limited explanatory power (Adjusted $R^2 ≈$ 0.05), and environmental taxes have a small impact on inequality (+3.0e-05) but is highly significant. GDP and time both show negative effects.

```{r, echo=FALSE}
#| message: false
#| warning: false
# taxes on the gender inequality index
model_gii_tax <- lm(GII_value ~ Envt_Tax_Revenues_value + log(GDP_value) + TIME_PERIOD, data = all_merged)

stargazer(model_gii_tax, type = "text", 
          title = "Regression Results",
          digits = 3, 
          single.row = TRUE)
```

Similarly, in the gender-inequality model the environmental-tax coefficient remains positive (+6.9e-07) and highly significant. GDP and time effects are negative and the model is substantially strong (Adjusted $R^2 ≈$ 0.40).

## 3. Between tax and economy

```{r, echo=FALSE}
#| message: false
#| warning: false
model_gdp_tax <- lm(log(GDP_value) ~ Envt_Tax_Revenues_value + TIME_PERIOD, data = all_merged)

stargazer(model_gdp_tax, type = "text", 
          title = "Regression Results",
          digits = 3, 
          single.row = TRUE)
```
The model shows that environmental tax revenues are positively associated with GDP, and the relationship is highly statistically significant, although the effect size is very small ($≈$ 0.0001 per unit of tax revenue). The time trend is positive and highly significant, reflecting steady GDP growth over the period. 

It also has substantial explanatory power (Adjusted $R^2 ≈$ 0.63). However, the relationship of GDP with taxes is correlational and does not imply causation, it indicates that environmental tax revenues tend to increase alongside GDP, consistent with the idea that wealthier economies can collect more from climate-related taxes.

{{< pagebreak >}}

# IV. Forecasting: no-policy

We forecast the impact of environmental taxes on various variables, in a scenario where no policy is implemented.

<<<<<<< HEAD
### 1. Greenhouse Gas emissions

```{r, echo=FALSE}
#| message: false
#| warning: false
EUGHG_agg_yearly <- EU27_GHG |>
  filter(!is.na(GHG_Emissions_value)) |>
  group_by(TIME_PERIOD) |>
  summarise(GHG_total = sum(GHG_Emissions_value, na.rm = TRUE))

ts_data <- ts(EUGHG_agg_yearly$GHG_total, start=min(EUGHG_agg_yearly$TIME_PERIOD))
model <- auto.arima(ts_data)
fc <- forecast(model, h=27)

fc_df <- as.data.frame(fc)
head(fc_df, 1)
tail(fc_df, 1)

print(model)

```

```{r, echo=FALSE}
#| message: false
#| warning: false
autoplot(fc)
```

This ARIMA-based forecast suggests a slow but steady decline in total EU27 greenhouse gas emissions from 2024 to 2050. The ARIMA(0,1,0) with drift implies an average annual reduction of about 52 million tonnes of CO$₂$-equivalent, with emissions falling from roughly 2.85 billion tonnes in 2024 to around 1.49 billion tonnes in 2050. 

Despite this downward trend, projected emissions in 2050 remain far above zero—even in the most optimistic scenario—indicating that current trends are insufficient to reach EU net-zero targets (about 363 million tonnes of CO$₂$-equivalent).

### 2. Social inequalities:

```{r, echo=FALSE}
#| message: false
#| warning: false

gii_ts <- ts(EU27_GII$GII_value, start = min(EU27_GII$TIME_PERIOD))
gii_model <- auto.arima(gii_ts)
gii_fc <- forecast(gii_model, h = 27)

coefineq_ts <- ts(EU27_Coefineq$coef_ineq_value, start = min(EU27_Coefineq$TIME_PERIOD))
coefineq_model <- auto.arima(coefineq_ts)
coefineq_fc <- forecast(coefineq_model, h = 27)

EU27_Warming_agg <- EU27_Warming |> 
  filter(!is.na(Warm_Households_value)) |> 
  group_by(TIME_PERIOD) |> 
  summarise(Warm_avg = mean(Warm_Households_value, na.rm = TRUE))

warm_ts <- ts(EU27_Warming_agg$Warm_avg, start = min(EU27_Warming_agg$TIME_PERIOD))
warm_model <- auto.arima(warm_ts)
warm_fc <- forecast(warm_model, h = 26)

gii_fc1 <- as.data.frame(gii_fc)
head(gii_fc1, 1)
tail(gii_fc1, 1)

coefineq_fc1 <- as.data.frame(gii_fc)
head(coefineq_fc1, 1)
tail(coefineq_fc1, 1)

warm_fc1 <- as.data.frame(gii_fc)
head(warm_fc1, 1)
tail(warm_fc1, 1)
```


```{r, echo=FALSE}
#| message: false
#| warning: false
plot(gii_fc, main = "Forecast of GII")
plot(coefineq_fc, main = "Forecast of Coefficient of Inequality")
plot(warm_fc, main = "Forecast of Warming Inability")
```

The GII is forecast to decline from 0.085 in 2023 to zero by 2050 (with negative values cut, as the index ranges from 0 to 1). 

Human inequality is expected to fall from 7.73 to about 5.04 over the same period, suggesting gradual improvement, though uncertainty increases over time. 

In contrast, the share of households unable to heat their homes is projected to remain broadly stable, edging up slightly from 14.65% in 2024 to 16.41% in 2050, with wide confidence intervals.

### 3. The economy: forecast of the GDP

```{r, echo=FALSE}
#| message: false
#| warning: false
gdp_ts <- ts(EU27_GDP$GDP_value, start = min(EU27_GDP$TIME_PERIOD)-1)
gdp_model <- auto.arima(gdp_ts)
gdp_fc <- forecast(gdp_model, h = 28)

plot(gdp_fc, main = "Forecast of GDP")
```

The EU27 aggregate GDP is projected to grow steadily from 19.4 trillion USD in 2024 to 29.5 trillion USD in 2050, representing approximately 50% growth over the period, though the widening confidence intervals indicate increasing uncertainty in long-term projections.

### 4. The economy: forecast of the environmental taxes {#forecastofenvttax}
=======
## 1. Forecasted environmental tax revenues 
>>>>>>> 7153cd54706bf27da4f51942cadb3c8f64d6ff4e

```{r, echo=FALSE}
#| message: false
#| warning: false
# Linear trend model
tax_lm <- lm(Envt_Tax_Revenues_value ~ TIME_PERIOD, data = EU27_Tax)
stargazer(tax_lm, type = "text", 
          title = "Regression Results",
          digits = 3, 
          single.row = TRUE)

# Forecast
future_years <- tibble(
  TIME_PERIOD = (max(EU27_Tax$TIME_PERIOD)+1):(max(EU27_Tax$TIME_PERIOD)+27)
)

tax_pred_95 <- predict(
  tax_lm,
  newdata = future_years,
  interval = "confidence", 
  level = 0.95
)

tax_pred_80 <- predict(
  tax_lm,
  newdata = future_years,
  interval = "confidence", 
  level = 0.80
)

tax_fc_lm <- bind_cols(
  future_years, 
  as.data.frame(tax_pred_95),
  lwr_80 = as.data.frame(tax_pred_80)$lwr,
  upr_80 = as.data.frame(tax_pred_80)$upr
)

```

```{r, echo=FALSE}
#| message: false
#| warning: false
tax_all_2050 <- ggplot() +
  geom_line(data = EU27_Tax, 
            aes(x = TIME_PERIOD, y = Envt_Tax_Revenues_value, color = "Observations"), linewidth = 0.5) +
  geom_ribbon(data = tax_fc_lm, 
              aes(x = TIME_PERIOD, ymin = lwr, ymax = upr, fill = "95% confidence interval"), alpha = 0.4) +
  geom_ribbon(data = tax_fc_lm, 
              aes(x = TIME_PERIOD, ymin = lwr_80, ymax = upr_80, fill = "80% confidence interval"), alpha = 0.6) +
  geom_line(data = tax_fc_lm, 
            aes(x = TIME_PERIOD, y = fit, color = "Forecast"), linewidth = 0.5) +
  scale_color_manual(name = "Lines",
                     values = c("Observations" = "black", 
                                "Forecast" = "darkblue")) +
  scale_fill_manual(name = "Confidence intervals",
                    values = c("80% confidence interval" = "#6666FF", 
                               "95% confidence interval" = "#6666FF")) +
  guides(fill = guide_legend(order = 2), color = guide_legend(order = 1)) +
  labs(
    title = "No-policy forecasted European Environmental\ntax revenues",
    x = "Year",
    y = "Tax revenues (millions of USD)"
  ) +
  scale_y_continuous(limits = c(0, 0.75e+06)) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0),
    axis.title = element_text(face = "bold", size = 12),
    plot.margin = margin(t = 40, r = 10, b = 10, l = 10),
    legend.position = "right",       
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10)
  )

if (knitr::is_html_output()) {plotly::ggplotly(tax_all_2050)} else {tax_all_2050}
```

If no new tax policy is implemented, environmental tax revenues in the EU27 are projected to increase from 362.4 billion USD in 2023 to 617.4 billion USD in 2050, representing a 70% growth based on an extrapolation of current trends. This steady upward trajectory suggests continued expansion of green taxation policies, however, it's important to note that this forecast assumes the continuation of existing trends under a no-change-in-policy scenario.

## 2. Greenhouse Gas emissions

```{r, echo=FALSE}
#| message: false
#| warning: false
EUGHG_agg_yearly <- EU27_GHG |>
  filter(!is.na(GHG_Emissions_value)) |>
  group_by(TIME_PERIOD) |>
  summarise(GHG_total = sum(GHG_Emissions_value, na.rm = TRUE))

ts_data <- ts(EUGHG_agg_yearly$GHG_total, start=min(EUGHG_agg_yearly$TIME_PERIOD))
model <- auto.arima(ts_data)
fc <- forecast(model, h=27)

fc_df <- as.data.frame(fc)
fc_df |>
  slice(c(1, n()))

# Check if coefficients exist
if(length(coef(model)) > 0){
  coefs <- coef(model)
  se <- sqrt(diag(model$var.coef))
  
  coef_df <- data.frame(
    Term = names(coefs),
    Estimate = formatC(coefs, format="e", digits=3),
    `Std. Error` = formatC(se, format="e", digits=3)
  )
} else {
  coef_df <- data.frame(
    Term = character(0),
    Estimate = character(0),
    `Std. Error` = character(0)
  )
}

# Model-level statistics
stats_df <- data.frame(
  Term = c("sigma^2", "log likelihood", "AIC", "AICc", "BIC"),
  Estimate = c(
    formatC(model$sigma2, format="e", digits=3),
    round(model$loglik,2),
    round(model$aic,2),
    round(model$aicc,2),
    round(model$bic,2)
  ),
  `Std. Error` = ""
)

# Combine
final_arima_table <- bind_rows(coef_df, stats_df)
rownames(final_arima_table) <- NULL

# Print nicely
kable(final_arima_table, caption = "ARIMA(0,1,0) with drift: Summary")
```

```{r, echo=FALSE}
#| message: false
#| warning: false
autoplot(fc)+
  ggtitle("No-policy forecasted European GHG emissions",
          subtitle = "EU−27, in thousand tonnes of CO2") +
  xlab("Year") +
  ylab("Emissions") +
  theme_minimal()+
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(face = "bold", size = 12), 
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)  
  )
```

This ARIMA-based forecast suggests a slow but steady decline in total EU27 greenhouse gas emissions from 2024 to 2050. The ARIMA(0,1,0) with drift implies an average annual reduction of about 52 million tonnes of CO$₂$-equivalent, with emissions falling from roughly 2.85 billion tonnes in 2024 to around 1.49 billion tonnes in 2050. 

Despite this downward trend, projected emissions in 2050 remain far above zero—even in the most optimistic scenario—indicating that current trends are insufficient to reach EU net-zero targets (about 363 million tonnes of CO$₂$-equivalent).

## 3. Social inequalities:

#### a. Gender Inequality Index
```{r, echo=FALSE}
#| message: false
#| warning: false
gii_ts <- ts(EU27_GII$GII_value, start = min(EU27_GII$TIME_PERIOD))
gii_model <- auto.arima(gii_ts)
gii_fc <- forecast(gii_model, h = 27)
gii_df <- as.data.frame(gii_fc)
gii_df |>
  slice(c(1, n()))

autoplot(gii_fc) +
  ggtitle("No-policy forecasted European GII") +
  xlab("Year") +
  ylab("Gender Inequality Index") +
  theme_minimal()+
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(face = "bold", size = 12), 
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)  
  )
```

The GII is forecast to decline from 0.079 in 2024 to zero by 2050 (with negative values cut, as the index ranges from 0 to 1). 

{{< pagebreak >}}

#### b. Coefficient of Human Inequality

```{r, echo=FALSE}
#| message: false
#| warning: false
coefineq_ts <- ts(EU27_Coefineq$coef_ineq_value, start = min(EU27_Coefineq$TIME_PERIOD))
coefineq_model <- auto.arima(coefineq_ts)
coefineq_fc <- forecast(coefineq_model, h = 27)
coefineq_df <- as.data.frame(coefineq_fc)
coefineq_df |>
  slice(c(1, n()))

autoplot(coefineq_fc) +
  ggtitle("No-policy forecasted European Coefficient\nof human inequality") +
  xlab("Year") +
  ylab("Coefficient of human inequality") +
  theme_minimal()+
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(face = "bold", size = 12), 
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)  
  )
```

Human inequality is expected to fall from 7.63 to about 5.04 over the same period, suggesting gradual improvement, though uncertainty increases over time. 

{{< pagebreak >}}

#### c. Energy poverty

```{r}
EU27_Warming_agg <- EU27_Warming |> 
  filter(!is.na(Warm_Households_value)) |> 
  group_by(TIME_PERIOD) |> 
  summarise(Warm_avg = mean(Warm_Households_value, na.rm = TRUE))

warm_ts <- ts(EU27_Warming_agg$Warm_avg, start = min(EU27_Warming_agg$TIME_PERIOD))
warm_model <- auto.arima(warm_ts)
warm_fc <- forecast(warm_model, h = 26)
warm_df <- as.data.frame(warm_fc)
warm_df |>
  slice(c(1, n()))

autoplot(warm_fc) +
  ggtitle("No-policy forecasted European Inability\nto keep the House Warm") +
  xlab("Year") +
  ylab("Percentage of population") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle = element_text(size = 12, hjust = 0),
    axis.title = element_text(face = "bold", size = 12),
    plot.margin = margin(t = 40, r = 10, b = 10, l = 10)
  )
```

In contrast, the share of households unable to heat their homes is projected to remain broadly stable, edging up slightly from 14.15% in 2024 to 16.41% in 2050, with wide confidence intervals.

{{< pagebreak >}}

## 3. Gross Domestic Product

```{r, echo=FALSE}
#| message: false
#| warning: false
gdp_ts <- ts(EU27_GDP$GDP_value, start = min(EU27_GDP$TIME_PERIOD)-1)
gdp_model <- auto.arima(gdp_ts)
gdp_fc <- forecast(gdp_model, h = 28)
gdp_df <- as.data.frame(gdp_fc)
gdp_df |>
  slice(c(1, n()))

autoplot(gdp_fc) +
  ggtitle("No-policy forecasted European GDP",
          subtitle = "in millions of USD") +
  xlab("Year") +
  ylab("GDP") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle = element_text(size = 12, hjust = 0),
    axis.title = element_text(face = "bold", size = 12),
    plot.margin = margin(t = 40, r = 10, b = 10, l = 10)
  )
```

The EU27 aggregate GDP is projected to grow steadily from 19.4 trillion USD in 2024 to 29.5 trillion USD in 2050, representing approximately 50% growth over the period, though the widening confidence intervals indicate increasing uncertainty in long-term projections.

{{< pagebreak >}}

# V. Forecasting: net-zero emission target

We are trying to evaluate the needed level of tax to reach the Zero net emissions by 2050 desired by the European Union, as if there only was a tax policy available; then we check the impact of this level of tax on the social inequalities and on the economy.

## 1. Deriving a realistic tax path to reach net-zero by 2050

### a. The forecast model

We use a linear regression relating aggregate GHG emissions to Environmental tax revenues at the EU-27 level. The 2023 tax level is anchored to the last observed value. 

```{r, echo=FALSE}
#| message: false
#| warning: false
data_forecast_tax_0net <- EUGHG_agg_yearly |>
  left_join(EU27_Tax, by = "TIME_PERIOD") |>
  filter(!is.na(GHG_total), !is.na(Envt_Tax_Revenues_value)) |>
  mutate(year = as.numeric(TIME_PERIOD))

# Estimate level-level model 
mod_LL <- lm(GHG_total ~ Envt_Tax_Revenues_value, data = data_forecast_tax_0net)

stargazer(mod_LL, type = "text", 
          title = "Regression Results",
          digits = 3, 
          single.row = TRUE)
```

The estimated regression reveals a statistically significant negative relationship between environmental tax revenues and total GHG emissions (−3.86, p-value < 0.001). This indicates that, historically, an increase of one million USD in environmental tax revenues is associated with a reduction of approximately 3.9 thousand tonnes of CO$₂$-equivalent in aggregate emissions. The model explains a meaningful share of the observed variation in emissions, with an Adjusted $R^2$ of 0.49.

```{r}
# If beta >= 0, stop (no reduction possible by raising tax)
if(coef(mod_LL)[2] >= 0) stop("Model does not indicate that raising tax reduces GHG.")

# Build linear target path (absolute decline) as before
years <- seq((max(data_forecast_tax_0net$year)+1), 2050)
GHG_last <- tail(data_forecast_tax_0net$GHG_total, 1)
target_GHG <- seq(GHG_last, 1e-6, length.out = length(years))
 
# Invert model: Tax_t = (target - alpha) / beta
tax_required <- (target_GHG - coef(mod_LL)[1]) / coef(mod_LL)[2]
tax_required <- ifelse(tax_required < 0 | tax_required > 1e12, NA_real_, tax_required)

tax_path <- tibble(
  year = years,
  target_GHG = target_GHG,
  tax_required = tax_required,
  pred_GHG = predict(mod_LL, newdata = tibble(Envt_Tax_Revenues_value = tax_required))
)


tax_path <- tax_path |>
  mutate(pred_GHG = predict(mod_LL, newdata = tibble(Envt_Tax_Revenues_value = tax_required)))

ggplot(tax_path, aes(x = year)) +
  geom_line(aes(y = pred_GHG), color = "darkblue") +
  labs(title="Targeted GHG (level-level)", y="GHG", x="Year") +
  theme_minimal()+
  scale_y_continuous(limits = c(0, 4e+06))

ggplot(tax_path, aes(x = year)) +
  geom_line(aes(y = tax_required), color = "darkblue") +
  labs(title="Required tax path (level-level model)", y="Tax (millions USD)", x="Year") +
  theme_minimal()+
  scale_y_continuous(limits = c(0, 2e+06))
```

Using this estimated relationship, we construct a counterfactual emissions trajectory from approximately 2.9 billion tonnes in 2023 to net-zero in 2050. After 2024, the tax path follows a linear trajectory toward 2050, reflecting the progressively stronger fiscal effort needed to reduce emissions to zero.

According to the tax path, environmental tax revenues would need to rise from 550 billion USD in 2024 to about 1.3 trillion USD in 2050, highlighting both the scale of the fiscal effort required and the limitations of relying solely on environmental taxation, given the modest explanatory power of the reduced-form regression.

### b. Comparing with the no-policy forecast

We evaluate the gap between historical trends and climate targets before assessing the implications of the required tax paths.

#### i. Tax policy: targeted vs predicted

To assess the adequacy of the policy, we compare the forecasted path of total environmental tax revenues based on historical trends, with the one required to reach net-zero emissions by 2050.

```{r, echo=FALSE}
#| message: false
#| warning: false
tax_fc_lm <- tax_fc_lm |> rename(year = TIME_PERIOD)

tax_plot_df <- tax_fc_lm |>
  left_join(
    tax_path |> select(year, tax_required),
    by = "year"
  )

ggplot(tax_plot_df, aes(x = year)) +
  geom_ribbon(
    aes(ymin = lwr, ymax = upr, fill = "95% confidence interval"),
    alpha = 0.35
  ) +
  geom_ribbon(
    aes(ymin = lwr_80, ymax = upr_80, fill = "80% confidence interval"),
    alpha = 0.5
  ) +
  geom_line(
    aes(y = fit, color = "No-policy forecasted tax revenues"),
    linewidth = 1
  ) +
  geom_line(
    aes(y = tax_required, color = "Targeted tax path for net-zero by 2050"),
    linewidth = 1
  ) +
  scale_color_manual(
    name = "Lines",
    values = c(
      "No-policy forecasted tax revenues" = "darkblue",
      "Targeted tax path for net-zero by 2050" = "red"
    )
  ) +
  scale_fill_manual(
    name = "Confidence intervals",
    values = c(
      "80% confidence interval" = "#6666FF",
      "95% confidence interval" = "#9999FF"
    )
  ) +
  labs(
    title = "Environmental Tax Revenues: Trend Forecast vs\nClimate-Required Path",
    x = "Year",
    y = "Tax revenues (millions USD)"
  ) +
  scale_y_continuous(limits = c(0, 1.5e6)) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )
```

Comparing the two paths shows a significant and growing shortfall. The historical trend regression predicts that total environmental tax revenues would increase by approximately 7.3 billion USD per year and in contrast, the required one to achieve net-zero calls for an annual increase of roughly 28.9 billion USD per year. From the first year (2024), forecasted revenues fall well below the required level, and by 2050 the gap widens to roughly 684 billion USD. These results highlight that in order to achieve the net-zero emissions target, environmental taxation would need to be increased rapidly.

#### ii. GHG emissions: targeted vs predicted

We can compare the emissions trajectory implied by a continuation of historical tax trends (ARIMA seen before) with the counter-factual path required to achieve net-zero emissions by 2050.

```{r, echo=FALSE}
#| message: false
#| warning: false
fc_df <- data.frame(
  year = as.numeric(time(fc$mean)),
  mean = as.numeric(fc$mean),
  lower80 = as.numeric(fc$lower[, 1]),
  upper80 = as.numeric(fc$upper[, 1]),
  lower95 = as.numeric(fc$lower[, 2]),
  upper95 = as.numeric(fc$upper[, 2])
)

ggplot(tax_path, aes(x = year)) +
  geom_ribbon(data = fc_df, aes(
      x = year,
      ymin = lower95,
      ymax = upper95,
      fill = "95% confidence interval"),
    inherit.aes = FALSE,
    alpha = 0.35) +
  geom_ribbon(data = fc_df, aes(
      x = year,
      ymin = lower80,
      ymax = upper80,
      fill = "80% confidence interval"),
    inherit.aes = FALSE,
    alpha = 0.5) +
  geom_line(data = fc_df, aes(
    x = year, y = mean, color = "No-policy forecasted European GHG emissions"),
    inherit.aes = FALSE,
    linewidth = 1) +
  geom_line(
    aes(y = target_GHG, color = "Targeted GHG emissions for net-zero by 2050"),
    linewidth = 1) +
  scale_color_manual(
    name = "Lines",
    values = c(
      "Targeted GHG emissions for net-zero by 2050" = "red",
      "No-policy forecasted European GHG emissions" = "darkblue")) +
  scale_fill_manual(
    name = "Confidence intervals",
    values = c(
      "80% confidence interval" = "#6666FF",
      "95% confidence interval" = "#9999FF")) +
  labs(
    title = "Greenhouse gases emissions: \nTrend Forecast vs Climate-Required Path",
    x = "Year",
    y = "GHG Emissions") +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )
```

As the graph shows, achieving the GHG emissions target would require significant, but not dramatic, changes to current tax practices. From around 2035 onwards, the required tax-path falls outside the 95% confidence interval, indicating that the probability of reaching the targeted emissions level is very low.

## 2. Is that policy fair? Its impact on social inequalities

### a. The forecast model

We project the future evolution of our three inequality indicators: average warming inability, the coefficient of human inequality, and the Gender Inequality Index (GII). 

Starting from their last observed values, we simulate paths up to 2050 by combining the estimated marginal effects of environmental tax revenues and time trends obtained from the regressions. The required path is expressed as an average annual increase, applied uniformly over time. The projections are established by applying these marginal effects cumulatively year after year. For the GII, predicted values are constrained to remain within 0 and 1.

```{r, echo=FALSE}
#| message: false
#| warning: false
# Average yearly tax increase
avg_tax_increment <- (last(tax_path$tax_required) -
                      first(tax_path$tax_required)) /
                      (nrow(tax_path) - 1)

# Last observed levels
last_obs <- tibble(
  Warm = tail(EU27_Warming_agg$Warm_avg, 1),
  Ineq = tail(EU27_Coefineq$coef_ineq_value, 1),
  GII  = tail(EU27_GII$GII_value, 1)
)

future_years <- length(tax_path$year)

# Extract coefficients
b_warm_tax  <- coef(model_warm_tax)["Envt_Tax_Revenues_value"]
b_warm_time <- coef(model_warm_tax)["TIME_PERIOD"]

b_ineq_tax  <- coef(model_ineq_tax)["Envt_Tax_Revenues_value"]
b_ineq_time <- coef(model_ineq_tax)["TIME_PERIOD"]

b_gii_tax   <- coef(model_gii_tax)["Envt_Tax_Revenues_value"]
b_gii_time  <- coef(model_gii_tax)["TIME_PERIOD"]

# Build paths
future_predictions <- tibble(
  year = tax_path$year,

  Predicted_Warming =
    last_obs$Warm +
    cumsum(rep(b_warm_tax * avg_tax_increment + b_warm_time, future_years)),

  Predicted_Inequality =
    last_obs$Ineq +
    cumsum(rep(b_ineq_tax * avg_tax_increment + b_ineq_time, future_years)),

  Predicted_GII =
    pmax(0, pmin(1,last_obs$GII +
        cumsum(rep(b_gii_tax * avg_tax_increment + b_gii_time, future_years))
      )))

future_predictions |>
  slice(c(1, n())) |>
  kable(digits = 3, col.names = c("Year", "Predicted Warming", "Predicted Inequality", "Predicted GII"))
```

Projections show that the required trajectory significantly increases the three inequality indicators over the period 2024-2050. Indeed, the average warming inability rises from 15.8% of the population in 2024 to 44.8% in 2050, the human inequality coefficient increases from around 8.5 to 29.2, indicating a considerable widening of social disparities. Similarly, the gender inequality index will increase from around 0.10 to around 0.53.

### b. Comparing with the no-policy forecast

To assess how the proposed environmental tax policy affects social inequalities relative to a baseline scenario without policy intervention, we compare the projected evolution of each indicator under the tax trajectory with their ARIMA-based forecasts in the absence of environmental taxation.

```{r, echo=FALSE}
#| message: false
#| warning: false

## WARM PLOT
warm_fc_df <- data.frame(
  year = as.numeric(time(warm_fc$mean)),
  mean = as.numeric(warm_fc$mean),
  lower80 = as.numeric(warm_fc$lower[, 1]),
  upper80 = as.numeric(warm_fc$upper[, 1]),
  lower95 = as.numeric(warm_fc$lower[, 2]),
  upper95 = as.numeric(warm_fc$upper[, 2])
)

p_warming <- ggplot() +
  geom_ribbon(
    data = warm_fc_df,
    aes(x = year, ymin = lower95, ymax = upper95, fill = "95% confidence interval"),
    alpha = 0.35
  ) +
  geom_ribbon(
    data = warm_fc_df,
    aes(x = year, ymin = lower80, ymax = upper80, fill = "80% confidence interval"),
    alpha = 0.5
  ) +
  geom_line(
    data = warm_fc_df,
    aes(x = year, y = mean, color = "No-policy forecasted Average Warming Inability"),
    linewidth = 1
  ) +
  geom_line(
    data = future_predictions,
    aes(x = year, y = Predicted_Warming, color = "Net-zero policy impact on Average Warming Inability"),
    linewidth = 1
  ) +
  scale_color_manual(
    name = "Scenarios",
    values = c(
      "No-policy forecasted Average Warming Inability" = "darkblue",
      "Net-zero policy impact on Average Warming Inability" = "darkgreen"
    )
  ) +
  scale_fill_manual(
    name = "Confidence intervals",
    values = c(
      "80% confidence interval" = "#6666FF",
      "95% confidence interval" = "#9999FF"
    )
  ) +
  labs(
    title = "Average Warming Inability: \nTax Policy vs No-Policy Scenario",
    x = "Year",
    y = "Average Warming Inability"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )



## INEQ PLOT 
ineq_fc_df <- data.frame(
  year = as.numeric(time(coefineq_fc$mean)),
  mean = as.numeric(coefineq_fc$mean),
  lower80 = as.numeric(coefineq_fc$lower[, 1]),
  upper80 = as.numeric(coefineq_fc$upper[, 1]),
  lower95 = as.numeric(coefineq_fc$lower[, 2]),
  upper95 = as.numeric(coefineq_fc$upper[, 2])
)

p_inequality <- ggplot() +
  geom_ribbon(
    data = ineq_fc_df,
    aes(x = year, ymin = lower95, ymax = upper95, fill = "95% confidence interval"),
    alpha = 0.35
  ) +
  geom_ribbon(
    data = ineq_fc_df,
    aes(x = year, ymin = lower80, ymax = upper80, fill = "80% confidence interval"),
    alpha = 0.5
  ) +
  geom_line(
    data = ineq_fc_df,
    aes(x = year, y = mean, color = "No-policy forecasted Coefficient of Human Inequality"),
    linewidth = 1
  ) +
  geom_line(
    data = future_predictions,
    aes(x = year, y = Predicted_Inequality, color = "Net-zero policy impact on the Coefficient of Human Inequality"),
    linewidth = 1
  ) +
  scale_color_manual(
    name = "Scenarios",
    values = c(
      "No-policy forecasted Coefficient of Human Inequality" = "darkblue",
      "Net-zero policy impact on the Coefficient of Human Inequality" = "orange"
    )
  ) +
  scale_fill_manual(
    name = "Confidence intervals",
    values = c(
      "80% confidence interval" = "#6666FF",
      "95% confidence interval" = "#9999FF"
    )
  ) +
  labs(
    title = "Coefficient of Human Inequality: \nTax Policy vs No-Policy Scenario",
    x = "Year",
    y = "Coefficient of Human Inequality"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )


## GII PLOT
gii_fc_df <- data.frame(
  year = as.numeric(time(gii_fc$mean)),
  mean = as.numeric(gii_fc$mean),
  lower80 = as.numeric(gii_fc$lower[, 1]),
  upper80 = as.numeric(gii_fc$upper[, 1]),
  lower95 = as.numeric(gii_fc$lower[, 2]),
  upper95 = as.numeric(gii_fc$upper[, 2])
)

p_gii <- ggplot() +
  geom_ribbon(
    data = gii_fc_df,
    aes(x = year, ymin = lower95, ymax = upper95, fill = "95% confidence interval"),
    alpha = 0.35
  ) +
  geom_ribbon(
    data = gii_fc_df,
    aes(x = year, ymin = lower80, ymax = upper80, fill = "80% confidence interval"),
    alpha = 0.5
  ) +
  geom_line(
    data = gii_fc_df,
    aes(x = year, y = mean, color = "No-policy forecasted GII"),
    linewidth = 1
  ) +
  geom_line(
    data = future_predictions,
    aes(x = year, y = Predicted_GII, color = "Net-zero policy impact on the GII"),
    linewidth = 1
  ) +
  scale_color_manual(
    name = "Scenarios",
    values = c(
      "No-policy forecasted GII" = "darkblue",
      "Net-zero policy impact on the GII" = "purple"
    )
  ) +
  scale_fill_manual(
    name = "Confidence intervals",
    values = c(
      "80% confidence interval" = "#6666FF",
      "95% confidence interval" = "#9999FF"
    )
  ) +
  labs(
    title = "Gender Inequality Index: \nTax Policy vs No-Policy Scenario",
    x = "Year",
    y = "GII"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )
```

```{r, echo=FALSE}
#| message: false
#| warning: false
p_warming
p_inequality
p_gii
```

The comparison between the net-zero target and no-policy scenarios highlights a clear trade-off. By contrast of the sharp increases seen before, without the required tax policy, warming inability remains around 16–17%, human inequality gradually declines, and gender inequality decreases over time. 

These projections suggest a potential trade-off between ambitious climate policy and social equity outcomes. Complementary redistributive or compensatory measures should be considered.

{{< pagebreak >}}

## 3. Is that policy viable ? Its impact on the GDP

To evaluate the potential economic impact of the proposed environmental tax policy, we forecast EU27 GDP under the tax scenario using the estimated coefficients from our GDP regression model. We construct the future path by applying the average annual increase in environmental tax revenues, combined with the underlying time trend, cumulatively on the last observed GDP level.

### a. The forecast model

```{r, echo=FALSE}
#| message: false
#| warning: false

# Average yearly tax increase
avg_tax_increment <- (last(tax_path$tax_required) -
                      first(tax_path$tax_required)) /
                      (nrow(tax_path) - 1)

# Last observed levels
last_obs <- tibble(GDP = tail(EU27_GDP$GDP_value, 1))

future_years <- length(tax_path$year)

# Extract coefficients
b_gdp_tax  <- coef(model_gdp_tax)["Envt_Tax_Revenues_value"]
b_gdp_time <- coef(model_gdp_tax)["TIME_PERIOD"]

# Build paths
future_gdp_predictions <- tibble(
  year = tax_path$year,
  Predicted_GDP =
    last_obs$GDP +
    cumsum(rep(b_gdp_tax * avg_tax_increment + b_gdp_time, future_years)))

ggplot(future_gdp_predictions,
                aes(x = year, y = Predicted_GDP)) +
  geom_line(color = "steelblue", linewidth = 1) +
  labs(
    title = "Forecasted impact of the tax policy on EU27 GDP",
    x = "Year",
    y = "GDP"
  ) +
  theme_minimal()+
  scale_y_continuous(limits = c(1.9434e+07, 1.9435e+07))
```

Under the tax policy scenario, projected GDP exhibits an almost flat trajectory over the period 2024–2050, increasing from approximately 19,434,512 million USD to 19,434,558 million USD. This corresponds to an average annual growth rate close to zero, indicating a near-complete stagnation of economic activity over the projection horizon. This result suggests that the required tax trajectory would substantially constrain aggregate economic growth. While these projections should not be interpreted as causal estimates, they point to a potentially significant macroeconomic cost associated to the the tax policy.

### b. Comparing with the no-policy forecast

We then project this GDP path alongside the ARIMA-based GDP forecast in the absence of any tax policy, allowing us to visually and quantitatively compare how the tax might influence aggregate econ

```{r, echo=FALSE}
#| message: false
#| warning: false
gdp_fc_df <- data.frame(
  year = as.numeric(time(gdp_fc$mean)),
  mean = as.numeric(gdp_fc$mean),
  lower80 = as.numeric(gdp_fc$lower[, 1]),
  upper80 = as.numeric(gdp_fc$upper[, 1]),
  lower95 = as.numeric(gdp_fc$lower[, 2]),
  upper95 = as.numeric(gdp_fc$upper[, 2])
)

ggplot() +
  geom_ribbon(
    data = gdp_fc_df,
    aes(x = year, ymin = lower95, ymax = upper95, fill = "95% confidence interval"),
    alpha = 0.35
  ) +
  geom_ribbon(
    data = gdp_fc_df,
    aes(x = year, ymin = lower80, ymax = upper80, fill = "80% confidence interval"),
    alpha = 0.5
  ) +
  geom_line(
    data = gdp_fc_df,
    aes(x = year, y = mean, color = "ARIMA forecast (no tax policy)"),
    linewidth = 1
  ) +
  geom_line(
    data = future_gdp_predictions,
    aes(x = year, y = Predicted_GDP, color = "Forecasted impact on Aggregate GDP"),
    linewidth = 1
  ) +
  scale_color_manual(
    name = "Scenarios",
    values = c(
      "ARIMA forecast (no tax policy)" = "darkblue",
      "Forecasted impact on Aggregate GDP" = "red"
    )
  ) +
  scale_fill_manual(
    name = "Confidence intervals",
    values = c(
      "80% confidence interval" = "#6666FF",
      "95% confidence interval" = "#9999FF"
    )
  ) +
  labs(
    title = "Aggregate GDP: Tax Policy vs No-Policy Scenario",
    x = "Year",
    y = "Aggregate GDP"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )
```

The comparison shows that the GDP forecast without the tax policy is consistently higher than the GDP projected under the tax scenario, with the no-policy path rising to 29.5 trillion USD in 2050, while the tax policy scenario remains nearly flat around 19.4 trillion USD. This indicates that, according to our model, implementing the environmental tax would largely slow the growth of aggregate GDP compared with the baseline. However, the tax-induced impact is relatively small in absolute terms early on, suggesting limited short-term economic disruption.

### c. Ratio Tax/GDP

To further assess the economic viability of the environmental tax policy, we examine the evolution of the tax burden relative to aggregate economic output by computing the environmental tax as a share of GDP under three distinct scenarios. 

First, we calculate the required-tax-to-GDP ratio, obtained by dividing the required tax trajectory to reach the net-zero objective by the baseline GDP forecast. 
Secondly, we compute a predicted-tax-to-GDP ratio, based on the forecasted environmental tax path in the absence of the policy, again expressed relative to forecasted GDP. 
Finally, we construct a tax-to-GDP ratio using GDP projected under the tax policy itself, which accounts for the potential macroeconomic feedback of the tax on economic activity. 

```{r, echo=FALSE}
#| message: false
#| warning: false
# Build GDP forecast dataframe
gdp_forecast_df <- tibble(
  year = seq(max(EU27_GDP$TIME_PERIOD),
             max(EU27_GDP$TIME_PERIOD) + 27),
  GDP_forecast = as.numeric(gdp_fc$mean)
)

# Required tax path → tax / GDP
required_data <- tax_path |>
  left_join(gdp_forecast_df, by = "year") |>
  mutate(
    tax_gdp_ratio = (tax_required / GDP_forecast) * 100,
    scenario = "Required tax-GDP ratio"
  ) |>
  select(year, tax_gdp_ratio, scenario)

# Forecasted tax path → tax / GDP
forecasted_data <- as.data.frame(tax_pred_95) |>
  mutate(year = tax_path$year) |>
  select(year, fit) |>
  left_join(gdp_forecast_df, by = "year") |>
  mutate(
    tax_gdp_ratio = (fit / GDP_forecast) * 100,
    scenario = "Forecasted tax-GDP ratio"
  ) |>
  select(year, tax_gdp_ratio, scenario)

# Tax/GDP using GDP under the tax scenario
future_gdp_df <- future_gdp_predictions |>
  rename(GDP_with_tax = Predicted_GDP) |>
  select(year, GDP_with_tax)

tax_gdp_with_taxGDP <- tax_path |>
  left_join(future_gdp_df, by = "year") |>
  mutate(
    tax_gdp_ratio = (tax_required / GDP_with_tax) * 100,
    scenario = "Tax/GDP using GDP with tax"
  ) |>
  select(year, tax_gdp_ratio, scenario)

# Combine both paths
comparison_data <- bind_rows(required_data, forecasted_data, tax_gdp_with_taxGDP)

#Viability zones
zones <- tibble(
  ymin = c(0, 3, 5),
  ymax = c(3, 5, Inf),
  zone = c("Viable", "Risk", "High Risk")
)

# Plot: Required vs Forecasted Tax as % of GDP
ggplot() +
  geom_rect(data = zones,
            aes(xmin = -Inf, xmax = Inf, ymin = ymin, ymax = ymax, fill = zone),
            alpha = 0.2) +
  geom_line(data = comparison_data,
            aes(x = year, y = tax_gdp_ratio, color = scenario),
            linewidth = 1.2) +
  geom_point(data = comparison_data,
             aes(x = year, y = tax_gdp_ratio, color = scenario),
             size = 2, alpha = 0.7) +
  geom_hline(yintercept = c(3, 5),
             linetype = "dashed", color = "gray40") +
  scale_color_manual(
    values = c(
      "Required tax-GDP ratio"   = "darkblue",
      "Forecasted tax-GDP ratio" = "firebrick",
      "Tax/GDP using GDP with tax" = "purple"
    ),
    name = "Scenario"
  ) +
  scale_fill_manual(
    values = c(
      "Viable" = "green",
      "Risk" = "gold",
      "High Risk" = "red"
    ),
    name = "Viability zone"
  ) +

  labs(
    title = "Economic Viability of Environmental Tax Policy",
    subtitle = "Tax as % of GDP under three scenarios",
    x = "Year",
    y = "Tax as % of GDP"
  ) +

  scale_y_continuous(breaks = seq(0, 10, 1)) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "right"
  )
```
These three trajectories are plotted against viability thresholds at 3% and 5% of GDP, defining zones of relative economic feasibility or risk. 

The comparison of tax-to-GDP ratios across the three scenarios shows the forecasted-tax ratio remains in the viable zone, while the required and the ratio with the GDP-under-policy increase over time respectively in the Risk and High risk zones. 

This framework allows us to assess how the interaction between taxation and economic output may affect the policy’s sustainability, including risks of growth disruption, reduced political acceptability, or broader macroeconomic instability.

{{< pagebreak >}}

# Conclusion

In conclusion, this research project has evaluated the effectiveness of environmental taxation as a tool for achieving the European Union’s goal of net-zero emissions by 2050. The study involved cleaning and merging large datasets covering the EU-27 from 1990 to 2024. By integrating indicators such as environmental tax revenues, the Gender Inequality Index (GII), the Coefficient of Human Inequality, and energy poverty, we assessed both environmental efficacy and socioeconomic sustainability.

Our findings indicate that while environmental taxes have a statistically significant negative association with GHG emissions, the impact has been small and we revealed that the current forecasted tax trajectory is insufficient to reach the European Green Deal’s objectives without important negative socioeconomic consequences. 

To determine the tax level required for the 2050 target, we inverted a level-level linear regression of total GHG emissions against tax revenues. This model allowed us to find a "required tax path" that would theoretically drive emissions to zero by itself. We found that environmental tax revenues would need to scale up to roughly 1.3 trillion USD by 2050 ($≈$684 billion USD more than predicted historical trends).

The impacts of such a massive tax increase on society and the economy would be severe:

• Social Inequality: The regressions show that environmental taxes are positively correlated with energy poverty, income inequality, and gender inequality. Implementing the required tax path would likely exacerbate these social costs, and increase these indicators largely above the levels expected in a no-policy scenario.
• Economic Viability: Although tax revenues tend to increase alongside GDP, the required fiscal burden would likely reach above 3% of the total GDP. This exceeds empirical thresholds for economic stability and could be highly disruptive.

If relying exclusively on environmental taxation proves to be socially regressive or economically unviable, what alternative strategies could the European Union use to fill the gap? Some possibilities could be the use of complementary green technology subsidies, the implementation of stricter sector-specific regulations, or the potential for economic sobriety and degrowth to reduce the overall energy demand.

# Limitations

This study has several limitations that should be considered when interpreting the results. First, the analysis focuses on environmental taxation as if it were the sole policy instrument, abstracting from other climate policies implemented at the EU level. Future research could address this by jointly modelling environmental taxes alongside complementary instruments such as subsidies, regulations, and redistributive measures within an integrated policy or general equilibrium framework.

Second, data availability limits the inequality analysis, as the Coefficient of Human Inequality is only available for the period 2010–2023. As longer time series become available, future studies could extend the estimation horizon or combine aggregate indicators with micro-level data to strengthen the robustness of inequality projections.

Third, although detailed information by household type and income group exists for energy poverty, this heterogeneity is not exploited, and the EU-level aggregation may mask substantial cross-country differences. Future work could rely on household-level microsimulation or country-specific analyses to better capture distributional effects. 


Finally, sector-level impacts are not analysed due to inconsistencies in sector classifications, a limitation that could be overcome by using harmonized input–output data or computable general equilibrium models to study sectoral adjustment mechanisms.