---
title: "dm-project"
author: "Daphné Perry and Justine Catta"
format: html
---

# Introduction


# Presenting our data

### Our data
1. 
Environmental tax revenues (in million of USD): energy, transport, pollution, resources
source : OECD
time period : 1994-2023

2. 
GDP and Value Added per capita annual growth per sector (in million of USD)
source : OECD
time period : 1990-2024

3.
Greenhouse gas emissions by source sector (in million of ton)
source : Eurostat
time period : 1990-2024

4.
Inability to keep home adequately warm
source : Eurostat
time period : 2003-2024

5.
HDR index
source : HDR
time period : 1990-2023

#### Justification for the Sector Mapping Between ISIC and CRF Classifications

The reconciliation of sectors across the ISIC-based dataset and the CRF (Common Reporting Format) classification relies on aligning economic activities with their dominant emission sources. Although the two systems are not structurally identical—ISIC being an economic activity classification and CRF being an emissions reporting framework—several robust correspondences can be established.

First, Agriculture, forestry and fishing (ISIC A) is mapped to CRF 3 (Agriculture). Both refer to primary-sector biological production and include emissions directly linked to agricultural practices such as enteric fermentation, manure management, and agricultural soils.

Second, Manufacturing (ISIC C) is associated with CRF 1.A.2 (Manufacturing industries and construction) and CRF 2 (Industrial processes and product use). Manufacturing activity generates emissions both from fuel combustion in industrial facilities (covered by 1.A.2) and from process-related sources such as cement, chemicals, or metals production (covered by CRF 2), making this linkage consistent with the structure of industrial emissions.

Third, Electricity, gas, steam and air conditioning supply (ISIC D)—even though aggregated within “Industry (except construction)” in the first dataset—logically corresponds to CRF 1.A.1 (Public electricity and heat production). This CRF category captures emissions from power generation and heat production, which are the core activities of ISIC D.

Finally, Transportation and storage (ISIC H)—included in a broader aggregated category in the dataset—is aligned with CRF 1.A.3 (Transport). This CRF category encompasses fuel combustion from road, rail, aviation, and navigation transport, which directly reflects the operational scope of ISIC H.

Overall, the mapping follows a principle of dominant emission source equivalence, ensuring that each ISIC activity is linked to the CRF category representing the primary origin of its greenhouse gas emissions. This approach allows the two datasets—despite structural differences—to be compared and integrated in a coherent and meaningful way for our analysis.


### Setup
```{r}
library(vroom)
library(ggplot2)
library(knitr)
library(dplyr)
library(here)
library(countrycode)
library(plotly)
library(tidyr)
```

```{r}
here::i_am("project.Rproj")
```


### Data loading
```{r}
#| message: false
Envt_Tax_Revenues <- vroom(here("data", "OECD_envt_tax.csv"))
GDP_and_AV <- vroom(here("data", "OECD_GDP_and_AV.csv"))
GHG_Emissions <- vroom(here("data", "Eurostat_ghg.csv"))
Warm_Households <- vroom(here("data","Eurostat_warm.csv"))
Soc_Indexes <- vroom(here("data", "HDR_indexes.csv"))
```

### Country keys

List of countries :
Austria, Belgium, Bulgaria, Croatia, Cyprus, Czechia, Denmark, Estonia, Finland, France, Germany, Greece, Hungary, Ireland, Italy, Latvia, Lithuania, Luxembourg, Malta, Netherlands, Poland, Portugal, Romania, Slovakia, Slovenia, Spain, Sweden

```{r}
eu_countries <- unique(GHG_Emissions$geo)

eu_country_key <- data.frame(name_en = eu_countries) |>
  mutate(
    iso3 = countrycode(name_en, origin = "country.name", destination = "iso3c"),
    iso2 = countrycode(name_en, origin = "country.name", destination = "iso2c"),
    name_fr = countrycode(name_en, origin = "country.name", destination = "country.name.fr")
  ) 
```


### Modifying the datasets

#### Removing unwanted variables
```{r}
Warm_Households <- Warm_Households |>
  select(geo, TIME_PERIOD, hhtyp, incgrp, OBS_VALUE) |>
  filter(geo %in% eu_country_key$iso2, 
         hhtyp %in% c("A1F", "A1M", "TOTAL"))

Envt_Tax_Revenues <- Envt_Tax_Revenues |>
  select(REF_AREA, 
         TIME_PERIOD, OBS_VALUE, CATEGORY,
         `Catégorie de recettes provenant des taxes liées à l'environnement`) |>
  filter(REF_AREA %in% eu_country_key$iso3)

GDP <- GDP_and_AV |>
  filter(Transaction == "Produit intérieur brut", 
         REF_AREA %in% eu_country_key$iso3) |>
  select(REF_AREA, TIME_PERIOD, OBS_VALUE)

Added_value <- GDP_and_AV |>
  filter(Transaction == "Valeur ajoutée, brute", 
         REF_AREA %in% eu_country_key$iso3, 
         `Activité économique` %in% c(
           "Agriculture, sylviculture et pêche",
           "Activités de fabrication",
           "Industrie (sauf construction)",
           "Commerce de gros et de détail; réparation de véhicules automobiles et de motocyclettes; transport et stockage; activités d'hébergement et de restauration"
           )) |>
  select(REF_AREA, `Activité économique`, 
         TIME_PERIOD, OBS_VALUE)

GHG_Emissions <- GHG_Emissions |>
  select(airpol, src_crf, geo, TIME_PERIOD, OBS_VALUE) |>
  filter(geo %in% eu_country_key$name_en, 
         src_crf != "Total (excluding memo items)")

Soc_Indexes <- Soc_Indexes |>
  pivot_longer(
    cols = -c(iso3, country, hdicode, region),
    names_to = c("Indicator", "TIME_PERIOD"),
    names_pattern = "^(.*)_(\\d{4})$",
    values_to = "OBS_VALUE"
  ) |>
  mutate(
    OBS_VALUE = as.numeric(OBS_VALUE),
    TIME_PERIOD = as.integer(TIME_PERIOD)
  )

GII <- Soc_Indexes|>
  filter(iso3 %in% eu_country_key$iso3,
         Indicator == "gii") |>
  select(iso3, TIME_PERIOD, OBS_VALUE)
  
coef_ineq <- Soc_Indexes|>
  filter(iso3 %in% eu_country_key$iso3,
         Indicator == "coef_ineq")|>
  select(iso3, TIME_PERIOD, OBS_VALUE)
  
```

#### Adding some entities

##### Add European average (accross countries)
```{r}
#| message: false
EU27_GDP <- GDP |>
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(`Zone de référence` = "EU Average",
         REF_AREA = "EU27") |>
  relocate(`Zone de référence`)
GDP <- bind_rows(GDP, EU27_GDP)

EU27_Tax <- Envt_Tax_Revenues |>
  group_by(TIME_PERIOD,
           `Catégorie de recettes provenant des taxes liées à l'environnement`) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(`Fréquence d'observation...8` = "EU Average", 
         REF_AREA = "EU27") |>
  relocate(`Fréquence d'observation...8`)
Envt_Tax_Revenues <- bind_rows(Envt_Tax_Revenues, EU27_Tax)

EU27_Warming <- Warm_Households |>
  group_by(TIME_PERIOD, hhtyp, incgrp) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(geo = "EU27") |>
  relocate(geo)
Warm_Households <- bind_rows(Warm_Households, EU27_Warming)

EU27_GII <- GII |>
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(country = "EU Average",
         iso3 = "EU27") |>
  relocate(country)
GII <- bind_rows(GII, EU27_GII)

EU27_Coefineq <- coef_ineq |>
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(country = "EU Average",
         iso3 = "EU27") |>
  relocate(country)
coef_ineq <- bind_rows(coef_ineq, EU27_Coefineq)
```


##### Add country average (accross sectors)
```{r}
#| message: false
EU27_AV <- Added_value |>
  group_by(TIME_PERIOD, `Activité économique`) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(`Zone de référence` = "EU average", 
         REF_AREA = "EU27") |>
  relocate(`Zone de référence`)
Added_value <- bind_rows(Added_value, EU27_AV)

EU27_GHG <- GHG_Emissions |>
  group_by(TIME_PERIOD, src_crf) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(geo = "EU average") |>
  relocate(geo)
GHG_Emissions <- bind_rows(GHG_Emissions, EU27_GHG)
```

#### Modifiying the characteristics of the variables
```{r}
GHG_Emissions$src_crf <- factor(GHG_Emissions$src_crf)
Added_value$`Activité économique` <- factor(Added_value$`Activité économique`)
```

```{r}
names(Warm_Households)[names(Warm_Households) == "OBS_VALUE"] <- "Warm_Households_value"
names(Envt_Tax_Revenues)[names(Envt_Tax_Revenues) == "OBS_VALUE"] <- "Envt_Tax_Revenues_value"
names(GDP)[names(GDP) == "OBS_VALUE"] <- "GDP_value"
names(Added_value)[names(Added_value) == "OBS_VALUE"] <- "AV_value"
names(GHG_Emissions)[names(GHG_Emissions) == "OBS_VALUE"] <- "GHG_Emissions_value"
names(GII)[names(GII) == "OBS_VALUE"] <- "GII_value"
names(coef_ineq)[names(coef_ineq) == "OBS_VALUE"] <- "coef_ineq_value"
#rename the OBS_VALUES
#rename the sectors
#sum the 2 sectors of manufacturing in GHG emissions
```

```{r}
Added_value <- Added_value |>
  mutate(`Activité économique` = recode(`Activité économique`,
                                        "Commerce de gros et de détail; réparation de véhicules automobiles et de motocyclettes; transport et stockage; activités d'hébergement et de restauration" = "Transport", 
                                        "Agriculture, sylviculture et pêche"= "Agriculture", 
                                        "Industrie (sauf construction)" = "Public electricity and heat production", 
                                        "Activités de fabrication" = "Manufacturing"
                                        ))

EU27_AV <- EU27_AV |>
  mutate(`Activité économique` = recode(`Activité économique`,
                                        "Commerce de gros et de détail; réparation de véhicules automobiles et de motocyclettes; transport et stockage; activités d'hébergement et de restauration" = "Transport", 
                                        "Agriculture, sylviculture et pêche"= "Agriculture", 
                                        "Industrie (sauf construction)" = "Public electricity and heat production", 
                                        "Activités de fabrication" = "Manufacturing"
                                        ))

GHG_Emissions <- GHG_Emissions|>
  mutate(
    src_crf = case_when(
      src_crf %in% c("Fuel combustion in manufacturing industries and construction", 
                     "Industrial processes and product use") ~ "Manufacturing",
      TRUE ~ src_crf
    )
  ) |>
  group_by(geo, TIME_PERIOD, src_crf) |>
  summarise(GHG_Emissions_value = sum(GHG_Emissions_value), .groups = "drop")

GHG_Emissions <- GHG_Emissions |>
  mutate(src_crf = recode(src_crf, 
                          "Fuel combustion in energy industries"="Public electricity and heat production", 
                          "Fuel combustion in transport"="Transport"))

EU27_GHG <- EU27_GHG|>
  mutate(
    src_crf = case_when(
      src_crf %in% c("Fuel combustion in manufacturing industries and construction", 
                     "Industrial processes and product use") ~ "Manufacturing",
      TRUE ~ src_crf
    )
  ) |>
  group_by(geo, TIME_PERIOD, src_crf) |>
  summarise(EU27_GHG = sum(OBS_VALUE), .groups = "drop")

EU27_GHG <- EU27_GHG |>
  mutate(src_crf = recode(src_crf, 
                          "Fuel combustion in energy industries"="Public electricity and heat production", 
                          "Fuel combustion in transport"="Transport"))
```


# Descriptive analyzes

### 1. Environmental tax revenues
```{r}
ggplot(EU27_Tax, aes(x = TIME_PERIOD, y = OBS_VALUE, color = `Catégorie de recettes provenant des taxes liées à l'environnement`, group = `Catégorie de recettes provenant des taxes liées à l'environnement`)) +
  geom_line(size = 1) +
  labs(
    title = "EU27 Environmental Tax Revenues by Category",
    x = "Year",
    y = "Revenue (million USD)",
    color = "Category"
  ) +
  theme_minimal()
```
COMMENT 

### 2. GDP per country
```{r}
p <- ggplot(GDP, aes(x = TIME_PERIOD, y = GDP_value, color = REF_AREA, group = REF_AREA)) +
  geom_line(size = 1) +
  labs(
    title = "GDP per Country (EU27)",
    x = "Year",
    y = "GDP (million USD)",
    color = "Country"
  ) +
  theme_minimal()

ggplotly(p)
```
COMMENT 



### 3. Added Value
```{r}
p_va <- ggplot(EU27_AV, aes(x = TIME_PERIOD, y = OBS_VALUE, color = `Activité économique`, group = `Activité économique`)) +
  geom_line(size = 1) +
  labs(
    title = "EU27 Average Added Value per Sector",
    x = "Year",
    y = "Value Added (million USD)",
    color = "Sector"
  ) +
  theme_minimal()

ggplotly(p_va)
```
COMMENT 


### 4. Households

```{r}

ggplot(EU27_Warming, 
       aes(x = TIME_PERIOD, y = OBS_VALUE,
           color = interaction(hhtyp, incgrp),
           group = interaction(hhtyp, incgrp))) +
  geom_line(size = 1) +
  labs(
    title = "EU27 – inability to Keep the House Warm (by Sex × Income Group)",
    x = "Year",
    y = "Percentage of population",
    color = "Sex × Income"
  ) +
  theme_minimal()
```

COMMENT 





# Correlations & regressions

## Between taxes and GHG 
```{r}
GHG_Emissions <- GHG_Emissions |> 
  left_join(eu_country_key, by = c("geo" = "name_en"))
```

```{r}
# keeping sectors
emitting_sectors <- c(
  "Agriculture",
  "Fuel combustion in manufacturing industries and construction",
  "Industrial processes and product use",
  "Fuel combustion in transport",
  "Fuel combustion in energy industries"
)

# filter GHG_Emissions
GHG_Emissions_filtered <- GHG_Emissions |>
  filter(src_crf %in% emitting_sectors)

# join with taxes
tax_ghg <- Envt_Tax_Revenues |>
  left_join(
    GHG_Emissions_filtered,
    by = c("REF_AREA" = "iso3", "TIME_PERIOD" = "TIME_PERIOD")
  ) |>
  rename(
    tax_value = OBS_VALUE.x,
    ghg_value = OBS_VALUE.y
  ) |>
  filter(!is.na(ghg_value)) |>
  select(REF_AREA, TIME_PERIOD, CATEGORY, tax_value, src_crf, ghg_value)


```

```{r}
tax_ghg$src_crf <- as.factor(tax_ghg$src_crf)

# GDP as control
tax_ghg_control <- tax_ghg %>%
  left_join(GDP %>% select(REF_AREA, TIME_PERIOD, OBS_VALUE) %>% rename(gdp = OBS_VALUE),
            by = c("REF_AREA", "TIME_PERIOD"))
```


```{r}
model <- lm(ghg_value ~ tax_value + gdp + src_crf + TIME_PERIOD, data = tax_ghg_control)
summary(model)
```
We estimated a linear regression to examine the relationship between environmental taxes and greenhouse gas emissions, while controlling for GDP, GHG sector, and year. The model includes 19,625 observations after excluding 625 rows with missing data. The regression explains approximately 57% of the variation in emissions, as indicated by the R-squared value of 0.571.

The coefficient on environmental taxes is -0.019, which is not statistically significant (p = 0.377). This suggests that, after accounting for GDP, sector, and year, there is no clear effect of environmental taxes on emissions in this dataset.

GDP has a positive and highly significant effect on emissions, with a coefficient of 0.0385 (p < 0.001), indicating that countries with higher GDP tend to have higher emissions.

The GHG sector also strongly affects emissions. Compared to the baseline sector, the estimated differences are as follows: fuel combustion in energy industries adds about 29,820 units, fuel combustion in manufacturing industries and construction adds about 3,944 units, fuel combustion in transport adds about 15,340 units, and industrial processes and product use reduces emissions by 615 units (the latter is not statistically significant).

Finally, the coefficient on the year variable is -827, which is highly significant, indicating a general downward trend in emissions over time.

Overall, the results suggest that sectoral composition and GDP are the main drivers of emissions, while environmental taxes, in this specification, do not show a statistically detectable effect. The negative time trend suggests that emissions have generally declined over the period considered, independent of tax levels.



# Forecasting
### tax needed 

```

### impact of tax


