---
title: "dm project - Phase 3"
author: "Daphne Perry and Justine Catta"
format: pdf
---

# Introduction

## Description of our research question

Our research question is the following: “To what extent do environmental taxes reduce Greenhouse Gas Emissions (GHG) in EU countries, and what are their long-term socioeconomic impacts across economy?”. As the European Union commits to achieving net-zero emissions by 2050 under the European Green Deal, we aim to better understand the effectiveness and consequences of environmental taxation, both on emissions and socially.

To answer this question, we will use panel data for all EU member states. Our data set includes GHG emissions, GDP, environmental tax revenues, and several indicators of social well-being. These include the Gender Inequality Index (GII), a coefficient of human inequality, and a measure of energy poverty that is the inability to adequately heat one’s home. Together, these variables allow us to evaluate the distributive consequences of taxes.

After cleaning, harmonizing, and describing the data sets, we aim to estimate several relationships. First, we will measure the link between environmental taxation and GHG emissions. Do higher taxes lead to statistically significant reductions in emissions at both the EU and country levels? **DO WE KEEP THAT QUESTION ? BECAUSE NO REF TO COUNTRY LVL AFTER**

We will then investigate the association between environmental taxes and social inequality indicators. Our core question here can be summarized as: “Is there a social cost associated with environmental taxation, and if so, how large is it?” Specifically, we will analyse whether environmental taxes correlate with increases or decreases in gender inequality, income inequality, and energy poverty. This step is essential because policies that are environmentally effective but socially regressive may have a reduced long-term viability.

Building on these empirical relationships, we will construct forecasts for the period 2026–2050 to estimate the level of taxation required to meet the European Green Deal’s emissions targets. We consider a policy scenario relying solely on environmental taxation applied to relevant sectors. Our analysis will assess whether such an approach is socially and economically sustainable: How would different tax levels affect inequalities and energy poverty? How much revenue would such taxes generate, and would the required tax burden be compatible with each country’s GDP levels?

By combining environmental effectiveness, economic feasibility, and social impact, we can then address our overarching question: Is it socially and economically optimal for the EU to rely exclusively on environmental taxation to achieve net-zero emissions by 2050? If taxation alone proves insufficient or too socially costly, our results may highlight the need for complementary policies such as subsidies for green technologies or targeted social transfers to ensure a fair and effective ecological transition.

## Link to the github project

Our project is available on github [here](https://github.com/cattajustine-spec/project).

## Setup

```{r echo=FALSE}
#| message: false
#| warning: false
library(vroom)
library(ggplot2)
library(knitr)
library(dplyr)
library(here)
library(countrycode)
library(plotly)
library(tidyr)
library(knitr)
library(plm)
library(tidyverse)
library(broom)
library(forecast)
library(purrr)
library(gridExtra)
```

```{r, echo=FALSE}
#| message: false
#| warning: false
here::i_am("project.Rproj")
```

## Presenting our data:
### 1. Environmental tax revenues (OECD, 1994–2023)

This data set reports annual environmentally related tax revenues for EU countries from the OECD ([link](https://data-explorer.oecd.org/vis?lc=en&tm=DF_ERTR&pg=0&snb=1&vw=tb&df%5Bds%5D=dsDisseminateFinalDMZ&df%5Bid%5D=DSD_ERTR%40DF_ERTR&df%5Bag%5D=OECD.ENV.EPI&df%5Bvs%5D=&pd=1994%2C2023&dq=A.HRV%2BLVA%2BLTU%2BLUX%2BMLT%2BNLD%2BCYP%2BAUT%2BBEL%2BCZE%2BDNK%2BEST%2BFIN%2BFRA%2BDEU%2BGRC%2BHUN%2BIRL%2BITA%2BPOL%2BPRT%2BSVK%2BSVN%2BESP%2BSWE%2BBGR%2BROU.TAXREV.CAT_ENE%2BCAT_TRA%2BCAT_POL%2BCAT_RES%2B_T._T.USD.&ly%5Brw%5D=REF_AREA&ly%5Bcl%5D=TIME_PERIOD&to%5BTIME_PERIOD%5D=false)). Values are expressed in millions of USD and they cover a time period from 1994 to 2023.

```{r, echo=FALSE}
#| message: false
#| warning: false
Envt_Tax_Revenues <- vroom(here("data", "OECD_envt_tax.csv"))
```

::: {columns}
|                   |                             |
|-------------------|-----------------------------|
| Name              | `r "Envt_Tax_Revenues"`     |
| Number of rows    | `r nrow(Envt_Tax_Revenues)` |
| Number of columns | `r ncol(Envt_Tax_Revenues)` |
:::

### 2. GDP (OECD, 1990-2024)

This data set provides annual GDP figures for EU countries from the OECD ([link](https://data-explorer.oecd.org/vis?df%5Bds%5D=dsDisseminateFinalDMZ&df%5Bid%5D=DSD_NAMAIN10%40DF_TABLE1_OUTPUT&df%5Bag%5D=OECD.SDD.NAD&df%5Bvs%5D=&dq=A.BEL%2BAUT%2BCZE%2BDNK%2BEST%2BFIN%2BFRA%2BDEU%2BGRC%2BHUN%2BIRL%2BITA%2BLVA%2BLTU%2BLUX%2BNLD%2BPOL%2BPRT%2BSVK%2BSVN%2BESP%2BSWE%2BBGR%2BHRV%2BCYP%2BMLT%2BROU...B1G%2BB1GQ.._Z%2B_T%2BC%2BA%2BBTE%2BF%2BGTI%2BJ%2BK%2BL%2BM_N%2BOTQ%2BRTU..USD_EXC.V..&to%5BTIME_PERIOD%5D=false&pd=1990%2C2024&vw=tb)), measured in millions of USD, using exchange rate and they cover the time period 1990 to 2024.

```{r, echo=FALSE}
#| message: false
#| warning: false
GDP_and_AV <- vroom(here("data", "OECD_GDP_and_AV.csv"))
```

::: {columns}
|                   |                      |
|-------------------|----------------------|
| Name              | `r "GDP_and_AV"`     |
| Number of rows    | `r nrow(GDP_and_AV)` |
| Number of columns | `r ncol(GDP_and_AV)` |
:::

### 3. Greenhouse gas emissions (Eurostat, 1990-2023)

This data set reports total greenhouse gas emissions for the 27 EU countries. Values are expressed in thousand tonnes of CO₂-equivalent. The data set is from Eurostat ([here](https://ec.europa.eu/eurostat/databrowser/view/env_air_gge__custom_18942108/default/table)) and covers the time period 1990 to 2023.

```{r, echo=FALSE}
#| message: false
#| warning: false
GHG_Emissions <- vroom(here("data", "Eurostat_ghg.csv"))
```

::: {columns}
|                   |                         |
|-------------------|-------------------------|
| Name              | `r "GHG_Emissions"`     |
| Number of rows    | `r nrow(GHG_Emissions)` |
| Number of columns | `r ncol(GHG_Emissions)` |
:::

### 4. Inability to keep home adequately warm (Eurostat, 2003-2024)

This data set measures the percentage of households unable to keep their home adequately warm. We chose 3 households type: total, single female and single male and 2 income situations: total and below 60% of the median (in situation of poverty). The data set is from Eurostat ([here](https://ec.europa.eu/eurostat/databrowser/view/ILC_MDES01__custom_18945942/default/table)) and covers 2003-2024. Note that it is a declarative observation of household who report that they are unable to keep their home adequately warm.

```{r, echo=FALSE}
#| message: false
#| warning: false
Warm_Households <- vroom(here("data","Eurostat_warm.csv"))
```

::: {columns}
|                   |                           |
|-------------------|---------------------------|
| Name              | `r "Warm_Households"`     |
| Number of rows    | `r nrow(Warm_Households)` |
| Number of columns | `r ncol(Warm_Households)` |
:::

### 5. Human index (UNDP HDR, 1990-2023)

We focus on the Coefficient of Human Inequality and Gender Inequality Index (GII) for most countries worldwide, from this HDR data set ([link](https://hdr.undp.org/sites/default/files/2025_HDR/HDR25_Composite_indices_complete_time_series.csv)) providing annual values for lots of Indexes, from 1990 to 2023.

```{r, echo=FALSE}
#| message: false
#| warning: false
Soc_Indexes <- vroom(here("data", "HDR_indexes.csv"))
```

::: {columns}
|                   |                       |
|-------------------|-----------------------|
| Name              | `r "Soc_Indexes"`     |
| Number of rows    | `r nrow(Soc_Indexes)` |
| Number of columns | `r ncol(Soc_Indexes)` |
:::

# Modifying the data sets

Because of the size of the data sets we chose, we remove all the variables we will not use, then add missing aggregates that we think are important for our analysis, and then modify the variables we kept.

Finally, to simplify our analysis, we will merge all our data sets in one data base.

## Country keys {#countrykeys}

Our analysis is on the 27 countries in the European Union. We construct a key that filters the EU-27 countries and harmonize the different country identifiers: ISO2 codes, ISO3 codes, or full English and French names, to ensure consistency across datasets and avoid mismatches during merging.

```{r, echo=FALSE}
#| message: false
#| warning: false
eu_countries <- unique(GHG_Emissions$geo)

eu_country_key <- data.frame(name_en = eu_countries) |>
  mutate(
    iso3 = countrycode(name_en, origin = "country.name", destination = "iso3c"),
    iso2 = countrycode(name_en, origin = "country.name", destination = "iso2c"),
    name_fr = countrycode(name_en, origin = "country.name", destination = "country.name.fr"))  |>
  bind_rows(tibble(name_en = "EU27", name_fr = "EU27", iso2 = "EU27", iso3 = "EU27"))
```

## Removing unwanted variables

The original datasets contain many variables and dimensions that are irrelevant to our research question (such as measurement units, alternative breakdowns, metadata, or categories outside our scope). For each dataset, we systematically retain only four essential components:

-   The geographical identifier (ISO2, ISO3, or name).
-   The reference year.
-   The relevant economic or social variable (e.g., revenues, GDP, emissions, inequality index).

This ensures a clean, comparable structure across datasets before we perform more complex transformations.

### 1. Environmental tax revenues (in million of USD)

In this data set, we select only the ISO3 code, the reference year, and the observed numerical value of Environmental related tax revenues.

```{r, echo=FALSE}
#| message: false
#| warning: false
Envt_Tax_Revenues <- Envt_Tax_Revenues |>
  filter(REF_AREA %in% eu_country_key$iso3,
         CATEGORY == "_T") |>
  select(REF_AREA, TIME_PERIOD, OBS_VALUE)
```

### 2. GDP (in million of USD)

In this data set, we select only the ISO3 code, the reference year and the observed numerical value of GDP.

```{r, echo=FALSE}
#| message: false
#| warning: false
GDP <- GDP_and_AV |>
  filter(Transaction == "Produit intérieur brut", 
         REF_AREA %in% eu_country_key$iso3) |>
  select(REF_AREA, TIME_PERIOD, OBS_VALUE)
```

### 3. Greenhouse gas emissions (in million of ton)

We keep the English name, the reference year and the observed numerical value of Greenhouse gas emissions. 

```{r, echo=FALSE}
#| message: false
#| warning: false
GHG_Emissions <- GHG_Emissions |>
  filter(geo %in% eu_country_key$name_en, 
         src_crf == "Total (excluding memo items)")|>
    select(geo, TIME_PERIOD, OBS_VALUE)
```

### 4. Inability to keep home adequately warm

In this data set, we keep the ISO2 code, the reference year, the different household types labels as well as the income groups labels, and the observed numerical values (percentage). 
In precision, the household types are the households composed by only one female, the households composed by only one male,  all households (whatever their composition) and the income groups are the households who have an income below 60% of the median income (meaning they are below the poverty threshold) and all households (whatever their revenue).

```{r, echo=FALSE}
#| message: false
#| warning: false
Warm_Households <- Warm_Households |>
  select(geo, TIME_PERIOD, hhtyp, incgrp, OBS_VALUE) |>
  filter(geo %in% eu_country_key$iso2, 
         hhtyp %in% c("A1F", "A1M", "TOTAL"))
```

### 5. HDR index

In this data set, we have two main variables of interest: the [Gender Inequality Index](#genderinequalityindex) and the [Coefficient of Human Inequality](#coefficientofhumaninequality).

They are described per year as different variables (meaning we have a gii_2020, gii_2021, ..., coef_ineq_2003, coef_ineq_2004, etc.). Therefore, we need to pivot our database, and then create two independent data sets, one for each variable of interest.

```{r, echo=FALSE}
#| message: false
#| warning: false
Soc_Indexes <- Soc_Indexes |>
  pivot_longer(
    cols = matches("_[0-9]{4}$"),   
    names_to = c("Indicator", "TIME_PERIOD"),
    names_pattern = "^(.*)_(\\d{4})$",
    values_to = "OBS_VALUE"
  ) |>
  mutate(
    OBS_VALUE = as.numeric(OBS_VALUE),
    TIME_PERIOD = as.integer(TIME_PERIOD)
  )
```

#### Gender inequality Index {#genderinequalityindex}

We keep the ISO3 code, the reference year and the observed numerical value of the GII.

```{r, echo=FALSE}
#| message: false
#| warning: false
GII <- Soc_Indexes|>
  filter(iso3 %in% eu_country_key$iso3,
         Indicator == "gii") |>
  select(iso3, TIME_PERIOD, OBS_VALUE)
```

#### Coefficient of Human inequality {#coefficientofhumaninequality}

We keep the ISO3 code, the reference year and the observed numerical value of the Coefficient of Human inequality.

```{r, echo=FALSE}
#| message: false
#| warning: false
coef_ineq <- Soc_Indexes|>
  filter(iso3 %in% eu_country_key$iso3,
         Indicator == "coef_ineq")|>
  select(iso3, TIME_PERIOD, OBS_VALUE)
```

## Adding some entities

### Add European average (across countries)

For the data sets that describe inequality indexes, we compute the European average across countries.

```{r, echo=FALSE}
#| message: false
#| warning: false
EU27_GII <- GII |>  #computing the average European gender inequality index. 
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(iso3 = "EU27") |>
  relocate(iso3)
GII <- bind_rows(GII, EU27_GII)

EU27_Coefineq <- coef_ineq |> #computing the average European coefficient of human inequality. 
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(iso3 = "EU27") |>
  relocate(iso3)
coef_ineq <- bind_rows(coef_ineq, EU27_Coefineq)

EU27_Warming <- Warm_Households |>  #computing the average European inability to keep home adequately warm, depending on the household composition and income.  
  group_by(TIME_PERIOD, hhtyp, incgrp) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(geo = "EU27") |>
  relocate(geo)
#adding in Warm households, using country keys to have by iso3
Warm_Households <- Warm_Households |> 
  rename(iso2 = geo) |>
  left_join(eu_country_key |> select(iso2, iso3), by = "iso2")
EU27_Warming <- EU27_Warming |> 
  rename(iso2 = geo) |> 
  left_join(eu_country_key |> select(iso2, iso3), by = "iso2")
Warm_Households <- bind_rows(Warm_Households, EU27_Warming) 
Warm_Households <- Warm_Households |> 
  select (iso3, TIME_PERIOD, hhtyp, incgrp, OBS_VALUE)
```

### Add European total

For other economic data sets (GDP, Taxes, GHG emissions), we compute the European total across countries.

```{r, echo=FALSE}
#| message: false
#| warning: false
EU27_GDP <- GDP |>                #computing the total European GDP 
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) |>
  mutate(REF_AREA = "EU27") |>
  relocate(REF_AREA)
GDP <- bind_rows(GDP, EU27_GDP)

EU27_Tax <- Envt_Tax_Revenues |>   #computing the total European environmental tax revenue
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) |>
  mutate(REF_AREA = "EU27") |>
  relocate(REF_AREA)
Envt_Tax_Revenues <- bind_rows(Envt_Tax_Revenues, EU27_Tax)

EU27_GHG <- GHG_Emissions |> #computing the total European greenhouse gas emissions
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) |>
  mutate(geo = "EU27") |>
  relocate(geo)
#adding in GHG_Emissions, using country keys to have by iso3
GHG_Emissions <- GHG_Emissions |> 
  rename(name_en = geo) |>
  left_join(eu_country_key |> select(name_en, iso3), by = "name_en")
EU27_GHG <- EU27_GHG  |> 
  rename(name_en = geo) |> 
  left_join(eu_country_key |> select(name_en, iso3), by = "name_en")
GHG_Emissions <- bind_rows(GHG_Emissions, EU27_GHG) 
GHG_Emissions <- GHG_Emissions  |> select (iso3, TIME_PERIOD, OBS_VALUE)
```

## Modifiying the characteristics of the variables

### Changing the type of some variables

For the inability to keep home adequately warm variable, we consider the household type variable and the income group variable as *factor*, rather than as *character* variables.

```{r, echo=FALSE}
#| message: false
#| warning: false
Warm_Households$hhtyp <- factor(Warm_Households$hhtyp)
EU27_Warming$hhtyp <- factor(EU27_Warming$hhtyp)
Warm_Households$incgrp <- factor(Warm_Households$incgrp)
EU27_Warming$incgrp <- factor(EU27_Warming$incgrp)
```

### Renaming some variables {#renamingsomevariables}

To be able to merge our data sets in a data base, we need to modify the name of the variable "**OBS_VALUE**" contained in several data sets to be able to merge. These variables are renamed as "**dataset_value**".

```{r, echo=FALSE}
#| message: false
#| warning: false
names(Warm_Households)[names(Warm_Households) == "OBS_VALUE"] <- "Warm_Households_value"
names(EU27_Warming)[names(EU27_Warming) == "OBS_VALUE"] <- "Warm_Households_value"
names(Envt_Tax_Revenues)[names(Envt_Tax_Revenues) == "OBS_VALUE"] <- "Envt_Tax_Revenues_value"
names(EU27_Tax)[names(EU27_Tax) == "OBS_VALUE"] <- "Envt_Tax_Revenues_value"
names(GDP)[names(GDP) == "OBS_VALUE"] <- "GDP_value"
names(EU27_GDP)[names(EU27_GDP) == "OBS_VALUE"] <- "GDP_value"
names(GHG_Emissions)[names(GHG_Emissions) == "OBS_VALUE"] <- "GHG_Emissions_value"
names(EU27_GHG)[names(EU27_GHG) == "OBS_VALUE"] <- "GHG_Emissions_value"
names(GII)[names(GII) == "OBS_VALUE"] <- "GII_value"
names(EU27_GII)[names(EU27_GII) == "OBS_VALUE"] <- "GII_value"
names(coef_ineq)[names(coef_ineq) == "OBS_VALUE"] <- "coef_ineq_value"
names(EU27_Coefineq)[names(EU27_Coefineq) == "OBS_VALUE"] <- "coef_ineq_value"
```

We rename the country codes as the ones in the country key we created in [Country keys](#countrykeys).

```{r, echo=FALSE}
#| message: false
#| warning: false
Envt_Tax_Revenues <- Envt_Tax_Revenues |>
  rename(iso3 = REF_AREA)
EU27_Tax <- EU27_Tax |>
  rename(iso3 = REF_AREA)

GDP <- GDP |> 
  rename(iso3 = REF_AREA)
EU27_GDP <- EU27_GDP |> 
  rename(iso3 = REF_AREA)
```

# Merging all data sets on country-year pairs

We cleaned and harmonized all datasets, and can now merge them into a single database combining economic, environmental and social indicators. The environmental tax dataset is used as the base for the merge; we then enrich it other datasets through matching on iso3 code, reference year. 

All merges use left_join(). This choice ensures that countries or years missing some variables are retained in the dataset rather than dropped — which would happen with inner_join(). Keeping these rows is important for transparency and for allowing later analyses to handle missingness explicitly.

This merge will allow us to study environmental taxes, emissions, economic structure, and social inequalities jointly within a single analytical framework.

```{r, echo=FALSE}
#| message: false
#| warning: false
all_merged <- Envt_Tax_Revenues |>
  left_join(GDP, by = c("iso3", "TIME_PERIOD")) |>
  left_join(GHG_Emissions, by = c("iso3", "TIME_PERIOD")) |>
  left_join(Warm_Households, by = c("iso3", "TIME_PERIOD")) |>
  left_join(GII, by = c("iso3", "TIME_PERIOD")) |>
  left_join(coef_ineq, by = c("iso3", "TIME_PERIOD")) |>
  select(iso3, TIME_PERIOD, hhtyp, incgrp, Warm_Households_value, GII_value, coef_ineq_value, Envt_Tax_Revenues_value, GDP_value, GHG_Emissions_value) |>
  filter(iso3 != "EU27")
```

# Descriptive analyzes

Now that we have cleaned our datasets, we can move on to some descriptive analyses of our main target variables.

## 1. European environmental tax revenues

```{r, echo=FALSE}
#| message: false
#| warning: false
ggplot(EU27_Tax, aes(x = TIME_PERIOD, y = Envt_Tax_Revenues_value)) +
  geom_line(linewidth = 1, color = "darkgreen") +
  labs(
    title = "Total European environmental tax revenues",
    x = "Year",
    y = "Environmental tax revenues (millions of USD)"
  ) +
  theme_minimal()
```

```{r, echo=FALSE}
#| message: false
#| warning: false
summary_stats_taxrevenues <- EU27_Tax |>
  summarise(
    Min = min(Envt_Tax_Revenues_value, na.rm = TRUE),
    Max = max(Envt_Tax_Revenues_value, na.rm = TRUE),
    Mean = mean(Envt_Tax_Revenues_value, na.rm = TRUE),
    Median = median(Envt_Tax_Revenues_value, na.rm = TRUE),
    SD = sd(Envt_Tax_Revenues_value, na.rm = TRUE)
    )

kable(summary_stats_taxrevenues, 
      caption = "Descriptive statistics of the total European environmental tax revenues (all years combined)")
```

The total European environmental tax revenues amounted to roughly 376 billion USD in the late 1990s. Over the 2000s, these revenues grew steadily and nearly doubled, reaching over 700 billion USD by 2010. Since then, the trend has remained broadly upward, though with noticeable fluctuations from year to year.

Revenues peaked at approximately 809 billion USD, while the long-run average (≈ 622 billion USD) reflects sustained and historically high levels of environmental taxation in the European Union. The standard deviation (≈ 152 billion USD) signals substantial variation over time, consistent with cyclical movements, recent economic crisis, and policy adjustments across the Union.

## 2. European GDP

```{r, echo=FALSE}
#| message: false
#| warning: false
ggplot(EU27_GDP, aes(x = TIME_PERIOD, y = GDP_value)) +
  geom_line(linewidth = 1, color = "darkgreen") +
  labs(
    title = "European GDP (in millions of USD)",
    x = "Year",
    y = "GDP (million USD)"
  ) +
  theme_minimal()
```

```{r, echo=FALSE}
#| message: false
#| warning: false
summary_stats_GDP <- EU27_GDP |>
  summarise(
    Min = min(GDP_value, na.rm = TRUE),
    Max = max(GDP_value, na.rm = TRUE),
    Mean = mean(GDP_value, na.rm = TRUE),
    Median = median(GDP_value, na.rm = TRUE),
    SD = sd(GDP_value, na.rm = TRUE)
    )

kable(summary_stats_GDP, 
      caption = "Descriptive statistics of the European GDP (all years combined)")
```

European GDP has increased a lot since the early 1990s. Starting from roughly 6.3 trillion USD, GDP expanded strongly throughout the 2000s, with a brief slowdown around the early 2010s (due to the subprime crisis), before resuming its upward trajectory. By 2024, it reached nearly 19.4 trillion USD, its historical maximum.

The long-run average GDP (≈ 12.2 trillion USD) and the very large standard deviation (≈ 4.1 trillion USD) reflect substantial economic growth over time as well as cyclical fluctuations. The median (≈ 13.7 trillion USD) suggests that GDP has remained consistently high over the last two decades, despite temporary downturns.

## 3. Greenhouse gas emissions 

```{r, echo=FALSE}
#| message: false
#| warning: false
ggplot(EU27_GHG, aes(x = TIME_PERIOD, y = GHG_Emissions_value)) +
  geom_line(linewidth = 1, color = "darkgreen") +
  labs(
    title = "Total European Greenhouse gas emissions (in millions of USD)",
    x = "Year",
    y = "Greenhouse gas emissions (thousands of ton of CO2)"
  ) +
  theme_minimal()
```


```{r, echo=FALSE}
#| message: false
#| warning: false
summary_stats_GHG <- EU27_GHG |>
  summarise(
    Min = min(GHG_Emissions_value, na.rm = TRUE),
    Max = max(GHG_Emissions_value, na.rm = TRUE),
    Mean = mean(GHG_Emissions_value, na.rm = TRUE),
    Median = median(GHG_Emissions_value, na.rm = TRUE),
    SD = sd(GHG_Emissions_value, na.rm = TRUE)
    )

kable(summary_stats_GHG, caption = "Descriptive statistics of EU total greenhouse gases emissions (all years combined)")
```

The EU27 total data show the total greenhouse gas emissions, displaying a downward trend. Their maximal value, obtained in 1990 is around 4.6 million units *WHAT UNIT ? *, and the lower one in 2023 is around 2.9 million units.

To see the descriptive analysis per sector, please refer to the document for the phase 2.

## 4. Inability to keep home adequately warm per household type

```{r, echo=FALSE}
#| message: false
#| warning: false
ggplot(EU27_Warming, 
       aes(x = TIME_PERIOD, y = Warm_Households_value,
           color = interaction(hhtyp, incgrp),
           group = interaction(hhtyp, incgrp))) +
  geom_line(linewidth = 1) +
  labs(
    title = "EU27 – inability to Keep the House Warm (by Sex × Income Group)",
    x = "Year",
    y = "Percentage of population",
    color = "Sex × Income"
  ) +
  theme_minimal()
```

```{r, echo=FALSE}
#| message: false
#| warning: false
summary_stats_Warm <- EU27_Warming |>
  group_by(hhtyp, incgrp) |>                    
  summarise(
    Min = min(Warm_Households_value, na.rm = TRUE),
    Max = max(Warm_Households_value, na.rm = TRUE),
    Mean = mean(Warm_Households_value, na.rm = TRUE),
    Median = median(Warm_Households_value, na.rm = TRUE),
    SD = sd(Warm_Households_value, na.rm = TRUE), 
    .groups = "drop"
  )

kable(summary_stats_Warm, caption = "Descriptive statistics of the inability to keep house adequately warm, by household type (all years combined)")
```

For the EU27 average, six curves represent different subgroups defined by sex and income. The values correspond to the percentage of each subgroup unable to keep their home adequately warm. All groups follow roughly the same pattern: a sharp increase between 2003 and 2005, a gradual decline afterward, and two significant shocks: a drop in 2008–2009 and a spike in 2022–2023, which can be linked to rising energy prices following the Russia–EU energy crisis.

Among all subgroups, the total-population categories show the lowest percentages, remaining below 10% since 2015. Conversely, households with incomes below 60% of the national median (below the poverty threshold) consistently show the highest rates, around 20% on average.

Women in low-income single households consistently report slightly higher percentages than men. For instance, low-income women (A1F, B_MD60) have an average inability rate of ≈ 21.1%, compared with ≈ 23.1% for low-income men (A1M, B_MD60), while total-population rates remain much lower at ≈ 10% across categories.

## 5. GII per country

```{r, echo=FALSE}
#| message: false
#| warning: false
p_GII <- ggplot(GII, aes(x = TIME_PERIOD, y = GII_value, color = iso3, group = iso3)) +
  geom_line(linewidth = 1) +
  labs(
    title = "GII per Country (EU27)",
    x = "Year",
    y = "Gender inequality index",
    color = "Country"
  ) +
  scale_y_continuous(limits = c(0, 1)) +   
  theme_minimal()

ggplotly(p_GII)
```

```{r, echo=FALSE}
#| message: false
#| warning: false
summary_stats_GII <- GII |>
  group_by(iso3) |>                        
  summarise(
    Min = min(GII_value, na.rm = TRUE),
    Max = max(GII_value, na.rm = TRUE),
    Mean = mean(GII_value, na.rm = TRUE),
    Median = median(GII_value, na.rm = TRUE),
    SD = sd(GII_value, na.rm = TRUE)
    )

kable(summary_stats_GII, 
      caption = "Descriptive statistics of the GII by country (all years combined)")
```

The Gender Inequality Index (GII) shows a clear downward trend across all European countries between 1990 and 2024. At the beginning of the period, Romania exhibited by far the highest level of inequality, followed by Cyprus and Estonia, while the EU27 average stood around 0.25. In contrast, the lowest GII levels were found in Sweden, Denmark, Finland, and the Netherlands, a pattern that still holds in 2024.

By 2024, the countries with the highest GII values are Cyprus, Romania, and Hungary, while the EU27 average has fallen to approximately 0.08, reflecting substantial progress over time. This aligns with the descriptive statistics: Romania has the highest mean GII (≈ 0.36), while Sweden and Denmark are among the lowest, with mean values of ≈ 0.05 and ≈ 0.054, respectively.

## 6. Coefficient of human inequality per country

```{r, echo=FALSE}
#| message: false
#| warning: false
p_coef <- ggplot(coef_ineq, aes(x = TIME_PERIOD, y = coef_ineq_value, color = iso3, group = iso3)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Coefficient of human inequality per Country (EU27)",
    x = "Year",
    y = "Coefficient of human inequality",
    color = "Country"
  ) +
  theme_minimal()

ggplotly(p_coef)
```

```{r, echo=FALSE}
#| message: false
#| warning: false
summary_stats_coefineq <- coef_ineq |>
  group_by(iso3) |>                        
  summarise(
    Min = min(coef_ineq_value, na.rm = TRUE),
    Max = max(coef_ineq_value, na.rm = TRUE),
    Mean = mean(coef_ineq_value, na.rm = TRUE),
    Median = median(coef_ineq_value, na.rm = TRUE),
    SD = sd(coef_ineq_value, na.rm = TRUE)
    )

kable(summary_stats_coefineq, 
      caption = "Descriptive statistics of the coefficient of human inequality by country (all years combined)")
```

The coefficient of human inequality varies widely across Europe. Spain records the highest values overall, with a mean of ≈ 13.48, while Slovenia and the Czech Republic stand out as the countries with the lowest inequality, both with mean values slightly above 5.

The EU27 average sits at ≈ 8.64, but the dispersion across countries is notable. Several Southern and Eastern European countries—such as Portugal (≈ 12.75), Italy (≈ 12.08), and Bulgaria (≈ 11.84)—exhibit consistently high levels of human inequality. In contrast, Nordic and Central European countries such as Denmark (≈ 5.77) and Finland (≈ 5.73) maintain some of the lowest inequality levels across the data set.

# Linear Regressions

## 1. Between taxes and Greenhouse Gases emissions

We try to answer one of our fundamental question : "Do taxes reduce emissions ?".

### a. Between taxes (n) and GHG (n)

First we try to evaluate the coefficient between an increase in taxes in the year n and the impact on Greenhouse gases emissions in the same year n.

```{r, echo=FALSE}
#| message: false
#| warning: false
model <- lm(GHG_Emissions_value ~ Envt_Tax_Revenues_value + GDP_value + TIME_PERIOD, data = all_merged)

summary(model)
```
We estimated a linear regression to examine the relationship between environmental tax revenues and greenhouse gas emissions, while controlling for GDP and year. The model explains about 84% of the variation in emissions (Adjusted $R^2$ = 0.8456).

The coefficient on environmental tax revenues is -1.029, and it is highly statistically significant (p \< 0.001). This suggests that, holding GDP and time constant, higher environmental tax revenues are associated with slightly lower greenhouse gas emissions. However, the magnitude of the effect remains very small.

GDP has a positive and small significant effect on emissions, with a coefficient of 0.2355 (p \< 0.001). This indicates that countries with higher economic output tend to emit more, consistent with scale effects: larger economies produce more, and thus emit more.

The coefficient on the time variable is -4,484, highly significant (p \< 0.001), indicating a robust downward trend in emissions over time. 

We do not calculate a log - level regression, because our following analysis and tax proposition will be made accordingly to levels of GHG.

### b. Between taxes (n) and GHG (n+k, given k € \[1; 5\])

Now we try to evaluate the coefficient between an increase in taxes in the year n and the impact on Greenhouse gases emissions in the following years n+1, to n+5.

```{r, echo=FALSE}
#| message: false
#| warning: false
# Create leads 1 to 5
for (k in 1:5) {
  GHG_Emissions[[paste0("GHG_Emissions_value.lead", k)]] <- ave(
    GHG_Emissions$GHG_Emissions_value, 
    GHG_Emissions$iso3, 
    FUN = function(x) c(x[(1+k):length(x)], rep(NA, k)))}

# add them in all merged
all_merged_plus_leads <- merge(all_merged, GHG_Emissions,
  by = c("iso3", "TIME_PERIOD", "GHG_Emissions_value"),
  all.x = TRUE)
all_merged_plus_leads <- all_merged_plus_leads |>
    select(iso3, TIME_PERIOD, GDP_value, Envt_Tax_Revenues_value, GHG_Emissions_value, GHG_Emissions_value.lead1, GHG_Emissions_value.lead2, GHG_Emissions_value.lead3, GHG_Emissions_value.lead4, GHG_Emissions_value.lead5 )
```

```{r, echo=FALSE}
#| message: false
#| warning: false
reg_results <- lapply(1:5, function(k) {formula_k <- as.formula(
    paste0("GHG_Emissions_value.lead", k,
           " ~ Envt_Tax_Revenues_value + GDP_value + TIME_PERIOD"))
  lm(formula_k, data = all_merged_plus_leads)})

lapply(reg_results, summary)
```

We estimated a series of linear regressions to examine whether environmental tax revenues in year n have an effect on greenhouse gas emissions in subsequent years, from n+1 to n+5. Across the five models, the number of observations decreases gradually as the lead increases, because later lead variables contain more missing data. Despite this reduction in sample size, all models explain roughly 84% of the variation in emissions ($R^2$ between 0.8444 and 0.835), which is highly consistent across specifications.

Across the five specifications, the coefficient on environmental tax revenues is negative, statistically significant, and remarkably stable. It increases from –1.516 at horizon n+1 to -2.083 at n+4, and then slightly decreases to -2.012 at n+5. This pattern suggests that higher environmental tax revenues are systematically associated with lower future GHG emissions. Statistical significance is very strong in all models (p \< 0.001).

All other coefficients are close to the ones saw before (GDP around 0.25, and time has coefficients between –4,320 and –4,552, all highly significant).

## 2. Between taxes and social inequalities

```{r, echo=FALSE}
#| message: false
#| warning: false
# taxes on inability to warm house (economic inequalities)
model_warm_tax <- lm(Warm_Households_value ~ Envt_Tax_Revenues_value +  GII_value + coef_ineq_value + hhtyp + incgrp + log(GDP_value) + TIME_PERIOD, data = all_merged)
summary(model_warm_tax)

# taxes on the coefficient of inequalities
model_ineq_tax <- lm(coef_ineq_value ~ Envt_Tax_Revenues_value + log(GDP_value) + TIME_PERIOD, data = all_merged)
summary(model_ineq_tax)

# taxes on the gender inequality index
model_gii_tax <- lm(GII_value ~ Envt_Tax_Revenues_value + log(GDP_value) + TIME_PERIOD, data = all_merged)
summary(model_gii_tax)

```

We estimate three regressions to evaluate the association between environmental tax revenues and three social outcomes, in percentage: energy poverty, income inequality, and gender inequality. In all models, we also add GDP and time (years). We chose to not include the Gender Inequality index and the Income inequality coefficient in one each others regression because of collinearity problems. Also, the coefficient of energy poverty is rather a consequence of both than a cause, so we do not include it in each as well.

In the model explaining energy poverty, we include the income-inequality coefficient. The resulting model has strong explanatory power (Adjusted $R^2$ ≈ 0.60). The coefficient on environmental tax revenues remains small, positive, stable (+3.7e-05) and significant. Also, the income-inequality coefficient shows strong positive and significant associations with energy poverty.

For income inequality, the model remains statistically valid but with limited explanatory power (Adjusted $R^2$ ≈ 0.05), indicating that environmental taxes explain only a small part of the variation in inequality (+3.0e-05 ) once GDP and time effects are controlled for.

Similarly, in the gender-inequality model (log GII) again, the inclusion of GDP and time fixed effects strengthens the model substantially (Adjusted $R^2$ ≈ 0.40), and the environmental-tax coefficient remains positive (+6.9e-07) and highly significant.

## 3. Between tax and economy

We try to answer the following question: "What is the impact of environmental taxes on the economy ?"

```{r, echo=FALSE}
#| message: false
#| warning: false
model_gdp_tax <- lm(log(GDP_value) ~ Envt_Tax_Revenues_value + TIME_PERIOD, data = all_merged)
summary(model_gdp_tax)
```

The regression of GDP on environmental tax revenues and year fixed effects shows a strong positive and statistically significant association between the two variables.The coefficient on environmental tax revenues is 6.02 × 10⁻⁵ and is significant at the 1% level (p \< 0.001), holding the time trend constant.

The model has substantial explanatory power (Adjusted $R^2$ ≈ 0.63). The year coefficient is also positive and significant, reflecting general GDP growth over the period. But this relationship is correlational and does not imply causation, it indicates that environmental tax revenues tend to increase alongside GDP, consistent with the idea that wealthier economies can collect more from climate-related taxes.

# Forecasting : no-policy

We use this forecast to try to predict the impact of environmental taxes on different variables.

### 1. Greenhouse Gas emissions (#forecasttotalGHGemissions)

```{r, echo=FALSE}
#| message: false
#| warning: false
EUGHG_agg_yearly <- EU27_GHG |>
  filter(!is.na(GHG_Emissions_value)) |>
  group_by(TIME_PERIOD) |>
  summarise(GHG_total = sum(GHG_Emissions_value, na.rm = TRUE))

ts_data <- ts(EUGHG_agg_yearly$GHG_total, start=min(EUGHG_agg_yearly$TIME_PERIOD))
model <- auto.arima(ts_data)
fc <- forecast(model, h=27)
fc

auto.arima(ts_data)
```

```{r}
autoplot(fc)
```

This ARIMA-based forecast indicates a steady but slow decline in total greenhouse gas emissions for the aggreagte EU27 from 2024 to 2050. The ARIMA(0,1,0) with drift models GHG emissions as a random walk with a linear trend, using first differencing (integrated) but no auto regressive (AR) or moving average (MA) terms. The drift of -52,365 indicates emissions decrease on average by \~52 million tons of CO₂-equivalent per year. Emissions fall from roughly 2.85 billion tons in 2024 to about 1.49 billion tons by 2050, representing a long-term downward trajectory.

However, despite this continuous decrease, the decline is far from the EU’s net-zero objective for 2050. Even in the most optimistic confidence interval, projected emissions in 2050 remain well above 0 (about 363 million tons of CO₂-equivalent), suggesting that under current trends and without major additional interventions—the EU27 is not on track to reach carbon neutrality.

### 2. Social inequalities:

```{r, echo=FALSE}
#| message: false
#| warning: false

gii_ts <- ts(EU27_GII$GII_value, start = min(EU27_GII$TIME_PERIOD))
gii_model <- auto.arima(gii_ts)
gii_fc <- forecast(gii_model, h = 27)

coefineq_ts <- ts(EU27_Coefineq$coef_ineq_value, start = min(EU27_Coefineq$TIME_PERIOD))
coefineq_model <- auto.arima(coefineq_ts)
coefineq_fc <- forecast(coefineq_model, h = 27)

EU27_Warming_agg <- EU27_Warming |> 
  filter(!is.na(Warm_Households_value)) |> 
  group_by(TIME_PERIOD) |> 
  summarise(Warm_avg = mean(Warm_Households_value, na.rm = TRUE))

warm_ts <- ts(EU27_Warming_agg$Warm_avg, start = min(EU27_Warming_agg$TIME_PERIOD))
warm_model <- auto.arima(warm_ts)
warm_fc <- forecast(warm_model, h = 26)


gii_fc
coefineq_fc
warm_fc
```


```{r, echo=FALSE}
#| message: false
#| warning: false
plot(gii_fc, main = "Forecast of GII")
plot(coefineq_fc, main = "Forecast of Coefficient of Inequality")
plot(warm_fc, main = "Forecast of Warm Households")
```

##### Comment :

GII (Gender Inequality Index): The forecast projects a decline from 0.085 in 2023 to 0 (it goes in the negative but the index is between 0 and 1) in 2050.
Coefficient of Inequality: Inequality is expected to decrease from 7.73 in 2023 to 5.04 in 2050, reflecting a gradual improvement in income distribution with moderate uncertainty bands widening over time.
Warm Households: The proportion of households unable to heat properly is forecast to remain relatively stable, rising slightly from 14.65% in 2024 to 16.41% in 2050, though with considerable uncertainty as shown by the wide confidence intervals.

### 3. The economy : forecast of the GDP

```{r, echo=FALSE}
#| message: false
#| warning: false
gdp_ts <- ts(EU27_GDP$GDP_value, start = min(EU27_GDP$TIME_PERIOD))
gdp_model <- auto.arima(gdp_ts)
gdp_fc <- forecast(gdp_model, h = 27)

plot(gdp_fc, main = "Forecast of GDP")
```

##### Comment :

The EU27 aggregate GDP is projected to grow steadily from €19.4 trillion in 2024 to €29.5 trillion in 2050, representing approximately 50% growth over the period, though the widening confidence intervals indicate increasing uncertainty in long-term projections.

### 4. The economy : forecast of the environmental taxes {#forecastofenvttax}

```{r, echo=FALSE}
#| message: false
#| warning: false
# Linear trend model
tax_lm <- lm(Envt_Tax_Revenues_value ~ TIME_PERIOD, data = EU27_Tax)
summary(tax_lm)

# Forecast
future_years <- tibble(
  TIME_PERIOD = (max(EU27_Tax$TIME_PERIOD)+1):(max(EU27_Tax$TIME_PERIOD)+27)
)

tax_pred <- predict(
  tax_lm,
  newdata = future_years,
  interval = "confidence", 
  level = 0.95
)

tax_pred_80 <- predict(
  tax_lm,
  newdata = future_years,
  interval = "confidence", 
  level = 0.80
)

tax_fc_lm <- bind_cols(
  future_years, 
  as.data.frame(tax_pred_95),
  lwr_80 = as.data.frame(tax_pred_80)$lwr,
  upr_80 = as.data.frame(tax_pred_80)$upr
)

```

```{r}
tax_all_2050 <- ggplot() +
  geom_line(data=EU27_Tax, aes(x = TIME_PERIOD, y = Envt_Tax_Revenues_value, color = "Observations")) +
  geom_ribbon(data=tax_fc_lm, aes(x = TIME_PERIOD, ymin = lwr, ymax = upr), fill = "#C8C8C8", alpha = 0.5) +
  geom_ribbon(data=tax_fc_lm, aes(x = TIME_PERIOD, ymin = lwr_80, ymax = upr_80), fill = "#9999CC", alpha = 0.5) +
  geom_line(data=tax_fc_lm, aes(x = TIME_PERIOD, y = fit, color = "Forecast")) +
  scale_color_manual(name = "Legend",
                     values = c("Observations" = "black", 
                               "Forecast" = "steelblue")) +
  theme_minimal() +
  labs(
    title = "Forecast of Environmental Tax Revenues (95% CI)",
    x = "Year",
    y = "Tax revenues"
  ) +
  scale_y_continuous(limits = c(0, 1e+06)) 

ggplotly(tax_all_2050)
```


##### Comment :

Environmental tax revenues in the EU27 are projected to increase from €362.4 billion in 2023 to €617.4 billion in 2050, representing a 70% growth based on an extrapolation of current trends. This steady upward trajectory suggests continued expansion of green taxation policies, however, it's important to note that this forecast assumes the continuation of existing trends under a no-change-in-policy scenario.

# Forecasting: if we plan to get to the 0 net emission by 2050

In this project, we are trying to look at the level of tax that would be necessary to reach the 0 net emissions by 2050 required by the European Union. To do so, we should first look at the level of tax that allows for this evolution, and then check the impact of this level of tax on the social inequalities and on the economy.

### 1. Deriving a realistic tax path to reach net-zero by 2050

#### a. The forecast model

To identify a tax trajectory consistent with the European Union’s goal of net-zero greenhouse gas emissions by 2050, we first estimate a linear regression relating aggregate GHG emissions to environmental tax revenues at the EU-27 level. The 2023 tax level is anchored to the last observed value, while a large increase in 2024 is introduced to ensure that predicted emissions reach zero by 2050. After 2024, the tax path follows a convex trajectory, meaning that annual tax increases accelerate over time, reflecting the progressively stronger measures needed to achieve net-zero emissions. Without this initial jump, emissions would remain far above zero (around 719,773 thousand tons of CO₂ equivalent). This method produces a stylized tax path under the assumption that environmental taxation is the primary policy instrument driving emissions reductions.

```{r, echo=FALSE}
#| message: false
#| warning: false
# --- DATA PREP ---
data_forecast_tax_0net <- EUGHG_agg_yearly |>
  left_join(EU27_Tax, by = "TIME_PERIOD") |>
  filter(!is.na(GHG_total), !is.na(Envt_Tax_Revenues_value)) |>
  mutate(year = as.numeric(TIME_PERIOD))

mod_LL <- lm(GHG_total ~ Envt_Tax_Revenues_value, data = data_forecast_tax_0net)
summary(mod_LL)
alpha <- coef(mod_LL)[1]
beta <- coef(mod_LL)[2]

if(beta >= 0) stop("Regression slope >= 0: tax does not reduce GHG.")

# --- PARAMETERS ---
years <- 2023:2050
n_years <- length(years) # length = 28

GHG_last <- tail(data_forecast_tax_0net$GHG_total, 1)
tax_2023_obs <- tail(data_forecast_tax_0net$Envt_Tax_Revenues_value, 1)
alpha_star <- GHG_last - beta * tax_2023_obs
tax_2050 <- (0 - alpha_star) / beta

# Convexity parameter
gamma <- 2

# Normalized time for 2024→2050
t_norm <- 0:(n_years - 2) / (n_years - 2)  # length = 27

# Compute 2024 tax to start trajectory
tax_2024 <- tax_2023_obs + (tax_2050 - tax_2023_obs)/2  # large jump

# Build smooth convex path for 2024→2050 (length = 27)
tax_smooth <- tax_2024 + (tax_2050 - tax_2024) * t_norm^gamma

# Combine 2023 + 2024→2050
tax_required <- c(tax_2023_obs, tax_smooth)  # length = 28, matches years

# Predicted GHG
pred_GHG <- alpha_star + beta * tax_required

# Build tibble
tax_path <- tibble(
  year = years,
  target_GHG = seq(GHG_last, 0, length.out = n_years),
  tax_required = tax_required,
  pred_GHG = pred_GHG
)

# check
tax_path <- tax_path |>
  mutate(pred_GHG = predict(mod_LL, newdata = tibble(Envt_Tax_Revenues_value = tax_required)))

# Plot
ggplot(tax_path, aes(x = year)) +
  geom_line(aes(y = tax_required), color = "darkblue") +
  labs(title="Required tax path (level-level model)", y="Tax (millions USD)", x="Year") +
  theme_minimal()

# Check predicted
ggplot(tax_path, aes(x = year)) +
  geom_line(aes(y = pred_GHG), color = "darkblue") +
  labs(title="Targeted GHG (level-level)", y="GHG", x="Year") +
  theme_minimal()

```

The estimated regression reveals a statistically significant negative relationship between environmental tax revenues and total GHG emissions. The coefficient on environmental tax revenues is −3.864 (p-value < 0.001), indicating that, historically, an increase of one million USD in environmental tax revenues is associated with a reduction of approximately 3.9 thousand tonnes of CO₂-equivalent in aggregate emissions. The intercept suggests a baseline emissions level of about 5,027,000 thousand tonnes of CO₂-equivalent in the absence of environmental taxation. While the model is intentionally parsimonious, it explains a meaningful share of the observed variation in emissions, with an Adjusted $R^2$ of 0.49.

Using this estimated relationship, we construct a counterfactual emissions trajectory from approximately 2,907,204 thousand tonnes in 2023 to net-zero emissions in 2050. To ensure that predicted emissions reach net zero, the 2023 tax level is anchored to the observed value (362,363 million USD), while a substantial increase in 2024 is introduced. After 2024, the tax path follows a convex trajectory toward 2050, reflecting the progressively stronger fiscal effort needed to reduce emissions to zero. Without this initial jump, predicted emissions would plateau around 719,773 thousand tonnes of CO₂-equivalent by 2050, far above net zero.

According to the tax path, environmental tax revenues would need to rise from 362,363 million USD in 2023 to 738,542 millions USD in 2024, to about 1,114,722 million USD in 2050, producing the targeted emissions trajectory illustrated in the GHG emissions path graph. These results highlight both the scale of the fiscal effort required and the limitations of relying solely on environmental taxation, given the modest explanatory power of the reduced-form regression.

#### b. Comparing with the no-policy forecast

Having established the tax path required to reach net-zero emissions by 2050, we now turn to a systematic comparison of forecasted and targeted trajectories. First, we will examine the total environmental tax revenues to highlight the gap between historical trends and climate targets. Next, we will assess the implications of these tax paths for aggregate GHG emissions, and finally, we will explore sector-level GHG emissions to identify which sectors require the most significant reductions. This stepwise analysis provides a comprehensive view of the fiscal and environmental adjustments necessary to meet the EU’s net-zero objectives.

##### Tax policy : targeted vs predicted

To assess the adequacy of current environmental tax trends, we now compare the forecasted path of total EU-27 environmental tax revenues, based on historical trends (the “forecast model” from the section: [The economy : forecast of the environmental taxes](#forecastofenvttax)), with the tax path required to reach net-zero emissions by 2050. This comparison highlights the gap between the continuation of past tax policies and the ambitious fiscal effort necessary to achieve the EU’s climate target. The code below generates the forecasted confidence intervals and overlays them with the required tax path.

```{r, echo=FALSE}
#| message: false
#| warning: false
tax_pred_95 <- predict(
  tax_lm,
  newdata = future_years,
  interval = "confidence",
  level = 0.95
)
tax_pred_80 <- predict(
  tax_lm,
  newdata = future_years,
  interval = "confidence",
  level = 0.80
)
tax_fc_lm_ci <- future_years |>
  mutate(
    mean = tax_pred_95[, "fit"],
    lower95 = tax_pred_95[, "lwr"],
    upper95 = tax_pred_95[, "upr"],
    lower80 = tax_pred_80[, "lwr"],
    upper80 = tax_pred_80[, "upr"]
  ) |>
  rename(year = TIME_PERIOD)
tax_plot_df <- tax_fc_lm_ci |>
  left_join(
    tax_path |> select(year, tax_required),
    by = "year"
  )

ggplot(tax_plot_df, aes(x = year)) +
  geom_ribbon(
    aes(ymin = lower95, ymax = upper95, fill = "95% confidence interval"),
    alpha = 0.35
  ) +
  geom_ribbon(
    aes(ymin = lower80, ymax = upper80, fill = "80% confidence interval"),
    alpha = 0.5
  ) +
  geom_line(
    aes(y = mean, color = "Forecasted tax revenues (trend continuation)"),
    linewidth = 1
  ) +
  geom_line(
    aes(y = tax_required, color = "Tax path required for net-zero by 2050"),
    linewidth = 1
  ) +
  scale_color_manual(
    name = "Lines",
    values = c(
      "Forecasted tax revenues (trend continuation)" = "red",
      "Tax path required for net-zero by 2050" = "darkblue"
    )
  ) +
  scale_fill_manual(
    name = "Confidence intervals",
    values = c(
      "80% confidence interval" = "grey50",
      "95% confidence interval" = "grey70"
    )
  ) +
  labs(
    title = "Environmental Tax Revenues: Trend Forecast vs Climate-Required Path",
    x = "Year",
    y = "Tax revenues (millions USD)"
  ) +
  scale_y_continuous(limits = c(0, 1.5e6)) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )
```

Comparing the two paths shows a significant and growing shortfall. The historical trend regression predicts that total environmental tax revenues would increase by approximately 14,767 million USD per year (intercept = −29,038,537; $R^2 = 0.73$), corresponding to a continuation of past trends. In contrast, the tax path required to achieve net-zero emissions calls for a steep and consistent annual increase of roughly 65,5 million USD per year, starting from about 1,2 trillion USD in 2023 and reaching 2,9 trillion USD by 2050. Already in 2024, forecasted revenues (≈ 851 billion USD) fall well below the required level (≈ 1,242 billion USD), and by 2050 the gap widens to roughly 1,7 trillion USD, indicating that current policy trajectories imply structural under-taxation. These results underscore that meeting the net-zero target would require a substantial policy departure from historical trends and a rapid scaling up of environmental taxation.

##### GHG emissions : targeted vs predicted

Having analyzed the environmental tax paths, we now turn to the environmental outcome they aim to influence: greenhouse gas (GHG) emissions. This subsection compares the emissions trajectory implied by a continuation of historical tax trends with the counter factual path required to achieve net-zero emissions by 2050.

Using the ARIMA model (cf. [Time series ARIMA forecast of the total Greenhouse Gas emissions](#forecasttotalGHGemissions)), we forecast total emissions under a continuation of historical trends and compute confidence intervals (80% and 95%) to capture uncertainty in the projection. We then overlay the **targeted emissions trajectory**, which represents the linear decline needed to reach net-zero by 2050, as derived from the inverted tax–GHG relationship. The resulting graph allows a visual comparison between the likely path of emissions if current policies continue and the ambitious reduction trajectory required by the EU’s climate objectives.

```{r, echo=FALSE}
#| message: false
#| warning: false
fc_df <- data.frame(
  year = as.numeric(time(fc$mean)),
  mean = as.numeric(fc$mean),
  lower80 = as.numeric(fc$lower[, 1]),
  upper80 = as.numeric(fc$upper[, 1]),
  lower95 = as.numeric(fc$lower[, 2]),
  upper95 = as.numeric(fc$upper[, 2])
)

ggplot(tax_path, aes(x = year)) +
  geom_ribbon(data = fc_df, aes(
      x = year,
      ymin = lower95,
      ymax = upper95,
      fill = "95% confidence interval"),
    inherit.aes = FALSE,
    alpha = 0.35) +
  geom_ribbon(data = fc_df, aes(
      x = year,
      ymin = lower80,
      ymax = upper80,
      fill = "80% confidence interval"),
    inherit.aes = FALSE,
    alpha = 0.5) +
  geom_line(data = fc_df, aes(
    x = year, y = mean, color = "ARIMA forecast (no-policy)"),
    inherit.aes = FALSE,
    linewidth = 1) +
  geom_line(
    aes(y = target_GHG, color = "Targeted GHG (if we plan to get zero net emissions)"),
    linewidth = 1) +
  scale_color_manual(
    name = "Lines",
    values = c(
      "Targeted GHG (if we plan to get zero net emissions)" = "darkblue",
      "ARIMA forecast (no-policy)" = "red")) +
  scale_fill_manual(
    name = "Confidence intervals",
    values = c(
      "80% confidence interval" = "grey50",
      "95% confidence interval" = "grey70")) +
  labs(
    title = "GHG Emissions: Target Path vs ARIMA Forecast (2023–2050)",
    x = "Year",
    y = "GHG Emissions") +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )
```
As shown in the figure, achieving the targeted GHG emissions trajectory would require a substantial change in current taxing practices. The target path consistently lies at the lower bound of the forecast confidence intervals for emissions, and from around 2035 onwards it even falls outside the 95% confidence interval. This suggests that, under the continuation of current trends captured by the ARIMA forecast, the probability of reaching the targeted emissions level is low, highlighting the need for stronger or additional policy interventions.

### 2. Impact on social inequalities

#### a. The forecast model

```{r, echo=FALSE}
#| message: false
#| warning: false
# Average yearly tax increase
avg_tax_increment <- (last(tax_path$tax_required) -
                      first(tax_path$tax_required)) /
                      (nrow(tax_path) - 1)

# Last observed levels
last_obs <- tibble(
  Warm = tail(EU27_Warming_agg$Warm_avg, 1),
  Ineq = tail(EU27_Coefineq$coef_ineq_value, 1),
  GII  = tail(EU27_GII$GII_value, 1)
)

future_years <- length(tax_path$year)

# Extract coefficients
b_warm_tax  <- coef(model_warm_tax)["Envt_Tax_Revenues_value"]
b_warm_time <- coef(model_warm_tax)["TIME_PERIOD"]

b_ineq_tax  <- coef(model_ineq_tax)["Envt_Tax_Revenues_value"]
b_ineq_time <- coef(model_ineq_tax)["TIME_PERIOD"]

b_gii_tax   <- coef(model_gii_tax)["Envt_Tax_Revenues_value"]
b_gii_time  <- coef(model_gii_tax)["TIME_PERIOD"]

# Build paths
future_predictions <- tibble(
  year = tax_path$year,

  Predicted_Warming =
    last_obs$Warm +
    cumsum(rep(b_warm_tax * avg_tax_increment + b_warm_time, future_years)),

  Predicted_Inequality =
    last_obs$Ineq +
    cumsum(rep(b_ineq_tax * avg_tax_increment + b_ineq_time, future_years)),

  Predicted_GII =
    pmax(0, pmin(1,last_obs$GII +
        cumsum(rep(b_gii_tax * avg_tax_increment + b_gii_time, future_years))
      )))
```

```{r, echo=FALSE}
#| message: false
#| warning: false
# Plot 1: Warming inability
p_warming <- ggplot(future_predictions, aes(x = year, y = Predicted_Warming)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  scale_y_continuous(limits = c(0, 50)) +
  labs(title = "Forecasted impact on Average Warming Inability", x = "Year", y = "Warming") +
  theme_minimal()

# Plot 2: Coefficient of Inequality
p_inequality <- ggplot(future_predictions, aes(x = year, y = Predicted_Inequality)) +
  geom_line(color = "orange", linewidth = 1) +
    scale_y_continuous(limits = c(0, 50)) +
  labs(title = "Forecasted impact on the Coefficient of Human Inequality", x = "Year", y = "Gini Coefficient") +
  theme_minimal()

# Plot 3: Gender Inequality Index
p_gii <- ggplot(future_predictions, aes(x = year, y = Predicted_GII)) +
  geom_line(color = "purple", linewidth = 1) +
    scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Forecasted impact on the Gender Inequality Index (GII)", x = "Year", y = "GII") +
  theme_minimal()

# Stack vertically using grid.arrange
grid.arrange(p_warming, p_inequality, p_gii, ncol = 1)
```


#### b. Comparing with the no-policy forecast

```{r, echo=FALSE}
#| message: false
#| warning: false
warm_fc_df <- data.frame(
  year = as.numeric(time(warm_fc$mean)),
  mean = as.numeric(warm_fc$mean),
  lower80 = as.numeric(warm_fc$lower[, 1]),
  upper80 = as.numeric(warm_fc$upper[, 1]),
  lower95 = as.numeric(warm_fc$lower[, 2]),
  upper95 = as.numeric(warm_fc$upper[, 2])
)

ggplot() +
  geom_ribbon(
    data = warm_fc_df,
    aes(x = year, ymin = lower95, ymax = upper95, fill = "95% confidence interval"),
    alpha = 0.35
  ) +
  geom_ribbon(
    data = warm_fc_df,
    aes(x = year, ymin = lower80, ymax = upper80, fill = "80% confidence interval"),
    alpha = 0.5
  ) +
  geom_line(
    data = warm_fc_df,
    aes(x = year, y = mean, color = "ARIMA forecast (no tax policy)"),
    linewidth = 1
  ) +
  geom_line(
    data = future_predictions,
    aes(x = year, y = Predicted_Warming, color = "Forecasted impact on Average Warming Inability"),
    linewidth = 1
  ) +
  scale_color_manual(
    name = "Scenarios",
    values = c(
      "ARIMA forecast (no tax policy)" = "red",
      "Forecasted impact on Average Warming Inability" = "darkgreen"
    )
  ) +
  scale_fill_manual(
    name = "Confidence intervals",
    values = c(
      "80% confidence interval" = "grey50",
      "95% confidence interval" = "grey70"
    )
  ) +
  labs(
    title = "Average Warming Inability: Tax Policy vs No-Policy Scenario",
    x = "Year",
    y = "Average Warming Inability"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )
```


```{r, echo=FALSE}
#| message: false
#| warning: false
ineq_fc_df <- data.frame(
  year = as.numeric(time(coefineq_fc$mean)),
  mean = as.numeric(coefineq_fc$mean),
  lower80 = as.numeric(coefineq_fc$lower[, 1]),
  upper80 = as.numeric(coefineq_fc$upper[, 1]),
  lower95 = as.numeric(coefineq_fc$lower[, 2]),
  upper95 = as.numeric(coefineq_fc$upper[, 2])
)

ggplot() +
  geom_ribbon(
    data = ineq_fc_df,
    aes(x = year, ymin = lower95, ymax = upper95, fill = "95% confidence interval"),
    alpha = 0.35
  ) +
  geom_ribbon(
    data = ineq_fc_df,
    aes(x = year, ymin = lower80, ymax = upper80, fill = "80% confidence interval"),
    alpha = 0.5
  ) +
  geom_line(
    data = ineq_fc_df,
    aes(x = year, y = mean, color = "ARIMA forecast (no tax policy)"),
    linewidth = 1
  ) +
  geom_line(
    data = future_predictions,
    aes(x = year, y = Predicted_Inequality, color = "Forecasted impact on the Coefficient of Human Inequality"),
    linewidth = 1
  ) +
  scale_color_manual(
    name = "Scenarios",
    values = c(
      "ARIMA forecast (no tax policy)" = "red",
      "Forecasted impact on the Coefficient of Human Inequality" = "orange"
    )
  ) +
  scale_fill_manual(
    name = "Confidence intervals",
    values = c(
      "80% confidence interval" = "grey50",
      "95% confidence interval" = "grey70"
    )
  ) +
  labs(
    title = "Coefficient of Human Inequality: Tax Policy vs No-Policy Scenario",
    x = "Year",
    y = "Coefficient of Human Inequality"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )
```


```{r, echo=FALSE}
#| message: false
#| warning: false
gii_fc_df <- data.frame(
  year = as.numeric(time(gii_fc$mean)),
  mean = as.numeric(gii_fc$mean),
  lower80 = as.numeric(gii_fc$lower[, 1]),
  upper80 = as.numeric(gii_fc$upper[, 1]),
  lower95 = as.numeric(gii_fc$lower[, 2]),
  upper95 = as.numeric(gii_fc$upper[, 2])
)

ggplot() +
  geom_ribbon(
    data = gii_fc_df,
    aes(x = year, ymin = lower95, ymax = upper95, fill = "95% confidence interval"),
    alpha = 0.35
  ) +
  geom_ribbon(
    data = gii_fc_df,
    aes(x = year, ymin = lower80, ymax = upper80, fill = "80% confidence interval"),
    alpha = 0.5
  ) +
  geom_line(
    data = gii_fc_df,
    aes(x = year, y = mean, color = "ARIMA forecast (no tax policy)"),
    linewidth = 1
  ) +
  geom_line(
    data = future_predictions,
    aes(x = year, y = Predicted_GII, color = "Forecasted impact on the GII"),
    linewidth = 1
  ) +
  scale_color_manual(
    name = "Scenarios",
    values = c(
      "ARIMA forecast (no tax policy)" = "red",
      "Forecasted impact on the GII" = "purple"
    )
  ) +
  scale_fill_manual(
    name = "Confidence intervals",
    values = c(
      "80% confidence interval" = "grey50",
      "95% confidence interval" = "grey70"
    )
  ) +
  labs(
    title = "Gender Inequality Index: Tax Policy vs No-Policy Scenario",
    x = "Year",
    y = "GII"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )
```


### 3. Is the model viable ? Its impact on the GDP (forecast & comparison)

```{r, echo=FALSE}
#| message: false
#| warning: false

# Average yearly tax increase
avg_tax_increment <- (last(tax_path$tax_required) -
                      first(tax_path$tax_required)) /
                      (nrow(tax_path) - 1)

# Last observed levels
last_obs <- tibble(GDP = tail(EU27_GDP$GDP_value, 1))

future_years <- length(tax_path$year)

# Extract coefficients
b_gdp_tax  <- coef(model_gdp_tax)["Envt_Tax_Revenues_value"]
b_gdp_time <- coef(model_gdp_tax)["TIME_PERIOD"]

# Build paths
future_gdp_predictions <- tibble(
  year = tax_path$year,
  Predicted_GDP =
    last_obs$GDP +
    cumsum(rep(b_gdp_tax * avg_tax_increment + b_gdp_time, future_years)))

ggplot(future_gdp_predictions,
                aes(x = year, y = Predicted_GDP)) +
  geom_line(color = "steelblue", linewidth = 1) +
  labs(
    title = "Forecasted impact of the tax policy on EU27 GDP",
    x = "Year",
    y = "GDP"
  ) +
  theme_minimal()
```



```{r, echo=FALSE}
#| message: false
#| warning: false

gdp_fc_df <- data.frame(
  year = as.numeric(time(gdp_fc$mean)),
  mean = as.numeric(gdp_fc$mean),
  lower80 = as.numeric(gdp_fc$lower[, 1]),
  upper80 = as.numeric(gdp_fc$upper[, 1]),
  lower95 = as.numeric(gdp_fc$lower[, 2]),
  upper95 = as.numeric(gdp_fc$upper[, 2])
)

ggplot() +
  geom_ribbon(
    data = gdp_fc_df,
    aes(x = year, ymin = lower95, ymax = upper95, fill = "95% confidence interval"),
    alpha = 0.35
  ) +
  geom_ribbon(
    data = gdp_fc_df,
    aes(x = year, ymin = lower80, ymax = upper80, fill = "80% confidence interval"),
    alpha = 0.5
  ) +
  geom_line(
    data = gdp_fc_df,
    aes(x = year, y = mean, color = "ARIMA forecast (no tax policy)"),
    linewidth = 1
  ) +
  geom_line(
    data = future_gdp_predictions,
    aes(x = year, y = Predicted_GDP, color = "Forecasted impact on Aggregate GDP"),
    linewidth = 1
  ) +
  scale_color_manual(
    name = "Scenarios",
    values = c(
      "ARIMA forecast (no tax policy)" = "red",
      "Forecasted impact on Aggregate GDP" = "darkblue"
    )
  ) +
  scale_fill_manual(
    name = "Confidence intervals",
    values = c(
      "80% confidence interval" = "grey50",
      "95% confidence interval" = "grey70"
    )
  ) +
  labs(
    title = "Aggregate GDP: Tax Policy vs No-Policy Scenario",
    x = "Year",
    y = "Aggregate GDP"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )
```


##### Ratio Tax/GDP

```{r, echo=FALSE}
#| message: false
#| warning: false
# Build GDP forecast dataframe
gdp_forecast_df <- tibble(
  year = seq(max(EU27_GDP$TIME_PERIOD) + 1,
             max(EU27_GDP$TIME_PERIOD) + 27),
  GDP_forecast = as.numeric(gdp_fc$mean)
)

# Required tax path → tax / GDP
required_data <- tax_path |>
  left_join(gdp_forecast_df, by = "year") |>
  mutate(
    tax_gdp_ratio = (tax_required / GDP_forecast) * 100,
    scenario = "Required tax-GDP ratio"
  ) |>
  select(year, tax_gdp_ratio, scenario)

# Forecasted tax path → tax / GDP
forecasted_data <- as.data.frame(tax_pred) |>
  mutate(year = tax_path$year[-1]) |>
  select(year, fit) |>
  left_join(gdp_forecast_df, by = "year") |>
  mutate(
    tax_gdp_ratio = (fit / GDP_forecast) * 100,
    scenario = "Forecasted tax-GDP ratio"
  ) |>
  select(year, tax_gdp_ratio, scenario)



# Combine both paths
comparison_data <- bind_rows(required_data, forecasted_data)

#Viability zones
zones <- tibble(
  ymin = c(0, 1, 3, 5),
  ymax = c(1, 3, 5, Inf),
  zone = c("Viable", "Caution", "Risk", "High Risk")
)

# Plot: Required vs Forecasted Tax as % of GDP
ggplot() +
  geom_rect(data = zones,
            aes(xmin = -Inf, xmax = Inf, ymin = ymin, ymax = ymax, fill = zone),
            alpha = 0.2) +
  geom_line(data = comparison_data,
            aes(x = year, y = tax_gdp_ratio, color = scenario),
            linewidth = 1.2) +
  geom_point(data = comparison_data,
             aes(x = year, y = tax_gdp_ratio, color = scenario),
             size = 2, alpha = 0.7) +
  geom_hline(yintercept = c(1, 3, 5),
             linetype = "dashed", color = "gray40") +
  scale_color_manual(
    values = c(
      "Required tax-GDP ratio"   = "darkblue",
      "Forecasted tax-GDP ratio" = "firebrick"
    ),
    name = "Scenario"
  ) +
  scale_fill_manual(
    values = c(
      "Viable" = "green",
      "Caution" = "yellow",
      "Risk" = "orange",
      "High Risk" = "red"
    ),
    name = "Viability zone"
  ) +

  labs(
    title = "Economic Viability of Environmental Tax Policy",
    subtitle = "Required vs Forecasted Environmental Tax as % of GDP",
    x = "Year",
    y = "Tax as % of GDP"
  ) +

  scale_y_continuous(breaks = seq(0, 10, 1)) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "right"
  )
```
The thresholds (1%, 3%, 5%) are based on empirical observations from existing tax systems.
We say that above 5% it would be Very Disruptive: At this scale, capital flight risks increase, sectoral lobbying intensifies, and social acceptance becomes critical. 

# Conclusion

In conclusion, this research project has evaluated the effectiveness of environmental taxation as a tool for achieving the European Union’s goal of net-zero emissions by 2050. The study involved cleaning and merging large datasets covering the EU-27 from 1990 to 2024, focusing on four key sectors: Agriculture, Industry (including energy), Manufacturing, and Transport. By integrating indicators such as environmental tax revenues, the Gender Inequality Index (GII), the Coefficient of Human Inequality, and energy poverty, we assessed both environmental efficacy and socioeconomic sustainability.
Our findings indicate that while environmental taxes have a statistically significant negative association with GHG emissions, the impact has been small. 
Our analysis revealed that the current forecasted tax trajectory is insufficient to reach the European Green Deal’s objectives. Under current trends as GHG emissions are projected to fall only to approximately 1.43 million units by 2050. 
To determine the tax level required for the 2050 target, we inverted a level-level linear regression of total GHG emissions against tax revenues. This model allowed us to find a "required tax path" that would theoretically drive emissions to zero by itself. We found that environmental tax revenues would need to scale up to roughly 2.9 trillion USD by 2050, (1.7 trillion USD more than predicted historical trends).

The impacts of such a massive tax increase on society and the economy would be severe:
• Social Inequality: The regressions show that environmental taxes are positively correlated with energy poverty, income inequality, and gender inequality. Implementing the required tax path would likely exacerbate these social costs, and increase these indicators largely above the levels expected in a no-policy scenario.
• Economic Viability: Although tax revenues tend to increase alongside GDP, the required fiscal burden would likely reach between 6% and 10% of the total GDP. This exceeds empirical thresholds for economic stability and could be highly disruptive.

## Opening

If relying exclusively on environmental taxation proves to be socially regressive or economically unviable, what alternative strategies could the European Union use to fill the gap? Some possibilities could be the use of complementary green technology subsidies, the implementation of stricter sector-specific regulations, or the potential for economic sobriety and degrowth to reduce the overall energy demand.
