---
title: "dm-project"
author: "Daphné Perry and Justine Catta"
format: html
---

# Summary 

# Introduction

## Description of our research question

Our research question is the following: “To what extent do environmental taxes reduce Greenhouse Gas Emissions (GHG) in EU countries, and what are their long-term socioeconomic impacts across economic sectors?”. As the European Union commits to achieving net-zero emissions by 2050 under the European Green Deal, we aim to better understand the effectiveness and consequences of environmental taxation, both on emissions and socially.

To answer this question, we will use sector-level panel data for all EU member states. Our data set includes GHG emissions, sectoral GDP (measured as Added Value), environmental tax revenues, and several indicators of social well-being. These include the Gender Inequality Index (GII), a coefficient of human inequality, and a measure of energy poverty that is the inability to adequately heat one’s home. Together, these variables allow us to evaluate the distributive consequences of taxes.

After cleaning, harmonizing, and describing the data sets, we aim to estimate several relationships. First, we will measure the link between environmental taxation and sectoral GHG emissions. Do higher taxes lead to statistically significant reductions in emissions at both the EU and country levels? Are some sectors more responsive to taxation than others? To maintain analytical clarity and comparability across states, we will focus on four key sectors: Agriculture, Energy, Manufacturing, and Transport, which together account for the bulk of emissions and policy intervention.

We will then investigate the association between environmental taxes and social inequality indicators. Our core question here can be summarized as: “Is there a social cost associated with environmental taxation, and if so, how large is it?” Specifically, we will analyse whether environmental taxes correlate with increases or decreases in gender inequality, income inequality, and energy poverty. This step is essential because policies that are environmentally effective but socially regressive may have a reduced long-term viability.

Building on these empirical relationships, we will construct forecasts for the period 2026–2050 to estimate the level of taxation required to meet the European Green Deal’s emissions targets. We consider a policy scenario relying solely on environmental taxation applied to relevant sectors. Our analysis will assess whether such an approach is socially and economically sustainable: How would different tax levels affect inequalities and energy poverty? How much revenue would such taxes generate, and would the required tax burden be compatible with each country’s GDP levels?

By combining environmental effectiveness, economic feasibility, and social impact, we can then address our overarching question: Is it socially and economically optimal for the EU to rely exclusively on environmental taxation to achieve net-zero emissions by 2050? If taxation alone proves insufficient or too socially costly, our results may highlight the need for complementary policies such as subsidies for green technologies, targeted social transfers, or sector-specific regulations to ensure a fair and effective ecological transition.


## Link to the github project

Our project is available on github [here](https://github.com/cattajustine-spec/project).

### Setup

```{r, echo=FALSE}
#| message: false
library(vroom)
library(ggplot2)
library(knitr)
library(dplyr)
library(here)
library(countrycode)
library(plotly)
library(tidyr)
library(knitr)
```

```{r, echo=FALSE}
#| message: false
here::i_am("project.Rproj")
```


## Presenting our data :

### 1. Environmental tax revenues (OECD, 1994–2023) 

  This data set reports annual environmentally related tax revenues for EU countries, broken into categorises based on their environmental relevance by tax-base category (including energy, transport, pollution, and resources) that are mutually exclusive. According to the OECD, the database contains detailed qualitative and quantitative information on environmentally related taxes and fees, tradable permits and offsets, deposit-refund schemes, environmentally beneficial subsidies and payments and voluntary approaches used for environmental policy.
Values are expressed in millions of USD. It is from the OECD and is available on this [link](https://data-explorer.oecd.org/vis?lc=en&tm=DF_ERTR&pg=0&snb=1&vw=tb&df%5Bds%5D=dsDisseminateFinalDMZ&df%5Bid%5D=DSD_ERTR%40DF_ERTR&df%5Bag%5D=OECD.ENV.EPI&df%5Bvs%5D=&pd=1994%2C2023&dq=A.HRV%2BLVA%2BLTU%2BLUX%2BMLT%2BNLD%2BCYP%2BAUT%2BBEL%2BCZE%2BDNK%2BEST%2BFIN%2BFRA%2BDEU%2BGRC%2BHUN%2BIRL%2BITA%2BPOL%2BPRT%2BSVK%2BSVN%2BESP%2BSWE%2BBGR%2BROU.TAXREV.CAT_ENE%2BCAT_TRA%2BCAT_POL%2BCAT_RES%2B_T._T.USD.&ly%5Brw%5D=REF_AREA&ly%5Bcl%5D=TIME_PERIOD&to%5BTIME_PERIOD%5D=false). They cover a time period from 1994 to 2023.


```{r, echo=FALSE}
#| message: false
Envt_Tax_Revenues <- vroom(here("data", "OECD_envt_tax.csv"))
```

::: {columns}
|                    |                                           |
|--------------------|-------------------------------------------|
| Name               | `r "Envt_Tax_Revenues"`                   |
| Number of rows     | `r nrow(Envt_Tax_Revenues)`               |
| Number of columns  | `r ncol(Envt_Tax_Revenues)`               |
:::


### 2. GDP and Added Value (OECD, 1990-2024)

  This data set provides annual GDP and sectoral gross Value Added figures for EU countries. It follows an output approach : GDP is measured as the sum of gross value added. For the Value Added by sectors, we will keep these ones : "Agriculture,forestry and fishing", "Manufacturing", "Industry (except construction)" and "Wholesale and retail trade; repair of motor vehicles and motorcycles; transportation and storage; accommodation and food service activities". (cf plus tard) All variables are measured in millions of USD, using exchange rate. 
 The data set on GDP and Value Added is available at this [link](https://data-explorer.oecd.org/vis?df%5Bds%5D=dsDisseminateFinalDMZ&df%5Bid%5D=DSD_NAMAIN10%40DF_TABLE1_OUTPUT&df%5Bag%5D=OECD.SDD.NAD&df%5Bvs%5D=&dq=A.BEL%2BAUT%2BCZE%2BDNK%2BEST%2BFIN%2BFRA%2BDEU%2BGRC%2BHUN%2BIRL%2BITA%2BLVA%2BLTU%2BLUX%2BNLD%2BPOL%2BPRT%2BSVK%2BSVN%2BESP%2BSWE%2BBGR%2BHRV%2BCYP%2BMLT%2BROU...B1G%2BB1GQ.._Z%2B_T%2BC%2BA%2BBTE%2BF%2BGTI%2BJ%2BK%2BL%2BM_N%2BOTQ%2BRTU..USD_EXC.V..&to%5BTIME_PERIOD%5D=false&pd=1990%2C2024&vw=tb). 
The data covers the time period 1990 to 2024.


```{r, echo=FALSE}
#| message: false
GDP_and_AV <- vroom(here("data", "OECD_GDP_and_AV.csv"))
```

::: {columns}
|                    |                       |
|--------------------|-----------------------|
| Name               | `r "GDP_and_AV"`      |
| Number of rows     | `r nrow(GDP_and_AV)`  |
| Number of columns  | `r ncol(GDP_and_AV)`  |
:::

### 3. Greenhouse gas emissions by source sector (Eurostat, 1990-2023)

  This data set reports total greenhouse gas emissions by source sector for the 27 eu countries. Values are expressed in thousand tonnes of CO₂-equivalent. The GHG source sectors we chose are the following ones, to be able to match the Added Value sectors (energy, transport, agriculture, industry): Total, Fuel combustion in energy industries, Fuel combustion in manufacturing industries and construction, Industrial processes and product use, agriculture,  fuel combustion in transport
The data set on Greenhouse gas emissions is [here](https://ec.europa.eu/eurostat/databrowser/view/env_air_gge__custom_18942108/default/table).
The data covers the time period 1990 to 2023.
```{r, echo=FALSE}
#| message: false
GHG_Emissions <- vroom(here("data", "Eurostat_ghg.csv"))
```

::: {columns}
|                    |                          |
|--------------------|--------------------------|
| Name               | `r "GHG_Emissions"`      |
| Number of rows     | `r nrow(GHG_Emissions)`  |
| Number of columns  | `r ncol(GHG_Emissions)`  |
:::

### 4.  Inability to keep home adequately warm (Eurostat, 2003-2024)

  This data set measures the percentage of households unable to keep their home adequately warm. It reflects energy poverty and distributional effects of energy pricing across income groups and household types. We chose 3 households type : the total, single female and single male to be able to have an idea of the gender issues. We also chose two income situations : the total (all incomes) and the group below 60% of the median (in situation of poverty), because the income inequality is not available in the Human Index.
The data set on Inability to keep home adequately warm is linked [here](https://ec.europa.eu/eurostat/databrowser/view/ILC_MDES01__custom_18945942/default/table). The time period covered is 2003-2024.

```{r, echo=FALSE}
#| message: false
Warm_Households <- vroom(here("data","Eurostat_warm.csv"))
```

::: {columns}
|                    |                            |
|--------------------|----------------------------|
| Name               | `r "Warm_Households"`      |
| Number of rows     | `r nrow(Warm_Households)`  |
| Number of columns  | `r ncol(Warm_Households)`  |
:::

### 5.  Human index (UNDP HDR, 1990-2023)

  This data set provides annual values for lots of Human Development Indexes. We focus on the Coefficient of Human Inequality  and Gender Inequality Index (GII) for most countries worldwide. It combines indicators related to health, education, income, and gender disparities, enabling long-run social development comparisons.

Available on this [page](HDR%20index%20https://hdr.undp.org/data-center/documentation-and-downloads) from the UNDP Human Development Report (HDR) "All composite indices and components time series (1990-2023)" or directly with this [link](https://hdr.undp.org/sites/default/files/2025_HDR/HDR25_Composite_indices_complete_time_series.csv). 
The time period of this data set is 1990 to 2023. 

```{r, echo=FALSE}
#| message: false
Soc_Indexes <- vroom(here("data", "HDR_indexes.csv"))
```


::: {columns}
|                    |                        |
|--------------------|------------------------|
| Name               | `r "Soc_Indexes"`      |
| Number of rows     | `r nrow(Soc_Indexes)`  |
| Number of columns  | `r ncol(Soc_Indexes)`  |
:::

# Modifying the data sets

Because the data sets we chose are quite huge, we first need to remove all the variables we will not use, then add missing variables that we think are important for our analysis, and then modify the variables we kept.

Finally, to simplify our analysis, we will merge all our data sets in one data base. 

### Country keys {#countrykeys}

  Our analysis is on the 27 countries in the European Union, in the following list to be precise : Austria, Belgium, Bulgaria, Croatia, Cyprus, Czechia, Denmark, Estonia, Finland, France, Germany, Greece, Hungary, Ireland, Italy, Latvia, Lithuania, Luxembourg, Malta, Netherlands, Poland, Portugal, Romania, Slovakia, Slovenia, Spain, Sweden

```{r, echo=FALSE}
eu_countries <- unique(GHG_Emissions$geo)

eu_country_key <- data.frame(name_en = eu_countries) |>
  mutate(
    iso3 = countrycode(name_en, origin = "country.name", destination = "iso3c"),
    iso2 = countrycode(name_en, origin = "country.name", destination = "iso2c"),
    name_fr = countrycode(name_en, origin = "country.name", destination = "country.name.fr")
  ) 
```

### Removing unwanted variables

##### 1. Environmental tax revenues (in million of USD)

In this data set, we are only interested in the following variables:

-   **REF_AREA**: Identifies the country or reporting area using an ISO3 code (e.g., LVA for Latvia, LTU for Lithuania). This variable is the geographic dimension of the data set. 
-   **TIME_PERIOD**: Indicates the reference year of observation (from 1994 to 2023).
-   **OBS_VALUE**: Contains the numerical value of environmentally related tax revenues.
-   **Catégorie de recettes provenant des taxes liées à l'environnement**: A label in French describing the tax revenue category in plain language. In this data set, there are five mutually-exclusive categories:
    -   "Assiette relatives aux taxes sur l'énergie", i.e., energy-related taxes.
    -   "Assiette relatives aux taxes sur le transport", i.e., transport-related taxes.
    -   "Assiette relatives aux taxes sur les pollutions", i.e., pollution-related taxes.
    -   "Assiette relatives aux taxes sur les resources", i.e., resource-related taxes.
    -   “Toute assiette de taxes”, i.e., all environmental tax bases combined.

```{r, echo=FALSE}
Envt_Tax_Revenues <- Envt_Tax_Revenues |>
  select(REF_AREA, TIME_PERIOD, OBS_VALUE, 
         `Catégorie de recettes provenant des taxes liées à l'environnement`) |>
  filter(REF_AREA %in% eu_country_key$iso3)
```

##### 2. GDP and Value Added per capita annual growth per sector (in million of USD)

In this data set, we have two main variables of interest: the [Gross Domestic Product](#grossdomesticproduct) and the [Gross Added Value](#grossaddedvalue).

They are described by the **Transaction** variable, therefore, we need to create two independent data sets, one for each variable of interest.

###### Gross Domestic product {#grossdomesticproduct}

In this data set, we are only interested in the following variables:

-   **REF_AREA**: Identifies the country or reporting area using an ISO3 code. This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 1990 to 2024).
-   **OBS_VALUE**: Contains the numerical value of the GDP, for a given time and year.

```{r, echo=FALSE}
GDP <- GDP_and_AV |>
  filter(Transaction == "Produit intérieur brut", 
         REF_AREA %in% eu_country_key$iso3) |>
  select(REF_AREA, TIME_PERIOD, OBS_VALUE)
```

###### Gross Added Value {#grossaddedvalue}

In this data set, we are only interested in the following variables:

-   **REF_AREA**: Identifies the country or reporting area using an ISO3 code. This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 1990 to 2024).
-   **Activité économique**: A label in French describing the economic activity or industry sector associated with the added value produced. In this data set, we are only interested in four sectors (cf. [Justification for the Sector Mapping Between ISIC and CRF Classifications](#justificationforsectormapping)):
    -   "Activités de fabrication", i.e., manufacturing.
    -   "Agriculture, sylviculture et pêche", i.e., agriculture.
    -   "Industrie (sauf construction)", i.e., public electricity and heat production.
    -   "Commerce de gros et de détail; réparation de véhicules automobiles et de motocyclettes; transport et stockage; activités d'hébergement et de restauration", i.e., transport.\
-   **OBS_VALUE**: Contains the numerical value of the gross added value.

```{r, echo=FALSE}
Added_value <- GDP_and_AV |>
  filter(Transaction == "Valeur ajoutée, brute", 
         REF_AREA %in% eu_country_key$iso3, 
         `Activité économique` %in% c(
           "Agriculture, sylviculture et pêche",
           "Activités de fabrication",
           "Industrie (sauf construction)",
           "Commerce de gros et de détail; réparation de véhicules automobiles et de motocyclettes; transport et stockage; activités d'hébergement et de restauration"
           )) |>
  select(REF_AREA, `Activité économique`, 
         TIME_PERIOD, OBS_VALUE)
```

##### 3. Greenhouse gas emissions by source sector (in million of ton)

In this data set, we are only interested in the following variables:

-   **geo**: Identifies the country or reporting area with its English name. This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 1990 to 2024).
-   **src_crf**: Describes the economic activity or industry sector associated with the greenhouse gases emitted. In this data set, we are only interested in five sectors (cf. [Justification for the Sector Mapping Between ISIC and CRF Classifications](#justificationforsectormapping)):
    -   "Agriculture".
    -   "Fuel combustion in energy industries", i.e., public electricity and heat production.
    -   "Fuel combustion in manufacturing industries and construction" and “Industrial processes and product use”, i.e., manufacturing (we will combine this sectors in the [Renaming some variables](#renamingsomevariables) section).
    -   "Fuel combustion in transport", i.e., transport.
-   **OBS_VALUE**: Contains the numerical value of the Greenhouse Gases emitted.

```{r, echo=FALSE}
GHG_Emissions <- GHG_Emissions |>
  select(airpol, src_crf, geo, TIME_PERIOD, OBS_VALUE) |>
  filter(geo %in% eu_country_key$name_en, 
         src_crf != "Total (excluding memo items)")
```

###### Justification for the Sector Mapping Between ISIC and CRF Classifications {#justificationforsectormapping}

The reconciliation of sectors across the ISIC-based data set and the CRF (Common Reporting Format) classification relies on aligning economic activities with their dominant emission sources. Although the two systems are not structurally identical—ISIC being an economic activity classification and CRF being an emissions reporting framework—several robust correspondences can be established.

First, Agriculture, forestry and fishing (ISIC A) is mapped to CRF 3 (Agriculture). Both refer to primary-sector biological production and include emissions directly linked to agricultural practices such as enteric fermentation, manure management, and agricultural soils.

Second, Manufacturing (ISIC C) is associated with CRF 1.A.2 (Manufacturing industries and construction) and CRF 2 (Industrial processes and product use). Manufacturing activity generates emissions both from fuel combustion in industrial facilities (covered by 1.A.2) and from process-related sources such as cement, chemicals, or metals production (covered by CRF 2), making this linkage consistent with the structure of industrial emissions.

Third, Electricity, gas, steam and air conditioning supply (ISIC D)—even though aggregated within “Industry (except construction)” in the first data set—logically corresponds to CRF 1.A.1 (Public electricity and heat production). This CRF category captures emissions from power generation and heat production, which are the core activities of ISIC D.

Finally, Transportation and storage (ISIC H)—included in a broader aggregated category in the data set—is aligned with CRF 1.A.3 (Transport). This CRF category encompasses fuel combustion from road, rail, aviation, and navigation transport, which directly reflects the operational scope of ISIC H.

Overall, the mapping follows a principle of dominant emission source equivalence, ensuring that each ISIC activity is linked to the CRF category representing the primary origin of its greenhouse gas emissions. This approach allows the two data sets—despite structural differences—to be compared and integrated in a coherent and meaningful way for our analysis.

##### 4. Inability to keep home adequately warm

In this data set, we are only interested in the following variables:

-   **geo**: Identifies the country or reporting area using a ISO2 code. This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 2003 to 2024).
-   **hhtyp**: Indicates the composition of the household, here we are only interested in:
    -   "A1F", i.e., the households composed by only one female.
    -   "A1M", i.e., the households composed by only one male.
    -   "TOTAL", i.e., all households (whatever their composition).
-   **incrp**: Indicates the income group of the household, here we are only interested in:
    -   "B_MD60", i.e., the households who have an income below 60% of the median income (meaning they are below the poverty threshold).
    -   "TOTAL", i.e., all households (whatever their revenue).
-   **OBS_VALUE**: Contains the percentage of household who report that they are unable to keep their home adequately warm, within the designated group.

```{r, echo=FALSE}
Warm_Households <- Warm_Households |>
  select(geo, TIME_PERIOD, hhtyp, incgrp, OBS_VALUE) |>
  filter(geo %in% eu_country_key$iso2, 
         hhtyp %in% c("A1F", "A1M", "TOTAL"))
```

##### 5. HDR index

In this data set, we have two main variables of interest: the [Gender Inequality Index](#genderinequalityindex) and the [Coefficient of Human Inequality](#coefficientofhumaninequality).

They are described per year as different variables (meaning we have a gii_2020, gii_2021, ..., coef_ineq_2003, coef_ineq_2004, etc.). Therefore, we need to pivot our database, and then create two independent data sets, one for each variable of interest.

```{r, echo=FALSE}
Soc_Indexes <- Soc_Indexes |>
  pivot_longer(
    cols = matches("_[0-9]{4}$"),   
    names_to = c("Indicator", "TIME_PERIOD"),
    names_pattern = "^(.*)_(\\d{4})$",
    values_to = "OBS_VALUE"
  ) |>
  mutate(
    OBS_VALUE = as.numeric(OBS_VALUE),
    TIME_PERIOD = as.integer(TIME_PERIOD)
  )
```

###### Gender inequality Index {#genderinequalityindex}

In this data set, we are only interested in the following variables:

-   **iso3**: Identifies the country or reporting area using an ISO3 code. This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 1990 to 2023).
-   **OBS_VALUE**: Contains the numerical value of the GII.

```{r, echo=FALSE}
GII <- Soc_Indexes|>
  filter(iso3 %in% eu_country_key$iso3,
         Indicator == "gii") |>
  select(iso3, TIME_PERIOD, OBS_VALUE)
```

###### Coefficient of Human inequality {#coefficientofhumaninequality}

In this data set, we are only interested in the following variables:

-   **iso3**: Identifies the country or reporting area using an ISO3 code. This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 2010 to 2023).
-   **OBS_VALUE**: Contains the numerical value of the coefficient of human inequality.

```{r, echo=FALSE}
coef_ineq <- Soc_Indexes|>
  filter(iso3 %in% eu_country_key$iso3,
         Indicator == "coef_ineq")|>
  select(iso3, TIME_PERIOD, OBS_VALUE)
```

### Adding some entities

#### Add European average (across countries)
For the data sets that are not described per economic sector, we compute the European average across countries:

```{r, echo=FALSE}
#| message: false
EU27_GDP <- GDP |>                #computing the average European GDP 
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(REF_AREA = "EU27") |>
  relocate(REF_AREA)
GDP <- bind_rows(GDP, EU27_GDP)

EU27_Tax <- Envt_Tax_Revenues |>   #computing the average European environmental tax revenue, depending on the tax category. 
  group_by(TIME_PERIOD,
           `Catégorie de recettes provenant des taxes liées à l'environnement`) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(REF_AREA = "EU27") |>
  relocate(REF_AREA)
Envt_Tax_Revenues <- bind_rows(Envt_Tax_Revenues, EU27_Tax)

EU27_Warming <- Warm_Households |>  #computing the average European inability to keep home adequately warm, depending on the household composition and income.  
  group_by(TIME_PERIOD, hhtyp, incgrp) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(geo = "EU27") |>
  relocate(geo)
Warm_Households <- bind_rows(Warm_Households, EU27_Warming)

EU27_GII <- GII |>  #computing the average European gender inequality index. 
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(iso3 = "EU27") |>
  relocate(iso3)
GII <- bind_rows(GII, EU27_GII)

EU27_Coefineq <- coef_ineq |> #computing the average European coefficient of human inequality. 
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(iso3 = "EU27") |>
  relocate(iso3)
coef_ineq <- bind_rows(coef_ineq, EU27_Coefineq)
```

#### Add country average (accross sectors)

For the data sets that are described per economic sector, we have to compute the European average across sectors:

```{r, echo=FALSE}
#| message: false
EU27_AV <- Added_value |> #computing the average European gross added value, depending on the economic sector.  
  group_by(TIME_PERIOD, `Activité économique`) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(REF_AREA = "EU27") |>
  relocate(REF_AREA)
Added_value <- bind_rows(Added_value, EU27_AV)

EU27_GHG <- GHG_Emissions |> #computing the average European greenhouse gas emissions, depending on the economic sector.  
  group_by(TIME_PERIOD, src_crf) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(geo = "EU average") |>
  relocate(geo)
GHG_Emissions <- bind_rows(GHG_Emissions, EU27_GHG)
```

### Modifiying the characteristics of the variables

For the variables related to the economic sectors, it will be easier to consider them as *factor*, rather than as *character* variables.

```{r, echo=FALSE}
GHG_Emissions$src_crf <- factor(GHG_Emissions$src_crf)
EU27_GHG$src_crf <- factor(EU27_GHG$src_crf)
Added_value$`Activité économique` <- factor(Added_value$`Activité économique`)
EU27_AV$`Activité économique` <- factor(EU27_AV$`Activité économique`)
```

Same reasoning for the household types in the Warm_Households data set, and for the tax categories in the Envt_Tax_Revenues data set. 

```{r}
Warm_Households$hhtyp <- factor(Warm_Households$hhtyp)
EU27_Warming$hhtyp <- factor(EU27_Warming$hhtyp)
Warm_Households$incgrp <- factor(Warm_Households$incgrp)
EU27_Warming$incgrp <- factor(EU27_Warming$incgrp)
Envt_Tax_Revenues$`Catégorie de recettes provenant des taxes liées à l'environnement` <- factor(Envt_Tax_Revenues$`Catégorie de recettes provenant des taxes liées à l'environnement`)
EU27_Tax$`Catégorie de recettes provenant des taxes liées à l'environnement` <- factor(EU27_Tax$`Catégorie de recettes provenant des taxes liées à l'environnement`)
```

##### Renaming some variables {#renamingsomevariables}

To be able to merge our data sets in a data base, we need to modify the name of the variable **OBS_VALUE** that is in every data set to be able to identify which OBS_VALUE is associated to which data set.

```{r, echo=FALSE}
names(Warm_Households)[names(Warm_Households) == "OBS_VALUE"] <- "Warm_Households_value"
names(EU27_Warming)[names(EU27_Warming) == "OBS_VALUE"] <- "Warm_Households_value"
names(Envt_Tax_Revenues)[names(Envt_Tax_Revenues) == "OBS_VALUE"] <- "Envt_Tax_Revenues_value"
names(EU27_Tax)[names(EU27_Tax) == "OBS_VALUE"] <- "Envt_Tax_Revenues_value"
names(GDP)[names(GDP) == "OBS_VALUE"] <- "GDP_value"
names(EU27_GDP)[names(EU27_GDP) == "OBS_VALUE"] <- "GDP_value"
names(Added_value)[names(Added_value) == "OBS_VALUE"] <- "AV_value"
names(EU27_AV)[names(EU27_AV) == "OBS_VALUE"] <- "AV_value"
names(GHG_Emissions)[names(GHG_Emissions) == "OBS_VALUE"] <- "GHG_Emissions_value"
names(EU27_GHG)[names(EU27_GHG) == "OBS_VALUE"] <- "GHG_Emissions_value"
names(GII)[names(GII) == "OBS_VALUE"] <- "GII_value"
names(EU27_GII)[names(EU27_GII) == "OBS_VALUE"] <- "GII_value"
names(coef_ineq)[names(coef_ineq) == "OBS_VALUE"] <- "coef_ineq_value"
names(EU27_Coefineq)[names(EU27_Coefineq) == "OBS_VALUE"] <- "coef_ineq_value"
```

To be able to merge the data sets, it will be easier to have the tax categories in English.

```{r}
Envt_Tax_Revenues <- Envt_Tax_Revenues |>
  mutate(`Catégorie de recettes provenant des taxes liées à l'environnement` = recode(`Catégorie de recettes provenant des taxes liées à l'environnement`, 
                                                                                      "Assiette relatives aux taxes sur les resources" = "Tax base related to resource taxes", 
                                                                                      "Assiette relatives aux taxes sur les pollutions" = "Tax base related to pollution taxes", 
                                                                                      "Assiette relatives aux taxes sur le transport" = "Tax base related to transport taxes", 
                                                                                      "Assiette relatives aux taxes sur l'énergie" = "Tax base related to energy taxes", 
                                                                                      "Toute assiette de taxes" = "Any tax base"))

EU27_Tax <- EU27_Tax |>
  mutate(`Catégorie de recettes provenant des taxes liées à l'environnement` = recode(`Catégorie de recettes provenant des taxes liées à l'environnement`, 
                                                                                      "Assiette relatives aux taxes sur les resources" = "Tax base related to resource taxes", 
                                                                                      "Assiette relatives aux taxes sur les pollutions" = "Tax base related to pollution taxes", 
                                                                                      "Assiette relatives aux taxes sur le transport" = "Tax base related to transport taxes", 
                                                                                      "Assiette relatives aux taxes sur l'énergie" = "Tax base related to energy taxes", 
                                                                                      "Toute assiette de taxes" = "Any tax base"))
```

To be able to merge the data sets including variables related to the economic sectors, it will be easier to have the same sector names, in English.

```{r, echo=FALSE}
Added_value <- Added_value |>
  mutate(`Activité économique` = recode(`Activité économique`,
                                        "Commerce de gros et de détail; réparation de véhicules automobiles et de motocyclettes; transport et stockage; activités d'hébergement et de restauration" = "Transport", 
                                        "Agriculture, sylviculture et pêche"= "Agriculture", 
                                        "Industrie (sauf construction)" = "Industry (including energy)",
                                        "Activités de fabrication" = "Manufacturing"
                                        ))

EU27_AV <- EU27_AV |>
  mutate(`Activité économique` = recode(`Activité économique`,
                                        "Commerce de gros et de détail; réparation de véhicules automobiles et de motocyclettes; transport et stockage; activités d'hébergement et de restauration" = "Transport", 
                                        "Agriculture, sylviculture et pêche"= "Agriculture", 
                                        "Industrie (sauf construction)" = "Industry (including energy)",
                                        "Activités de fabrication" = "Manufacturing"
                                        ))

GHG_Emissions <- GHG_Emissions|>  #here we are combining the sectors related to manufacturing. 
  mutate(
    src_crf = case_when(
      src_crf %in% c("Fuel combustion in manufacturing industries and construction", 
                     "Industrial processes and product use") ~ "Manufacturing",
      TRUE ~ src_crf
    )
  ) |>
  group_by(geo, TIME_PERIOD, src_crf) |>
  summarise(GHG_Emissions_value = sum(GHG_Emissions_value), .groups = "drop")
GHG_Emissions <- GHG_Emissions |>
  mutate(src_crf = recode(src_crf, 
                          "Fuel combustion in energy industries"="Industry (including energy)", 
                          "Fuel combustion in transport"="Transport"))

EU27_GHG <- EU27_GHG|>
  mutate(
    src_crf = case_when(
      src_crf %in% c("Fuel combustion in manufacturing industries and construction", 
                     "Industrial processes and product use") ~ "Manufacturing",
      TRUE ~ src_crf
    )
  ) |>
  group_by(geo, TIME_PERIOD, src_crf) |>
  summarise(GHG_Emissions_value = sum(GHG_Emissions_value), .groups = "drop")
EU27_GHG <- EU27_GHG |>
  mutate(src_crf = recode(src_crf, 
                          "Fuel combustion in energy industries"="Industry (including energy)", 
                          "Fuel combustion in transport"="Transport"))
```

To be able to merge all the data sets, it will be easier to rename the country codes as the ones in the country key we created in [Country keys](#countrykeys).

We also rename the different sectors in the different data bases as "sector" to be able to merge more easily and have two values for the same sector : one for the Added Value and one for the Greenhouse gases emissions.

```{r}
Envt_Tax_Revenues <- Envt_Tax_Revenues |>
  rename(iso3 = REF_AREA, tax_category = `Catégorie de recettes provenant des taxes liées à l'environnement`)
EU27_Tax <- EU27_Tax |>
  rename(iso3 = REF_AREA, tax_category = `Catégorie de recettes provenant des taxes liées à l'environnement`)

GDP <- GDP |> rename(iso3 = REF_AREA)
EU27_GDP <- EU27_GDP |> rename(iso3 = REF_AREA)

Added_value <- Added_value |>
  rename(iso3 = REF_AREA,
    sector = `Activité économique`) 
EU27_AV <- EU27_AV |>
  rename(iso3 = REF_AREA,
    sector = `Activité économique`) 

GHG_Emissions <- GHG_Emissions |> rename(name_en = geo,
                                         sector = src_crf)
EU27_GHG <- EU27_GHG |> rename(name_en = geo,
                                         sector = src_crf)

Warm_Households <- Warm_Households |> rename(iso2 = geo)
EU27_Warming <- EU27_Warming |> rename(iso2 = geo)
```


# Merging all data sets on country-year pairs

We do this big join to have in only one data set all these variables:

 - Country
 - Year of observation
 - Environmentally related tax revenues
 - GHG emissions by economic sector
 - Added value by economic sector (using the same sector as the GHG emissions one)
 - GDP_value
 - Gender inequality index value
 - Income inequality index value
 - Share of households unable to adequately heat their home, by subgroup :
      - household type
      - income group

```{r}
GHG_Emissions <- GHG_Emissions |>
  left_join(eu_country_key |> select(name_en, iso3), by = "name_en")

Warm_Households <- Warm_Households |>
  left_join(eu_country_key |> select(iso2, iso3), by = "iso2")
```


```{r}
#| message: false
#| warning: false
all_merged <- Envt_Tax_Revenues |>
  left_join(Added_value |> distinct(iso3, TIME_PERIOD, sector), by = c("iso3", "TIME_PERIOD")) |>
  left_join(Added_value, by = c("iso3", "TIME_PERIOD", "sector")) |>
  left_join(GDP, by = c("iso3", "TIME_PERIOD")) |>
  left_join(GHG_Emissions, by = c("iso3", "TIME_PERIOD", "sector")) |>
  left_join(Warm_Households, by = c("iso3", "TIME_PERIOD")) |>
  left_join(GII, by = c("iso3", "TIME_PERIOD")) |>
  left_join(coef_ineq, by = c("iso3", "TIME_PERIOD")) |>
  select(iso3, TIME_PERIOD, tax_category, sector, hhtyp, incgrp, Envt_Tax_Revenues_value, AV_value, GDP_value, GHG_Emissions_value, Warm_Households_value, GII_value, coef_ineq_value)
```

# Descriptive analyzes

Now that we have cleaned our data sets, here are some descriptive analyzes of our main target variables:

## 1. Environmental tax revenues per tax category

### Graphical representation

```{r, echo=FALSE}
ggplot(EU27_Tax, aes(x = TIME_PERIOD, y = Envt_Tax_Revenues_value, color = tax_category, group = tax_category)) +
  geom_line(linewidth = 1) +
  labs(
    title = "EU average Environmental Tax Revenues by Category",
    x = "Year",
    y = "Revenue (million USD)",
    color = "Category"
  ) +
  theme_minimal()
```

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_EnvtTax <- EU27_Tax %>%
  group_by(tax_category) %>%
  summarise(
    Min = min(Envt_Tax_Revenues_value, na.rm = TRUE),
    Max = max(Envt_Tax_Revenues_value, na.rm = TRUE),
    Mean = mean(Envt_Tax_Revenues_value, na.rm = TRUE),
    Median = median(Envt_Tax_Revenues_value, na.rm = TRUE),
    SD = sd(Envt_Tax_Revenues_value, na.rm = TRUE)
  )

kable(summary_stats_EnvtTax, caption = "Descriptive statistics of EU average Environmental Tax Revenues by Category (all years combined)")
```

##### Comment :

The graph shows that, on average across Europe, environmental tax revenues have increased substantially between 1990 and 2023. They initially declined until the early 2000s, falling below 7.5 billion USD, before rising sharply and then stabilizing between 13 and 15 billion USD from the mid-2000s to the early 2020s.

We can also see that the vast majority of these revenues come from energy-related taxes, which may reflect the importance of energy both in emissions and in the European economy. This is consistent with the descriptive statistics, where energy taxes have by far the highest mean value (≈ 8.69 billion USD, compared with only ≈ 0.53 billion USD for pollution taxes).

By contrast, revenues from transport taxes and pollution taxes represent only a small share of the total. Transport tax revenues have increased the most visibly, especially between 2000 and 2005, rising from roughly 1.5 to 2.5 billion USD.

## 2. GDP per country

### Graphical representation

```{r, echo=FALSE}
p_GDP <- ggplot(GDP, aes(x = TIME_PERIOD, y = GDP_value, color = iso3, group = iso3)) +
  geom_line(linewidth = 1) +
  labs(
    title = "GDP per Country (EU27)",
    x = "Year",
    y = "GDP (million USD)",
    color = "Country"
  ) +
  theme_minimal()

ggplotly(p_GDP)
```

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_GDP <- GDP %>%
  group_by(iso3) %>%                        
  summarise(
    Min = min(GDP_value, na.rm = TRUE),
    Max = max(GDP_value, na.rm = TRUE),
    Mean = mean(GDP_value, na.rm = TRUE),
    Median = median(GDP_value, na.rm = TRUE),
    SD = sd(GDP_value, na.rm = TRUE)
    )

kable(summary_stats_GDP, 
      caption = "Descriptive statistics of the GDP by country (all years combined)")
```

##### Comment :

It is quite clear that GDP levels across European countries are generally rising over time. Four major economies lead by a large margin: Germany, France, Italy, and Spain. All of them experienced a noticeable boom around 2002, which may be linked to the consolidation following the introduction of the euro in 1999.

Germany is by far the largest economy, with an average GDP of ≈ 3.1 trillion USD, followed by France (≈ 2.19 trillion USD) and Italy (≈ 1.75 trillion USD). Spain remains lower but still significant, with a mean of ≈ 1.08 trillion USD.

Most other countries remain much smaller, except the Netherlands, which stands out with a mean GDP of ≈ 706 billion USD. Overall, the EU27 average GDP reaches roughly 780 billion USD in 2024, reflecting large disparities across the Union.


## 3. Added Value per economic sector

### Graphical representation
```{r, echo=FALSE}
p_va <- ggplot(EU27_AV, aes(x = TIME_PERIOD, y = AV_value, color = sector, group = sector)) +
  geom_line(linewidth = 1) +
  labs(
    title = "EU27 Average Added Value per Sector",
    x = "Year",
    y = "Value Added (million USD)",
    color = "Sector"
  ) +
  theme_minimal()

ggplotly(p_va)
```

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_AV <- EU27_AV %>%
  group_by(sector) %>%
  summarise(
    Min = min(AV_value, na.rm = TRUE),
    Max = max(AV_value, na.rm = TRUE),
    Mean = mean(AV_value, na.rm = TRUE),
    Median = median(AV_value, na.rm = TRUE),
    SD = sd(AV_value, na.rm = TRUE)
  )

kable(summary_stats_AV, caption = "Descriptive statistics of EU average gross added value by sector (all years combined)")
```

##### Comment :

Looking at the EU27 average gross value added by sector, three major sectors dominate and remain close in magnitude: manufacturing, transport, and industry (with energy related industries such as public electricity and heat production). Each of these sectors shows mean added values between 80 and 96 billion USD, making them central pillars of the European economy.

Agriculture, by contrast, contributes much less, with an average value added of only ≈ 9.5 billion USD, which is expected given its smaller role in modern European economies. The manufacturing sector alone ranges from 46 to 142 billion USD, highlighting both its scale and its variability over time.

## 4. Greenhouse gas emissions by economic sector

### Graphical representation

```{r, echo=FALSE}
p_GHG <- ggplot(EU27_GHG, aes(x = TIME_PERIOD, y = GHG_Emissions_value, color = sector, group = sector)) +
  geom_line(linewidth = 1) +
  labs(
    title = "EU27 Average Greenhouse gas emissions per Sector",
    x = "Year",
    y = "Greenhouse gas emissions (thousands of ton of CO2)",
    color = "Sector"
  ) +
  theme_minimal()

ggplotly(p_GHG)
```

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_GHG <- EU27_GHG %>%
  group_by(sector) %>%                
  summarise(
    Min = min(GHG_Emissions_value, na.rm = TRUE),
    Max = max(GHG_Emissions_value, na.rm = TRUE),
    Mean = mean(GHG_Emissions_value, na.rm = TRUE),
    Median = median(GHG_Emissions_value, na.rm = TRUE),
    SD = sd(GHG_Emissions_value, na.rm = TRUE)
  )

kable(summary_stats_GHG, caption = "Descriptive statistics of EU average greenhouse gases emissions by sector (all years combined)")
```
##### Comment :

The EU27 average data show greenhouse gas emissions across four sectors: agriculture, manufacturing, energy industries, and transport. All sectors display a downward trend except transport, which starts rising again after 2010. With the exception of agriculture, all sectors experienced a marked drop in 2020 due to the COVID-19 shock, followed by a partial rebound.

Industrial emissions, particularly from the energy sector, have declined sharply over the long run (from values reaching above 53,000 kt CO₂-eq to much lower levels). Manufacturing emissions also decreased, though at a more moderate pace, with a mean of ≈ 33,400 kt.

Since around 2009–2012, transport emissions have surpassed those of manufacturing, and by 2023 they also exceeded energy industries emissions. Transport has become the single largest emitting sector, consistent with its relatively high mean value (≈ 29,200 kt CO₂-eq), compared with ≈ 45,000 kt for electricity and heat production (energy) and ≈ 14,900 kt for agriculture.

## 5. Inability to keep home adequately warm by household type

### Graphical representation

```{r, echo=FALSE}
ggplot(EU27_Warming, 
       aes(x = TIME_PERIOD, y = Warm_Households_value,
           color = interaction(hhtyp, incgrp),
           group = interaction(hhtyp, incgrp))) +
  geom_line(linewidth = 1) +
  labs(
    title = "EU27 – inability to Keep the House Warm (by Sex × Income Group)",
    x = "Year",
    y = "Percentage of population",
    color = "Sex × Income"
  ) +
  theme_minimal()
```

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_Warm <- EU27_Warming %>%
  group_by(hhtyp, incgrp) %>%                    
  summarise(
    Min = min(Warm_Households_value, na.rm = TRUE),
    Max = max(Warm_Households_value, na.rm = TRUE),
    Mean = mean(Warm_Households_value, na.rm = TRUE),
    Median = median(Warm_Households_value, na.rm = TRUE),
    SD = sd(Warm_Households_value, na.rm = TRUE), 
    .groups = "drop"
  )

kable(summary_stats_Warm, caption = "Descriptive statistics of the inability to keep house adequately warm, by household type (all years combined)")
```

##### Comment :

For the EU27 average, six curves represent different subgroups defined by sex and income. The values correspond to the percentage of each subgroup unable to keep their home adequately warm. All groups follow roughly the same pattern: a sharp increase between 2003 and 2005, a gradual decline afterward, and two significant shocks: a drop in 2008–2009 and a spike in 2022–2023, which can be linked to rising energy prices following the Russia–EU energy crisis.

Among all subgroups, the total-population categories show the lowest percentages, remaining below 10% since 2015. Conversely, households with incomes below 60% of the national median (below the poverty threshold) consistently show the highest rates, around 20% on average.

Women in low-income single households consistently report slightly higher percentages than men. For instance, low-income women (A1F, B_MD60) have an average inability rate of ≈ 21.1%, compared with ≈ 23.1% for low-income men (A1M, B_MD60), while total-population rates remain much lower at ≈ 10–15% across categories.

## 6. GII per country

### Graphical representation

```{r, echo=FALSE}
p_GII <- ggplot(GII, aes(x = TIME_PERIOD, y = GII_value, color = iso3, group = iso3)) +
  geom_line(linewidth = 1) +
  labs(
    title = "GII per Country (EU27)",
    x = "Year",
    y = "Gender inequality index",
    color = "Country"
  ) +
  scale_y_continuous(limits = c(0, 1)) +   
  theme_minimal()

ggplotly(p_GII)
```

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_GII <- GII %>%
  group_by(iso3) %>%                        
  summarise(
    Min = min(GII_value, na.rm = TRUE),
    Max = max(GII_value, na.rm = TRUE),
    Mean = mean(GII_value, na.rm = TRUE),
    Median = median(GII_value, na.rm = TRUE),
    SD = sd(GII_value, na.rm = TRUE)
    )

kable(summary_stats_GII, 
      caption = "Descriptive statistics of the GII by country (all years combined)")
```
##### Comment :

The Gender Inequality Index (GII) shows a clear downward trend across all European countries between 1990 and 2024. At the beginning of the period, Romania exhibited by far the highest level of inequality, followed by Cyprus and Estonia, while the EU27 average stood around 0.25. In contrast, the lowest GII levels were found in Sweden, Denmark, Finland, and the Netherlands, a pattern that still holds in 2024.

By 2024, the countries with the highest GII values are Cyprus, Romania, and Hungary, while the EU27 average has fallen to approximately 0.08, reflecting substantial progress over time. This aligns with the descriptive statistics: Romania has the highest mean GII (≈ 0.36), while Sweden and Denmark are among the lowest, with mean values of ≈ 0.05 and ≈ 0.054, respectively.

## 7. Coefficient of human inequality per country

### Graphical representation

```{r, echo=FALSE}
p_coef <- ggplot(coef_ineq, aes(x = TIME_PERIOD, y = coef_ineq_value, color = iso3, group = iso3)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Coefficient of human inequality per Country (EU27)",
    x = "Year",
    y = "Coefficient of human inequality",
    color = "Country"
  ) +
  theme_minimal()

ggplotly(p_coef)
```

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_coefineq <- coef_ineq %>%
  group_by(iso3) %>%                        
  summarise(
    Min = min(coef_ineq_value, na.rm = TRUE),
    Max = max(coef_ineq_value, na.rm = TRUE),
    Mean = mean(coef_ineq_value, na.rm = TRUE),
    Median = median(coef_ineq_value, na.rm = TRUE),
    SD = sd(coef_ineq_value, na.rm = TRUE)
    )

kable(summary_stats_coefineq, 
      caption = "Descriptive statistics of the coefficient of human inequality by country (all years combined)")
```

##### Comment :

The coefficient of human inequality varies widely across Europe. Spain records the highest values overall, with a mean of ≈ 13.48, while Slovenia and the Czech Republic stand out as the countries with the lowest inequality, both with mean values slightly above 5.

The EU27 average sits at ≈ 8.64, but the dispersion across countries is notable. Several Southern and Eastern European countries—such as Portugal (≈ 12.75), Italy (≈ 12.08), and Bulgaria (≈ 11.84)—exhibit consistently high levels of human inequality. In contrast, Nordic and Central European countries such as Denmark (≈ 5.77) and Finland (≈ 5.73) maintain some of the lowest inequality levels across the data set.

# Linear Regressions
MAYBE ADD CORRELATION GRAPHS !!

## 1. Between taxes and Greenhouse Gases emissions

We try to answer one of our fundamental question : "Do taxes reduce emissions ?"

### a. Between taxes (n) and GHG (n)

First we try to evaluate the coefficient between an increase in taxes in the year n and the impact on Greenhouse gases emissions in the same year n.

```{r, echo=FALSE}
model <- lm(GHG_Emissions_value ~ Envt_Tax_Revenues_value + GDP_value + sector + TIME_PERIOD,
            data = all_merged)

summary(model)
```

##### Comment :

We estimated a linear regression to examine the relationship between environmental tax revenues and greenhouse gas emissions, while controlling for GDP, emission sector, and year. The model explains about 65% of the variation in emissions (R-squared = 0.6542).

The coefficient on environmental tax revenues is –0.043, and it is highly statistically significant (p < 0.001). This suggests that, holding GDP, sector, and time constant, higher environmental tax revenues are associated with slightly lower greenhouse gas emissions. However, the magnitude of the effect remains very small, implying a limited practical impact.

GDP has a positive and very large significant effect on emissions, with a coefficient of 0.043 (p < 0.001). This indicates that countries with higher economic output tend to emit more, consistent with scale effects: larger economies produce more goods and services, and thus more emissions.

Sectoral effects are strong and highly significant. Relative to the baseline category (Agriculture), estimated additional emissions are:
- Manufacturing: +16,440 units
- Energy industry: +27,370 units
- Transport: +16,150 units
These substantial differences reflect the much higher emission intensities of industrial, energy, and transport activities compared with agriculture.

The coefficient on the time variable is –913, highly significant (p < 0.001), indicating a robust downward trend in emissions over time. This is consistent with what we saw before in the descriptive data.

Overall, the results suggest that GDP and sectoral composition remain the dominant drivers of emissions. Environmental tax revenues show a statistically significant negative association with emissions, but the magnitude of the effect is small. The pronounced negative time trend indicates that greenhouse gas emissions have declined over the period independently of the direct influence of environmental tax revenues.

### b. Between taxes (n) and GHG (n+k, given k € [1; 5] to have enough data)

Now we try to evaluate the coefficient between an increase in taxes in the year n and the impact on Greenhouse gases emissions in the following years n+1, to n+5.

```{r}

# Create leads 1 to 15
for (k in 1:15) {
  GHG_Emissions[[paste0("GHG_Emissions_value.lead", k)]] <- ave(
    GHG_Emissions$GHG_Emissions_value, 
    GHG_Emissions$name_en, GHG_Emissions$sector,
    FUN = function(x) c(x[(1+k):length(x)], rep(NA, k))
  )
}

```

##### Comment :

### c. Between tax category and GHG emissions

Here we try to evaluate the coefficient between an increase in tax in one specific area (environmentally related taxes are grouped by what sector they impact : transport, energy, resource or pollution) in the year n and the impact on Greenhouse gases emissions in the year n+k, most efficient lag.

```{r}
all_merged$tax_category <- as.factor(all_merged$tax_category)

model_taxcat <- lm(GHG_Emissions_value ~ tax_category + GDP_value + sector + TIME_PERIOD,
                   data = all_merged)

summary(model_taxcat)
```

##### Comment :

We estimated a linear regression to examine the relationship between different types of environmental taxes and greenhouse gas emissions, while controlling for GDP, sector, and year. The model explains about 65.8% of the variation in emissions (R-squared = 0.658), slightly higher than the previous model that considered total tax revenues.

The coefficients on all tax categories—energy, pollution, resource, and transport taxes—are effectively zero and not statistically significant (p = 1 for all categories). This indicates that, after accounting for GDP, sector, and time, there is no detectable relationship between any specific type of tax and GHG emissions in this data set.

Other coefficient are roughly the same as before.

Answering the question, based on this regression, no specific type of environmental tax appears to be more efficient at reducing GHG emissions, as none of the tax category coefficients are statistically significant. This suggests that, at the observed levels and in this dataset, other factors—especially GDP, sectoral composition, and time trends—dominate emission outcomes.

### d. Between tax category and GHG sector

Here we try to evaluate the coefficient between an increase in tax in one specific area (environmentally related taxes are grouped by what sector they impact : transport, energy, resource or pollution) in the year n and the impact on Greenhouse gases emissions in each sector, in the year n+k, most efficient lag.

```{r}
all_merged$tax_category <- as.factor(all_merged$tax_category)

model_taxcat <- lm(GHG_Emissions_value ~ tax_category + GDP_value + sector + TIME_PERIOD,
                   data = all_merged)

summary(model_taxcat)
```

##### Comment :

## 2. Between taxes and social inequalities
### a. All taxes
```{r, echo=FALSE}
# taxes on inability to warm house (economic inequalities)
model_ineq <- lm(Warm_Households_value ~ Envt_Tax_Revenues_value + GII_value + coef_ineq_value + hhtyp + incgrp,
                 data = all_merged)

summary(model_ineq)

# taxes on social inequalities
model_ineq_tax <- lm(coef_ineq_value ~ Envt_Tax_Revenues_value + hhtyp + incgrp, data = all_merged)
summary(model_ineq_tax)

model_gii_tax <- lm(GII_value ~ Envt_Tax_Revenues_value + hhtyp + incgrp, data = all_merged)
summary(model_gii_tax)

```
##### Comment :

short comments (redo)

1. Heating difficulties Higher environmental taxes slightly reduce households’ inability to heat their homes (coefficient = -5.97×10⁻⁵, p \< 0.001). Gender inequality (GII) and overall inequality (coef_ineq) are the main drivers, with coefficients 71.2 and 3.02, respectively ??? (both p \< 0.001). Household type and income group also matter. Model explains 59% of variation (adj. R² = 0.590).

2.  Effect on income inequality (??warming ??) Higher environmental taxes are associated with a tiny increase in income inequality (coefficient = 1.705×10⁻⁵, p \< 0.001). The model explains less than 1% of inequality variation (R² = 0.0095), so the economic effect is very small despite statistical significance.

3.  Effect on gender inequality (GII) Higher environmental taxes slightly reduce gender inequality (coefficient = -1.115×10⁻⁶, p \< 0.001)**KUA?il faut qu'on explique pk !**!. The model explains only 4% of GII variation (R² = 0.042), indicating a statistically clear but economically minor effect.

#### Graphical representation

```{r}
ineq_plot_data <- all_merged %>%
  select(
    Envt_Tax_Revenues_value,
    Warm_Households_value,
    coef_ineq_value,
    GII_value
  ) %>%
  pivot_longer(
    cols = c(Warm_Households_value, coef_ineq_value, GII_value),
    names_to = "inequality_indicator",
    values_to = "ineq_value"
  ) %>%
    filter(
    !is.na(Envt_Tax_Revenues_value),
    !is.na(ineq_value),
    is.finite(Envt_Tax_Revenues_value),
    is.finite(ineq_value)
  )

ggplot(ineq_plot_data,
       aes(x = Envt_Tax_Revenues_value, y = ineq_value)) +
  geom_point(alpha = 0.1, size = 1, color = "darkgreen") +
  geom_smooth(method = "lm", se = TRUE, color = "darkred") +
  facet_wrap(~inequality_indicator, scales = "free_y") +
  labs(
    title = "Relationship Between Environmental Tax Revenues and Social Inequality Indicators",
    x = "Environmental Tax Revenues",
    y = "Value of Social Inequality Indicator"
  ) +
  theme_minimal()
```



## 3. Between tax and economy
->what is the impact of environmental taxes on the economy ? 

### 3.1. Between taxes and GDP
#### a. All taxes
```{r, echo=FALSE}
all_tax_data <- all_merged %>%
  filter(tax_category == "Any tax base")
model_gdp_tax <- lm(GDP_value ~ Envt_Tax_Revenues_value, data = all_tax_data)
summary(model_gdp_tax)
```

##### Comment :
The regression restricted to the “Any tax base” category reveals a very strong and highly significant relationship between GDP and environmental tax revenues. The coefficient on environmental tax revenues is positive and statistically significant at the 1% level (p < 0.001), with an exceptionally large t-value (406), indicating that the estimate is extremely precise. This suggests that within this broad and non-specific tax category, higher environmental tax revenues are strongly associated with higher GDP levels.

The intercept is not statistically significant, meaning that the model does not provide strong evidence about the expected GDP when environmental tax revenues are near zero. However, this has little substantive impact, since the main explanatory power comes from the revenue variable.

The model exhibits an exceptionally high explanatory power: the adjusted R² is 0.93, meaning that roughly 93% of the variation in GDP within this subset is explained solely by environmental tax revenues. This is an unusually high value for macroeconomic data and indicates a very tight empirical relationship between the two variables in this specific tax category.

The overall model significance is confirmed by the extremely large F-statistic (p < 0.001), showing that environmental tax revenues provide strong explanatory power for GDP in this subset. While the relationship is purely correlational, these results suggest that in contexts where environmental taxes are aggregated under a broad “any tax base” classification, higher environmental tax revenues tend to coincide with substantially higher levels of economic activity.

```{r}
clean_tax_data <- all_tax_data %>%
  filter(
    !is.na(GDP_value),
    !is.na(Envt_Tax_Revenues_value),
    is.finite(GDP_value),
    is.finite(Envt_Tax_Revenues_value)
  )

ggplot(clean_tax_data, aes(x = Envt_Tax_Revenues_value, y = GDP_value)) +
  geom_point(alpha = 0.1, size = 1.2, color = "darkgreen") +
  geom_smooth(method = "lm", se = TRUE, color = "darkred") +
  labs(
    title = "Relationship between Environmental Tax Revenues and GDP",
    x = "Environmental Tax Revenues",
    y = "GDP"
  ) +
  theme_minimal()
```

#### b. Per tax category
```{r, echo=FALSE}
all_merged$tax_category <- relevel(all_merged$tax_category, 
                                   ref = "Tax base related to energy taxes")
model_emissions_taxval <- lm(GDP_value ~ tax_category + Envt_Tax_Revenues_value,
                             data = all_merged)
summary(model_emissions_taxval)
```

##### Comment :
The regression results indicate a strong and highly significant relationship between GDP, environmental tax revenues, and tax categories. All estimated coefficients are statistically significant at the 1% level (p < 0.001), meaning that the variables included in the model reliably explain differences in GDP. The overall model is also highly significant (F-statistic p < 0.001), confirming that the set of predictors jointly contributes to explaining economic performance.

Compared to the reference category (“Tax base related to energy taxes”), the categories related to transport, pollution, and resource taxes are associated with higher levels of GDP. This suggests that countries (or years) where environmental taxation relies more heavily on these bases tend to exhibit stronger economic activity. Conversely, the negative coefficient for the “Any tax base” category may reflect contexts where environmental taxation is less targeted or less structured, potentially being associated with weaker economic outcomes.

Environmental tax revenues themselves have a positive and highly significant effect: higher receipts from environmental taxes are associated with higher GDP levels. While this correlation does not imply causality, it may indicate that stronger economies can raise more environmental tax revenue, or that well-designed environmental taxes can coexist with — or even support — economic performance.

The adjusted R² of about 0.47 indicates that nearly half of the variation in GDP is explained by the included variables, which is relatively high for macroeconomic cross-country or panel data. Overall, the results suggest that both the structure and the magnitude of environmental taxation are meaningfully related to economic activity.

### 3.2. Between taxes and added value
#### Regression
```{r, echo=FALSE}
sector <- c("Transport", "Agriculture", "Industry (including energy)", "Manufacturing")

model_AV_tax <- lapply(sector, function(s){
  data_s <- all_merged %>% filter(sector == s)
  lm(AV_value ~ Envt_Tax_Revenues_value, data = data_s)
})

names(model_AV_tax) <- sector

lapply(model_AV_tax, summary)
```

##### Comment :
Transport sector

The results show a strong and statistically significant positive relationship between environmental tax revenues and added value in the transport sector. The coefficient on environmental tax revenues is positive and highly significant (p < 0.001), indicating that higher environmental tax revenues are associated with higher sectoral added value. The model explains about 39% of the variation in added value, which is relatively substantial for sector-level data. Although the relationship is correlational, this result suggests that periods or countries with higher environmental tax revenues tend to coincide with stronger economic activity in the transport sector.

Agriculture sector

In the agricultural sector, environmental tax revenues are also positively and significantly related to added value. The coefficient is smaller in magnitude compared to other sectors, reflecting the smaller economic scale of agriculture, but remains highly significant (p < 0.001). The model explains about 32% of the variation in added value. This indicates a moderately strong association, suggesting that increases in environmental tax revenues align with higher value creation in agriculture, although the sector is structurally less responsive than transport or industry.

Industry (including energy)

The regression for industry (including energy) reveals a strong and highly significant positive association between environmental tax revenues and added value. The coefficient is the highest among all sectors, indicating that added value in industry is particularly sensitive to variations in environmental tax revenue. The model explains around 35% of the added value variability, highlighting a robust correlation. These results may reflect the capital-intensive and heavily regulated nature of industry and energy production, where environmental taxation and economic output often evolve together.

Manufacturing sector

The manufacturing sector also displays a positive and statistically significant effect of environmental tax revenues on added value (p < 0.001). The estimated coefficient is close to that of the transport sector, indicating a comparably strong association. The model explains about 34% of the variation in added value, suggesting a substantial link between fiscal environmental instruments and economic performance in manufacturing. As in other sectors, these results reflect correlation rather than causality, but they indicate that higher environmental tax revenues tend to coincide with higher manufacturing activity.



Global comparative summary

Across all four sectors—transport, agriculture, industry (including energy), and manufacturing—the relationship between environmental tax revenues and added value is consistently positive, highly significant, and economically meaningful. Although the strength of the effect varies across sectors, the overall pattern suggests that higher environmental tax revenues tend to coincide with higher levels of sectoral economic activity.

The transport sector exhibits one of the strongest relationships, with a sizeable coefficient and an R² of about 39%, indicating that environmental tax revenues explain a notable share of the variation in its added value. Manufacturing and industry (including energy) also show substantial effects, with R² values around 34–35%, reflecting the structural sensitivity of these capital-intensive sectors to regulatory and fiscal environments. In contrast, the agricultural sector displays a smaller coefficient and a more moderate explanatory power (R² ≈ 32%), which likely reflects its smaller economic scale and lower exposure to environmental taxation mechanisms.

Overall, while the results do not imply causality, they reveal a consistent pattern: sectors with higher environmental tax revenues tend to report higher added value. This suggests that environmental taxation—at least in aggregate—does not appear to hinder sectoral economic performance and may even evolve alongside periods of stronger economic activity.

#### Graphical representation
```{r}
ggplot(all_merged %>% drop_na(Envt_Tax_Revenues_value, AV_value),
       aes(x = Envt_Tax_Revenues_value, y = AV_value)) +
  geom_point(alpha = 0.1, color = "darkgreen") +
  geom_smooth(method = "lm", se = TRUE, color = "darkred") +
  facet_wrap(~ sector, scales = "free") +
  labs(
    title = "Relationship between Environmental Tax Revenues and Added Value by Sector",
    x = "Environmental Tax Revenues",
    y = "Added Value"
  ) +
  theme_minimal()
```
Agriculture

The relationship is positive but weaker than in other sectors, reflecting the smaller scale and lower exposure of agriculture to environmental taxation.

Industry (including energy)

Shows a strong positive association; this sector reacts the most, likely due to its capital intensity and direct exposure to environmental regulation.

Manufacturing

Displays a clear positive relationship, similar in strength to transport, indicating that higher environmental tax revenues coincide with higher manufacturing activity.

Transport

Strong linear relationship with added value; environmental tax revenue seems closely aligned with economic output in this sector.

Overall

Across all sectors, environmental tax revenues are positively and significantly related to added value. While the strength varies, the consistent upward trend suggests that environmental taxation is not associated with weaker sectoral performance; rather, periods with higher environmental tax revenues tend to coincide with stronger economic activity.

->which economic sector is the most economically efficient & produces the less emissions (à favoriser, éventuellement subsidies -> info à garder pour la conclusion) 
->which is the less efficient & emits the most ? ->target the tax on this sector ! 

# Forecasting : if we continue as we do

We use this forecast to try to predict the impact of environmental taxes on different variables.

### 1. Greenhouse Gas emissions


```{r, echo=FALSE}
model_ghg_tax <- lm(GHG_Emissions_value ~ Envt_Tax_Revenues_value, data = all_merged)
future <- data.frame(Envt_Tax_Revenues_value = seq(min(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
                                      max(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
                                      length.out = 25))
predicted_GHG <- predict(model_ghg_tax, newdata = future)
predicted_GHG
```
#### a. For all taxes 
```{r, echo=FALSE}

```

##### Comment :

#### b. Per tax
```{r, echo=FALSE}

```

##### Comment :

### 2. social inequalities: 


```{r, echo=FALSE}

model_gii_tax <- lm(GII_value ~ Envt_Tax_Revenues_value, data = all_merged)
predicted_GII <- predict(model_gii_tax, newdata = data.frame(Envt_Tax_Revenues_value = seq(
  min(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
  max(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
  length.out = 25)))
predicted_GII

## RAJOUTER COEF INEQ

model_warm_tax <- lm(Warm_Households_value ~ Envt_Tax_Revenues_value, data = all_merged)
predicted_Warm <- predict(model_warm_tax, newdata = data.frame(Envt_Tax_Revenues_value = seq(
  min(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
  max(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
  length.out = 25)))
predicted_Warm
```

##### Comment :



### 3. The economy : forecast of the GDP
```{r, echo=FALSE}

```

##### Comment :


# Forecasting: if we plan to get to the 0 net emission by 2050

We propose a tax of ... to reach the 0 net emission by 2050. blabla comment on l'a trouvée...

### 1. Levels greenhouse gas emissions reached with our proposed tax
->check for all taxes & per taxes (en regardant surtout la taxe la + efficace sur les GHG, déterminée dans la section régressions linéaires), quel montant ? 
```{r, echo=FALSE}
model_ghg_tax <- lm(GHG_Emissions_value ~ Envt_Tax_Revenues_value, data = all_merged)
future <- data.frame(Envt_Tax_Revenues_value = seq(min(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
                                      max(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
                                      length.out = 25))
predicted_GHG <- predict(model_ghg_tax, newdata = future)
predicted_GHG
```
##### Comment :


### 2. Social inequalities

```{r, echo=FALSE}

model_gii_tax <- lm(GII_value ~ Envt_Tax_Revenues_value, data = all_merged)
predicted_GII <- predict(model_gii_tax, newdata = data.frame(Envt_Tax_Revenues_value = seq(
  min(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
  max(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
  length.out = 25)))
predicted_GII

## Rajouter coef ineq

model_warm_tax <- lm(Warm_Households_value ~ Envt_Tax_Revenues_value, data = all_merged)
predicted_Warm <- predict(model_warm_tax, newdata = data.frame(Envt_Tax_Revenues_value = seq(
  min(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
  max(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
  length.out = 25)))
predicted_Warm
```
##### Comment :


### 3. Is the model viable ? Its impact on the GDP 
```{r, echo=FALSE}

```
##### Comment :

 
# Conclusion 

## Ouverture
Si une telle les 3 types de taxes ne sont pas compatibles / impact trop grand sur les inégalités et/ou sur l'économie / impact trop faible sur les GHG, quelles alternatives (en plus de la taxe / à la place de la taxe) ? 
 - subvention : à la recherche ? 
 - changement modes de consommation, institutions : sobriété éco ? décroissance
 - autre ? 

