---
title: "dm-project"
author: "Daphné Perry and Justine Catta"
format: html
---

# Summary 

# Introduction

## Description of our research question

Our research question is the following: “To what extent do environmental taxes reduce Greenhouse Gas Emissions (GHG) in EU countries, and what are their long-term socioeconomic impacts across economic sectors?”. As the European Union commits to achieving net-zero emissions by 2050 under the European Green Deal, we aim to better understand the effectiveness and consequences of environmental taxation, both on emissions and socially.

To answer this question, we will use sector-level panel data for all EU member states. Our data set includes GHG emissions, sectoral GDP (measured as Added Value), environmental tax revenues, and several indicators of social well-being. These include the Gender Inequality Index (GII), a coefficient of human inequality, and a measure of energy poverty that is the inability to adequately heat one’s home. Together, these variables allow us to evaluate the distributive consequences of taxes.

After cleaning, harmonizing, and describing the data sets, we aim to estimate several relationships. First, we will measure the link between environmental taxation and sectoral GHG emissions. Do higher taxes lead to statistically significant reductions in emissions at both the EU and country levels? Are some sectors more responsive to taxation than others? To maintain analytical clarity and comparability across states, we will focus on four key sectors: Agriculture, Energy, Manufacturing, and Transport, which together account for the bulk of emissions and policy intervention.

We will then investigate the association between environmental taxes and social inequality indicators. Our core question here can be summarized as: “Is there a social cost associated with environmental taxation, and if so, how large is it?” Specifically, we will analyse whether environmental taxes correlate with increases or decreases in gender inequality, income inequality, and energy poverty. This step is essential because policies that are environmentally effective but socially regressive may have a reduced long-term viability.

Building on these empirical relationships, we will construct forecasts for the period 2026–2050 to estimate the level of taxation required to meet the European Green Deal’s emissions targets. We consider a policy scenario relying solely on environmental taxation applied to relevant sectors. Our analysis will assess whether such an approach is socially and economically sustainable: How would different tax levels affect inequalities and energy poverty? How much revenue would such taxes generate, and would the required tax burden be compatible with each country’s GDP levels?

By combining environmental effectiveness, economic feasibility, and social impact, we can then address our overarching question: Is it socially and economically optimal for the EU to rely exclusively on environmental taxation to achieve net-zero emissions by 2050? If taxation alone proves insufficient or too socially costly, our results may highlight the need for complementary policies such as subsidies for green technologies, targeted social transfers, or sector-specific regulations to ensure a fair and effective ecological transition.


## link to the github project;

Our project is available on github [here](https://github.com/cattajustine-spec/project).

### Setup

```{r, echo=FALSE}
#| message: false
library(vroom)
library(ggplot2)
library(knitr)
library(dplyr)
library(here)
library(countrycode)
library(plotly)
library(tidyr)
library(knitr)
```

```{r, echo=FALSE}
#| message: false
here::i_am("project.Rproj")
```


## Presenting our data :

### 1. Environmental tax revenues (OECD, 1994–2023) 

This data set reports annual environmentally related tax revenues for EU countries, broken into categorises based on their environmental relevance by tax-base category (including energy, transport, pollution, and resources) that are mutually exclusive. According to the OECD, the database contains detailed qualitative and quantitative information on environmentally related taxes and fees, tradable permits and offsets, deposit-refund schemes, environmentally beneficial subsidies and payments and voluntary approaches used for environmental policy.
Values are expressed in millions of USD. It is from the OECD and is available on this [link](https://data-explorer.oecd.org/vis?lc=en&tm=DF_ERTR&pg=0&snb=1&vw=tb&df%5Bds%5D=dsDisseminateFinalDMZ&df%5Bid%5D=DSD_ERTR%40DF_ERTR&df%5Bag%5D=OECD.ENV.EPI&df%5Bvs%5D=&pd=1994%2C2023&dq=A.HRV%2BLVA%2BLTU%2BLUX%2BMLT%2BNLD%2BCYP%2BAUT%2BBEL%2BCZE%2BDNK%2BEST%2BFIN%2BFRA%2BDEU%2BGRC%2BHUN%2BIRL%2BITA%2BPOL%2BPRT%2BSVK%2BSVN%2BESP%2BSWE%2BBGR%2BROU.TAXREV.CAT_ENE%2BCAT_TRA%2BCAT_POL%2BCAT_RES%2B_T._T.USD.&ly%5Brw%5D=REF_AREA&ly%5Bcl%5D=TIME_PERIOD&to%5BTIME_PERIOD%5D=false). They cover a time period from 1994 to 2023.


```{r, echo=FALSE}
#| message: false
Envt_Tax_Revenues <- vroom(here("data", "OECD_envt_tax.csv"))
```

::: {columns}
|                    |                                           |
|--------------------|-------------------------------------------|
| Name               | `r "Envt_Tax_Revenues"`                   |
| Number of rows     | `r nrow(Envt_Tax_Revenues)`               |
| Number of columns  | `r ncol(Envt_Tax_Revenues)`               |
:::


### 2. GDP and Added Value (OECD, 1990-2024)

This data set provides annual GDP and sectoral gross Value Added figures for EU countries. It follows an output approach : GDP is measured as the sum of gross value added. For the Value Added by sectors, we will keep these ones : "Agriculture,forestry and fishing", "Manufacturing", "Industry (except construction)" and "Wholesale and retail trade; repair of motor vehicles and motorcycles; transportation and storage; accommodation and food service activities". (cf plus tard) All variables are measured in millions of USD, using exchange rate. 
 The data set on GDP and Value Added is available at this [link](https://data-explorer.oecd.org/vis?df%5Bds%5D=dsDisseminateFinalDMZ&df%5Bid%5D=DSD_NAMAIN10%40DF_TABLE1_OUTPUT&df%5Bag%5D=OECD.SDD.NAD&df%5Bvs%5D=&dq=A.BEL%2BAUT%2BCZE%2BDNK%2BEST%2BFIN%2BFRA%2BDEU%2BGRC%2BHUN%2BIRL%2BITA%2BLVA%2BLTU%2BLUX%2BNLD%2BPOL%2BPRT%2BSVK%2BSVN%2BESP%2BSWE%2BBGR%2BHRV%2BCYP%2BMLT%2BROU...B1G%2BB1GQ.._Z%2B_T%2BC%2BA%2BBTE%2BF%2BGTI%2BJ%2BK%2BL%2BM_N%2BOTQ%2BRTU..USD_EXC.V..&to%5BTIME_PERIOD%5D=false&pd=1990%2C2024&vw=tb). 
The data covers the time period 1990 to 2024.


```{r, echo=FALSE}
#| message: false
GDP_and_AV <- vroom(here("data", "OECD_GDP_and_AV.csv"))
```

::: {columns}
|                    |                       |
|--------------------|-----------------------|
| Name               | `r "GDP_and_AV"`      |
| Number of rows     | `r nrow(GDP_and_AV)`  |
| Number of columns  | `r ncol(GDP_and_AV)`  |
:::

### 3. Greenhouse gas emissions by source sector (Eurostat, 1990-2023)

This data set reports total greenhouse gas emissions by source sector for the 27 eu countries. Values are expressed in thousand tonnes of CO₂-equivalent. The GHG source sectors we chose are the following ones, to be able to match the Added Value sectors (energy, transport, agriculture, industry): Total, Fuel combustion in energy industries, Fuel combustion in manufacturing industries and construction, Industrial processes and product use, agriculture,  fuel combustion in transport
The data set on Greenhouse gas emissions is [here](https://ec.europa.eu/eurostat/databrowser/view/env_air_gge__custom_18942108/default/table).
The data covers the time period 1990 to 2023.
```{r, echo=FALSE}
#| message: false
GHG_Emissions <- vroom(here("data", "Eurostat_ghg.csv"))
```

::: {columns}
|                    |                          |
|--------------------|--------------------------|
| Name               | `r "GHG_Emissions"`      |
| Number of rows     | `r nrow(GHG_Emissions)`  |
| Number of columns  | `r ncol(GHG_Emissions)`  |
:::

### 4.  Inability to keep home adequately warm (Eurostat, 2003-2024)

This data set measures the percentage of households unable to keep their home adequately warm. It reflects energy poverty and distributional effects of energy pricing across income groups and household types. We chose 3 households type : the total, single female and single male to be able to have an idea of the gender issues. We also chose two income situations : the total (all incomes) and the group below 60% of the median (in situation of poverty), because the income inequality is not available in the Human Index.
The data set on Inability to keep home adequately warm is linked [here](https://ec.europa.eu/eurostat/databrowser/view/ILC_MDES01__custom_18945942/default/table). The time period covered is 2003-2024.

```{r, echo=FALSE}
#| message: false
Warm_Households <- vroom(here("data","Eurostat_warm.csv"))
```

::: {columns}
|                    |                            |
|--------------------|----------------------------|
| Name               | `r "Warm_Households"`      |
| Number of rows     | `r nrow(Warm_Households)`  |
| Number of columns  | `r ncol(Warm_Households)`  |
:::

### 5.  Human index (UNDP HDR, 1990-2023)

This data set provides annual values for lots of Human Development Indexes. We focus on the Coefficient of Human Inequality  and Gender Inequality Index (GII) for most countries worldwide. It combines indicators related to health, education, income, and gender disparities, enabling long-run social development comparisons.

Available on this [page](HDR%20index%20https://hdr.undp.org/data-center/documentation-and-downloads) from the UNDP Human Development Report (HDR) "All composite indices and components time series (1990-2023)" or directly with this [link](https://hdr.undp.org/sites/default/files/2025_HDR/HDR25_Composite_indices_complete_time_series.csv). 
The time period of this data set is 1990 to 2023. 

```{r, echo=FALSE}
#| message: false
Soc_Indexes <- vroom(here("data", "HDR_indexes.csv"))
```


::: {columns}
|                    |                        |
|--------------------|------------------------|
| Name               | `r "Soc_Indexes"`      |
| Number of rows     | `r nrow(Soc_Indexes)`  |
| Number of columns  | `r ncol(Soc_Indexes)`  |
:::

### Country keys

List of countries : Austria, Belgium, Bulgaria, Croatia, Cyprus, Czechia, Denmark, Estonia, Finland, France, Germany, Greece, Hungary, Ireland, Italy, Latvia, Lithuania, Luxembourg, Malta, Netherlands, Poland, Portugal, Romania, Slovakia, Slovenia, Spain, Sweden

```{r, echo=FALSE}
eu_countries <- unique(GHG_Emissions$geo)

eu_country_key <- data.frame(name_en = eu_countries) |>
  mutate(
    iso3 = countrycode(name_en, origin = "country.name", destination = "iso3c"),
    iso2 = countrycode(name_en, origin = "country.name", destination = "iso2c"),
    name_fr = countrycode(name_en, origin = "country.name", destination = "country.name.fr")
  ) 
```

### Modifying the data sets

#### Removing unwanted variables

```{r}
Warm_Households <- Warm_Households |>
  select(geo, TIME_PERIOD, hhtyp, incgrp, OBS_VALUE) |>
  filter(geo %in% eu_country_key$iso2, 
         hhtyp %in% c("A1F", "A1M", "TOTAL"))
```


#### Removing unwanted variables
Because the data sets we chose are quite huge, we first need to remove all the variables we will not use, then add missing variables that we think are important for our analysis, and finally modify the variables we kept.

##### 1. Environmental tax revenues (in million of USD)

In this data set, we are only interested in the following variables:

-   **REF_AREA**: Identifies the country or reporting area using an ISO-style three-letter code (e.g., LVA for Latvia, LTU for Lithuania). This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 1994 to 2023).
-   **OBS_VALUE**: Contains the numerical value of environmental tax revenues. When the value is NA, it means the revenue amount is not available for that country and year.
-   **Catégorie de recettes provenant des taxes liées à l'environnement**: A label in French describing the tax revenue category in plain language. In this data set, there are five categories:
    -   "Assiette relatives aux taxes sur l'énergie", i.e., energy-related taxes.
    -   "Assiette relatives aux taxes sur le transport", i.e., transport-related taxes.
    -   "Assiette relatives aux taxes sur les pollutions", i.e., pollution-related taxes.
    -   "Assiette relatives aux taxes sur les resources", i.e., resource-related taxes.
    -   “Toute assiette de taxes”, i.e., all environmental tax bases combined.

```{r, echo=FALSE}
Envt_Tax_Revenues <- Envt_Tax_Revenues |>
  select(REF_AREA, TIME_PERIOD, OBS_VALUE, 
         `Catégorie de recettes provenant des taxes liées à l'environnement`) |>
  filter(REF_AREA %in% eu_country_key$iso3)
```

##### 2. GDP and Value Added per capita annual growth per sector (in million of USD)

In this data set, we have two main variables of interest: the [Gross Domestic Product](#grossdomesticproduct) and the [Gross Added Value](#grossaddedvalue).

They are described by the **Transaction** variable, therefore, we need to create two independent data sets, one for each variable of interest.

###### Gross Domestic product {#grossdomesticproduct}

In this data set, we are only interested in the following variables:

-   **REF_AREA**: Identifies the country or reporting area using an ISO-style three-letter code (e.g., LVA for Latvia, LTU for Lithuania). This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 1990 to 2024).
-   **OBS_VALUE**: Contains the numerical value of the GDP.

```{r, echo=FALSE}
GDP <- GDP_and_AV |>
  filter(Transaction == "Produit intérieur brut", 
         REF_AREA %in% eu_country_key$iso3) |>
  select(REF_AREA, TIME_PERIOD, OBS_VALUE)
```

###### Gross Added Value {#grossaddedvalue}

In this data set, we are only interested in the following variables:

-   **REF_AREA**: Identifies the country or reporting area using an ISO-style three-letter code (e.g., LVA for Latvia, LTU for Lithuania). This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 1990 to 2024).
-   **Activité économique**: A label in French describing the economic activity or industry sector associated with the added value produced. In this data set, we are only interested in four sectors (cf. [Justification for the Sector Mapping Between ISIC and CRF Classifications](#justificationforsectormapping)):
    -   "Activités de fabrication", i.e., manufacturing.
    -   "Agriculture, sylviculture et pêche", i.e., agriculture.
    -   "Industrie (sauf construction)", i.e., public electricity and heat production.
    -   "Commerce de gros et de détail; réparation de véhicules automobiles et de motocyclettes; transport et stockage; activités d'hébergement et de restauration", i.e., transport.\
-   **OBS_VALUE**: Contains the numerical value of the gross added value.

```{r, echo=FALSE}
Added_value <- GDP_and_AV |>
  filter(Transaction == "Valeur ajoutée, brute", 
         REF_AREA %in% eu_country_key$iso3, 
         `Activité économique` %in% c(
           "Agriculture, sylviculture et pêche",
           "Activités de fabrication",
           "Industrie (sauf construction)",
           "Commerce de gros et de détail; réparation de véhicules automobiles et de motocyclettes; transport et stockage; activités d'hébergement et de restauration"
           )) |>
  select(REF_AREA, `Activité économique`, 
         TIME_PERIOD, OBS_VALUE)
```

##### 3. Greenhouse gas emissions by source sector (in million of ton)

In this data set, we are only interested in the following variables:

-   **geo**: Identifies the country or reporting area. This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 1990 to 2024).
-   **src_crf**: Describes the economic activity or industry sector associated with the greenhouse gases emitted. In this data set, we are only interested in five sectors (cf. [Justification for the Sector Mapping Between ISIC and CRF Classifications](#justificationforsectormapping)):
    -   "Agriculture".
    -   "Fuel combustion in energy industries", i.e., public electricity and heat production.
    -   "Fuel combustion in manufacturing industries and construction" and “Industrial processes and product use”, i.e., manufacturing (we will combine this sectors in the [Renaming some variables](#renamingsomevariables) section).
    -   "Fuel combustion in transport", i.e., transport.
-   **OBS_VALUE**: Contains the numerical value of the greenhouse gases emitted.

```{r, echo=FALSE}
GHG_Emissions <- GHG_Emissions |>
  select(airpol, src_crf, geo, TIME_PERIOD, OBS_VALUE) |>
  filter(geo %in% eu_country_key$name_en, 
         src_crf != "Total (excluding memo items)")
```

###### Justification for the Sector Mapping Between ISIC and CRF Classifications {#justificationforsectormapping}

The reconciliation of sectors across the ISIC-based data set and the CRF (Common Reporting Format) classification relies on aligning economic activities with their dominant emission sources. Although the two systems are not structurally identical—ISIC being an economic activity classification and CRF being an emissions reporting framework—several robust correspondences can be established.

First, Agriculture, forestry and fishing (ISIC A) is mapped to CRF 3 (Agriculture). Both refer to primary-sector biological production and include emissions directly linked to agricultural practices such as enteric fermentation, manure management, and agricultural soils.

Second, Manufacturing (ISIC C) is associated with CRF 1.A.2 (Manufacturing industries and construction) and CRF 2 (Industrial processes and product use). Manufacturing activity generates emissions both from fuel combustion in industrial facilities (covered by 1.A.2) and from process-related sources such as cement, chemicals, or metals production (covered by CRF 2), making this linkage consistent with the structure of industrial emissions.

Third, Electricity, gas, steam and air conditioning supply (ISIC D)—even though aggregated within “Industry (except construction)” in the first data set—logically corresponds to CRF 1.A.1 (Public electricity and heat production). This CRF category captures emissions from power generation and heat production, which are the core activities of ISIC D.

Finally, Transportation and storage (ISIC H)—included in a broader aggregated category in the data set—is aligned with CRF 1.A.3 (Transport). This CRF category encompasses fuel combustion from road, rail, aviation, and navigation transport, which directly reflects the operational scope of ISIC H.

Overall, the mapping follows a principle of dominant emission source equivalence, ensuring that each ISIC activity is linked to the CRF category representing the primary origin of its greenhouse gas emissions. This approach allows the two data sets—despite structural differences—to be compared and integrated in a coherent and meaningful way for our analysis.

##### 4. Inability to keep home adequately warm

In this data set, we are only interested in the following variables:

-   **geo**: Identifies the country or reporting area. This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 2003 to 2024).
-   **hhtyp**: Indicates the composition of the household, here we are only interested in:
    -   "A1F", i.e., the households composed by only one female.
    -   "A1M", i.e., the households composed by only one male.
    -   "TOTAL", i.e., all households (whatever their composition).
-   **incrp**: Indicates the income group of the household, here we are only interested in:
    -   "B_MD60", i.e., the households who have an income below 60% of median equivalised income (meaning they are below the poverty treshold).
    -   "TOTAL", i.e., all households (whatever their revenue).
-   **OBS_VALUE**: Contains the percentage of household that are not able to keep their home adequately warm.

```{r, echo=FALSE}
Warm_Households <- Warm_Households |>
  select(geo, TIME_PERIOD, hhtyp, incgrp, OBS_VALUE) |>
  filter(geo %in% eu_country_key$iso2, 
         hhtyp %in% c("A1F", "A1M", "TOTAL"))
```

##### 5. HDR index

In this data set, we have two main variables of interest: the [Gender Inequality Index](#genderinequalityindex) and the [Coefficient of Human Inequality](#coefficientofhumaninequality).

They are described per year as different variables (meaning we have a gii_2020, gii_2021, ..., coef_ineq_2003, coef_ineq_2004, etc.). Therefore, we need to pivot our database, and then create two independent data sets, one for each variable of interest.

```{r, echo=FALSE}
Soc_Indexes <- Soc_Indexes |>
  pivot_longer(
    cols = matches("_[0-9]{4}$"),   
    names_to = c("Indicator", "TIME_PERIOD"),
    names_pattern = "^(.*)_(\\d{4})$",
    values_to = "OBS_VALUE"
  ) |>
  mutate(
    OBS_VALUE = as.numeric(OBS_VALUE),
    TIME_PERIOD = as.integer(TIME_PERIOD)
  )
```

###### Gender inequality Index {#genderinequalityindex}

In this data set, we are only interested in the following variables:

-   **iso3**: Identifies the country or reporting area using an ISO-style three-letter code (e.g., LVA for Latvia, LTU for Lithuania). This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 1990 to 2023).
-   **OBS_VALUE**: Contains the numerical value of the GII.

```{r, echo=FALSE}
GII <- Soc_Indexes|>
  filter(iso3 %in% eu_country_key$iso3,
         Indicator == "gii") |>
  select(iso3, TIME_PERIOD, OBS_VALUE)
```

###### Coefficient of Human inequality {#coefficientofhumaninequality}

In this data set, we are only interested in the following variables:

-   **iso3**: Identifies the country or reporting area using an ISO-style three-letter code (e.g., LVA for Latvia, LTU for Lithuania). This variable is the geographic dimension of the data set.
-   **TIME_PERIOD**: Indicates the reference year of observation (from 2010 to 2023).
-   **OBS_VALUE**: Contains the numerical value of the coefficient of human inequality.

```{r, echo=FALSE}
coef_ineq <- Soc_Indexes|>
  filter(iso3 %in% eu_country_key$iso3,
         Indicator == "coef_ineq")|>
  select(iso3, TIME_PERIOD, OBS_VALUE)
```

#### Adding some entities

##### Add European average (across countries)
For the data sets that are not described per economic sector, we compute the European average across countries:


```{r, echo=FALSE}
#| message: false
EU27_GDP <- GDP |>                #computing the average European GDP 
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(REF_AREA = "EU27") |>
  relocate(REF_AREA)
GDP <- bind_rows(GDP, EU27_GDP)

EU27_Tax <- Envt_Tax_Revenues |>   #computing the average European environmental tax revenue, depending on the tax category. 
  group_by(TIME_PERIOD,
           `Catégorie de recettes provenant des taxes liées à l'environnement`) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(REF_AREA = "EU27") |>
  relocate(REF_AREA)
Envt_Tax_Revenues <- bind_rows(Envt_Tax_Revenues, EU27_Tax)

EU27_Warming <- Warm_Households |>  #computing the average European inability to keep home adequately warm, depending on the household composition and income.  
  group_by(TIME_PERIOD, hhtyp, incgrp) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(geo = "EU27") |>
  relocate(geo)
Warm_Households <- bind_rows(Warm_Households, EU27_Warming)

EU27_GII <- GII |>  #computing the average European gender inequality index. 
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(iso3 = "EU27") |>
  relocate(iso3)
GII <- bind_rows(GII, EU27_GII)

EU27_Coefineq <- coef_ineq |> #computing the average European coefficient of human inequality. 
  group_by(TIME_PERIOD) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(iso3 = "EU27") |>
  relocate(iso3)
coef_ineq <- bind_rows(coef_ineq, EU27_Coefineq)
```


##### Add country average (accross sectors)
For the data sets that are described per economic sector, we have to compute the European average across sectors:

```{r, echo=FALSE}
#| message: false
EU27_AV <- Added_value |> #computing the average European gross added value, depending on the economic sector.  
  group_by(TIME_PERIOD, `Activité économique`) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(REF_AREA = "EU27") |>
  relocate(REF_AREA)
Added_value <- bind_rows(Added_value, EU27_AV)

EU27_GHG <- GHG_Emissions |> #computing the average European greenhouse gas emissions, depending on the economic sector.  
  group_by(TIME_PERIOD, src_crf) |>
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) |>
  mutate(geo = "EU average") |>
  relocate(geo)
GHG_Emissions <- bind_rows(GHG_Emissions, EU27_GHG)
```

#### Modifiying the characteristics of the variables
For the variables related to the economic sectors, it will be easier to consider them as *factor*, rather than as *character* variables.

```{r, echo=FALSE}
GHG_Emissions$src_crf <- factor(GHG_Emissions$src_crf)
Added_value$`Activité économique` <- factor(Added_value$`Activité économique`)
```

##### Renaming some variables {#renamingsomevariables}

To be able to merge our data sets in a data base, we need to modify the name of the variable **OBS_VALUE** that is in every data set to be able to identify which OBS_VALUE is associated to which data set.

```{r, echo=FALSE}
names(Warm_Households)[names(Warm_Households) == "OBS_VALUE"] <- "Warm_Households_value"
names(Envt_Tax_Revenues)[names(Envt_Tax_Revenues) == "OBS_VALUE"] <- "Envt_Tax_Revenues_value"
names(GDP)[names(GDP) == "OBS_VALUE"] <- "GDP_value"
names(Added_value)[names(Added_value) == "OBS_VALUE"] <- "AV_value"
names(GHG_Emissions)[names(GHG_Emissions) == "OBS_VALUE"] <- "GHG_Emissions_value"
names(GII)[names(GII) == "OBS_VALUE"] <- "GII_value"
names(coef_ineq)[names(coef_ineq) == "OBS_VALUE"] <- "coef_ineq_value"
```

To be able to merge the data sets including variables related to the economic sectors, it will be easier to have the same sector names, in english.

```{r, echo=FALSE}
Added_value <- Added_value |>
  mutate(`Activité économique` = recode(`Activité économique`,
                                        "Commerce de gros et de détail; réparation de véhicules automobiles et de motocyclettes; transport et stockage; activités d'hébergement et de restauration" = "Transport", 
                                        "Agriculture, sylviculture et pêche"= "Agriculture", 
                                        "Industrie (sauf construction)" = "Public electricity and heat production", 
                                        "Activités de fabrication" = "Manufacturing"
                                        ))

EU27_AV <- EU27_AV |>
  mutate(`Activité économique` = recode(`Activité économique`,
                                        "Commerce de gros et de détail; réparation de véhicules automobiles et de motocyclettes; transport et stockage; activités d'hébergement et de restauration" = "Transport", 
                                        "Agriculture, sylviculture et pêche"= "Agriculture", 
                                        "Industrie (sauf construction)" = "Public electricity and heat production", 
                                        "Activités de fabrication" = "Manufacturing"
                                        ))

GHG_Emissions <- GHG_Emissions|>  #here we are combining the sectors related to manufacturing. 
  mutate(
    src_crf = case_when(
      src_crf %in% c("Fuel combustion in manufacturing industries and construction", 
                     "Industrial processes and product use") ~ "Manufacturing",
      TRUE ~ src_crf
    )
  ) |>
  group_by(geo, TIME_PERIOD, src_crf) |>
  summarise(GHG_Emissions_value = sum(GHG_Emissions_value), .groups = "drop")

GHG_Emissions <- GHG_Emissions |>
  mutate(src_crf = recode(src_crf, 
                          "Fuel combustion in energy industries"="Public electricity and heat production", 
                          "Fuel combustion in transport"="Transport"))

EU27_GHG <- EU27_GHG|>
  mutate(
    src_crf = case_when(
      src_crf %in% c("Fuel combustion in manufacturing industries and construction", 
                     "Industrial processes and product use") ~ "Manufacturing",
      TRUE ~ src_crf
    )
  ) |>
  group_by(geo, TIME_PERIOD, src_crf) |>
  summarise(EU27_GHG_value = sum(OBS_VALUE), .groups = "drop")

EU27_GHG <- EU27_GHG |>
  mutate(src_crf = recode(src_crf, 
                          "Fuel combustion in energy industries"="Public electricity and heat production", 
                          "Fuel combustion in transport"="Transport"))
```

```{r}
Envt_Tax_Revenues <- Envt_Tax_Revenues |>
  rename(iso3 = REF_AREA, tax_categorie = `Catégorie de recettes provenant des taxes liées à l'environnement`) 

GDP <- GDP |> rename(iso3 = REF_AREA) 

Added_value <- Added_value |>
  rename(iso3 = REF_AREA,
    sector = `Activité économique`) 

GHG_Emissions <- GHG_Emissions |> rename(name_en = geo,
                                         sector = src_crf)

Warm_Households <- Warm_Households |> rename(iso2 = geo)
```


# Merging all data sets on country-year pairs
We do this big join to have in only one data set all these informations :
 - TaxRevenue (environmental tax revenues)
 - GII_value (gender inequality)
 - coef_ineq_value (income inequality index)
 - WarmValue (share unable to heat home)
 - hhtyp, incgrp (household type, income group)
 - src_crf (GHG emission sector)
 - GHG_Emissions_value
 - GDP_value
 - sector, AV_value (added value by economic sector)



```{r}
GHG_Emissions <- GHG_Emissions |>
  left_join(eu_country_key |> select(name_en, iso3), by = "name_en")

Warm_Households <- Warm_Households |>
  left_join(eu_country_key |> select(iso2, iso3), by = "iso2")

```


```{r}
all_merged <- Envt_Tax_Revenues |>
  left_join(Added_value |> distinct(iso3, TIME_PERIOD, sector), by = c("iso3", "TIME_PERIOD")) |>
  left_join(Added_value, by = c("iso3", "TIME_PERIOD", "sector")) |>
  left_join(GDP, by = c("iso3", "TIME_PERIOD")) |>
  left_join(GHG_Emissions, by = c("iso3", "TIME_PERIOD", "sector")) |>
  left_join(Warm_Households, by = c("iso3", "TIME_PERIOD")) |>
  left_join(GII, by = c("iso3", "TIME_PERIOD")) |>
  left_join(coef_ineq, by = c("iso3", "TIME_PERIOD")) |>
  select(name_en, TIME_PERIOD, tax_categorie, Envt_Tax_Revenues_value, sector, AV_value, GDP_value, GHG_Emissions_value, hhtyp, incgrp, Warm_Households_value, GII_value, coef_ineq_value)

```


# Descriptive analyzes

Now that we have cleaned our data sets, here are some descriptive analyzes of our main target variables:

## 1. Environmental tax revenues per tax category

### Graphical representation

```{r, echo=FALSE}
ggplot(EU27_Tax, aes(x = TIME_PERIOD, y = OBS_VALUE, color = `Catégorie de recettes provenant des taxes liées à l'environnement`, group = `Catégorie de recettes provenant des taxes liées à l'environnement`)) +
  geom_line(linewidth = 1) +
  labs(
    title = "EU average Environmental Tax Revenues by Category",
    x = "Year",
    y = "Revenue (million USD)",
    color = "Category"
  ) +
  theme_minimal()
```

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_EnvtTax <- EU27_Tax %>%
  group_by(`Catégorie de recettes provenant des taxes liées à l'environnement`) %>%
  summarise(
    Min = min(OBS_VALUE, na.rm = TRUE),
    Max = max(OBS_VALUE, na.rm = TRUE),
    Mean = mean(OBS_VALUE, na.rm = TRUE),
    Median = median(OBS_VALUE, na.rm = TRUE),
    SD = sd(OBS_VALUE, na.rm = TRUE)
  )

kable(summary_stats_EnvtTax, caption = "Descriptive statistics of EU average Environmental Tax Revenues by Category (all years combined)")
```

COMMENT

## 2. GDP per country

### Graphical representation

```{r, echo=FALSE}
p_GDP <- ggplot(GDP, aes(x = TIME_PERIOD, y = GDP_value, color = REF_AREA, group = REF_AREA)) +
  geom_line(linewidth = 1) +
  labs(
    title = "GDP per Country (EU27)",
    x = "Year",
    y = "GDP (million USD)",
    color = "Country"
  ) +
  theme_minimal()

ggplotly(p_GDP)
```

COMMENT

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_GDP <- GDP %>%
  group_by(REF_AREA) %>%                        
  summarise(
    Min = min(GDP_value, na.rm = TRUE),
    Max = max(GDP_value, na.rm = TRUE),
    Mean = mean(GDP_value, na.rm = TRUE),
    Median = median(GDP_value, na.rm = TRUE),
    SD = sd(GDP_value, na.rm = TRUE)
    )

kable(summary_stats_GDP, 
      caption = "Descriptive statistics of the GDP by country (all years combined)")
```

## 3. Added Value per economic sector

### Graphical representation
```{r, echo=FALSE}
p_va <- ggplot(EU27_AV, aes(x = TIME_PERIOD, y = OBS_VALUE, color = `Activité économique`, group = `Activité économique`)) +
  geom_line(linewidth = 1) +
  labs(
    title = "EU27 Average Added Value per Sector",
    x = "Year",
    y = "Value Added (million USD)",
    color = "Sector"
  ) +
  theme_minimal()

ggplotly(p_va)
```

COMMENT

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_AV <- EU27_AV %>%
  group_by(`Activité économique`) %>%
  summarise(
    Min = min(OBS_VALUE, na.rm = TRUE),
    Max = max(OBS_VALUE, na.rm = TRUE),
    Mean = mean(OBS_VALUE, na.rm = TRUE),
    Median = median(OBS_VALUE, na.rm = TRUE),
    SD = sd(OBS_VALUE, na.rm = TRUE)
  )

kable(summary_stats_AV, caption = "Descriptive statistics of EU average gross added value by sector (all years combined)")
```

## 4. Greenhouse gas emissions by economic sector

### Graphical representation

```{r, echo=FALSE}
p_GHG <- ggplot(EU27_GHG, aes(x = TIME_PERIOD, y = EU27_GHG_value, color = src_crf, group = src_crf)) +
  geom_line(linewidth = 1) +
  labs(
    title = "EU27 Average Greenhouse gas emissions per Sector",
    x = "Year",
    y = "Greenhouse gas emissions (million ton of CO2)",
    color = "Sector"
  ) +
  theme_minimal()

ggplotly(p_GHG)
```

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_GHG <- EU27_GHG %>%
  group_by(src_crf) %>%                
  summarise(
    Min = min(EU27_GHG_value, na.rm = TRUE),
    Max = max(EU27_GHG_value, na.rm = TRUE),
    Mean = mean(EU27_GHG_value, na.rm = TRUE),
    Median = median(EU27_GHG_value, na.rm = TRUE),
    SD = sd(EU27_GHG_value, na.rm = TRUE)
  )

kable(summary_stats_GHG, caption = "Descriptive statistics of EU average greenhouse gases emissions by sector (all years combined)")
```

## 5. Inability to keep home adequately warm by household type

### Graphical representation

```{r, echo=FALSE}
ggplot(EU27_Warming, 
       aes(x = TIME_PERIOD, y = OBS_VALUE,
           color = interaction(hhtyp, incgrp),
           group = interaction(hhtyp, incgrp))) +
  geom_line(linewidth = 1) +
  labs(
    title = "EU27 – inability to Keep the House Warm (by Sex × Income Group)",
    x = "Year",
    y = "Percentage of population",
    color = "Sex × Income"
  ) +
  theme_minimal()
```

COMMENT


### Descriptive statistics

```{r, echo=FALSE}
summary_stats_Warm <- EU27_Warming %>%
  group_by(hhtyp, incgrp) %>%                    
  summarise(
    Min = min(OBS_VALUE, na.rm = TRUE),
    Max = max(OBS_VALUE, na.rm = TRUE),
    Mean = mean(OBS_VALUE, na.rm = TRUE),
    Median = median(OBS_VALUE, na.rm = TRUE),
    SD = sd(OBS_VALUE, na.rm = TRUE), 
    .groups = "drop"
  )

kable(summary_stats_Warm, caption = "Descriptive statistics of the inability to keep house adequately warm, by household type (all years combined)")
```

## 6. GII per country

### Graphical representation

```{r, echo=FALSE}
p_GII <- ggplot(GII, aes(x = TIME_PERIOD, y = GII_value, color = iso3, group = iso3)) +
  geom_line(linewidth = 1) +
  labs(
    title = "GII per Country (EU27)",
    x = "Year",
    y = "Gender inequality index",
    color = "Country"
  ) +
  theme_minimal()

ggplotly(p_GII)
```

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_GII <- GII %>%
  group_by(iso3) %>%                        
  summarise(
    Min = min(GII_value, na.rm = TRUE),
    Max = max(GII_value, na.rm = TRUE),
    Mean = mean(GII_value, na.rm = TRUE),
    Median = median(GII_value, na.rm = TRUE),
    SD = sd(GII_value, na.rm = TRUE)
    )

kable(summary_stats_GII, 
      caption = "Descriptive statistics of the GII by country (all years combined)")
```

## 7. Coefficient of human inequality per country

### Graphical representation

```{r, echo=FALSE}
p_coef <- ggplot(coef_ineq, aes(x = TIME_PERIOD, y = coef_ineq_value, color = iso3, group = iso3)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Coefficient of human inequality per Country (EU27)",
    x = "Year",
    y = "Coefficient of human inequality",
    color = "Country"
  ) +
  theme_minimal()

ggplotly(p_coef)
```

### Descriptive statistics

```{r, echo=FALSE}
summary_stats_coefineq <- coef_ineq %>%
  group_by(iso3) %>%                        
  summarise(
    Min = min(coef_ineq_value, na.rm = TRUE),
    Max = max(coef_ineq_value, na.rm = TRUE),
    Mean = mean(coef_ineq_value, na.rm = TRUE),
    Median = median(coef_ineq_value, na.rm = TRUE),
    SD = sd(coef_ineq_value, na.rm = TRUE)
    )

kable(summary_stats_coefineq, 
      caption = "Descriptive statistics of the coefficient of human inequality by country (all years combined)")
```

# linear regressions

## Between taxes and GHG

```{r, echo=FALSE}
model <- lm(GHG_Emissions_value ~ Envt_Tax_Revenues_value + GDP_value + sector + TIME_PERIOD,
            data = all_merged)

summary(model)
```
REDO THE COMMENT :
We estimated a linear regression to examine the relationship between environmental taxes and greenhouse gas emissions, while controlling for GDP, emission sector, and year. The model includes 16,700 observations after excluding 500 rows with missing data. It explains approximately 62% of the variation in emissions (R-squared = 0.624), which is higher than in the previous version.

The coefficient on environmental tax revenues is –0.024, which is not statistically significant (p = 0.363). This indicates that, after accounting for GDP, sectoral structure, and time, there is no detectable correlation between environmental tax revenues and GHG emissions in this data set.

GDP has a positive and highly significant effect on emissions, with a coefficient of 0.048 (p \< 0.001). This means that countries with higher GDP tend to have higher emissions, consistent with scale effects: larger economies produce more output and therefore more emissions.

Sectoral effects are strong and highly significant. Relative to the baseline category (Agriculture), estimated additional emissions are:

-   Manufacturing: +18,040 units
-   Public electricity and heat production: +29,820 units
-   Transport: +15,340 units

These large differences reflect the inherent emission intensities of these sectors.

The coefficient on the year variable is –1,034, highly significant (p \< 0.001), indicating a strong downward trend in emissions over time, likely due to EU climate policies, efficiency improvements, and structural economic changes.

Overall, the results suggest that GDP and sectoral composition are the dominant drivers of emissions, while environmental taxes, in the form measured here, show no statistically significant association with emission levels. The negative time trend indicates that emissions have declined over the studied period independently of environmental tax revenues.

## Between taxes and inequalities

```{r, echo=FALSE}
# taxes on inability to warm house
model_ineq <- lm(Warm_Households_value ~ Envt_Tax_Revenues_value + GII_value + coef_ineq_value + hhtyp + incgrp,
                 data = all_merged)

summary(model_ineq)

# taxes on inequalities
model_ineq_tax <- lm(coef_ineq_value ~ Envt_Tax_Revenues_value + hhtyp + incgrp, data = all_merged)
summary(model_ineq_tax)

model_gii_tax <- lm(GII_value ~ Envt_Tax_Revenues_value + hhtyp + incgrp, data = all_merged)
summary(model_gii_tax)

```


short comments (redo)

1. Heating difficulties Higher environmental taxes slightly reduce households’ inability to heat their homes (coefficient = -5.97×10⁻⁵, p \< 0.001). Gender inequality (GII) and overall inequality (coef_ineq) are the main drivers, with coefficients 71.2 and 3.02, respectively (both p \< 0.001). Household type and income group also matter. Model explains 59% of variation (adj. R² = 0.590).

2.  Effect on income inequality Higher environmental taxes are associated with a tiny increase in income inequality (coefficient = 1.705×10⁻⁵, p \< 0.001). The model explains less than 1% of inequality variation (R² = 0.0095), so the economic effect is very small despite statistical significance.

3.  Effect on gender inequality (GII) Higher environmental taxes slightly reduce gender inequality (coefficient = -1.115×10⁻⁶, p \< 0.001). The model explains only 4% of GII variation (R² = 0.042), indicating a statistically clear but economically minor effect.



-\> je dois revoir la suite pour utiliser le big merge si nécessaire -\> across : pour avoir impact de taxe / emission sur pays + quel niveau de taxe par pays !!!

conclu =\> social cost associated to tax ? == faire une politique en tenant compte


# Forecasting

### impact of tax on ghg

```{r, echo=FALSE}

model_ghg_tax <- lm(GHG_Emissions_value ~ Envt_Tax_Revenues_value, data = all_merged)
future <- data.frame(Envt_Tax_Revenues_value = seq(min(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
                                      max(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
                                      length.out = 25))
predicted_GHG <- predict(model_ghg_tax, newdata = future)
predicted_GHG
```

### Forecasting impact of tax on GII

```{r, echo=FALSE}

model_gii_tax <- lm(GII_value ~ Envt_Tax_Revenues_value, data = all_merged)
predicted_GII <- predict(model_gii_tax, newdata = data.frame(Envt_Tax_Revenues_value = seq(
  min(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
  max(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
  length.out = 25)))
predicted_GII
```


### Forecasting impact of tax on Households incapable of warming their home
```{r, echo=FALSE}

model_warm_tax <- lm(Warm_Households_value ~ Envt_Tax_Revenues_value, data = all_merged)
predicted_Warm <- predict(model_warm_tax, newdata = data.frame(Envt_Tax_Revenues_value = seq(
  min(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
  max(all_merged$Envt_Tax_Revenues_value, na.rm = TRUE),
  length.out = 25)))
predicted_Warm
```
