---
title: "dm-project"
author: "Daphné Perry and Justine Catta"
format: html
---

### Our data
1. 
Environmental tax revenues: energy, transport, pollution, resources
source : OECD
time period : 1994-2023

2. 
GDP and Added Value per capita annual growth per sector 
source : OECD
time period : 1990-2024

3.
Greenhouse gas emissions by source sector
source : Eurostat
time period : 1990-2024

4.
Inability to keep home adequately warm
source : Eurostat
time period : 1990-2024



### Setup
```{r}
library(vroom)
library(ggplot2)
library(knitr)
library(dplyr)
library(here)
library(countrycode)
```

```{r}
here::i_am("project.Rproj")
```


### Data loading
```{r}
Envt_Tax_Revenues <- vroom(here("data", "OECD_envt_tax.csv"))
GDP_and_AV <- vroom(here("data", "OECD_GDP_and_AV.csv"))
GHG_Emissions <- vroom(here("data", "Eurostat_ghg.csv"))
Warm_Households <- vroom(here("data","Eurostat_warm.csv"))
Soc_Indexes <- vroom(here("data", "HDR_indexes.csv"))
```

### Country keys

List of countries :
Austria, Belgium, Bulgaria, Croatia, Cyprus, Czechia, Denmark, Estonia, Finland, France, Germany, Greece, Hungary, Ireland, Italy, Latvia, Lithuania, Luxembourg, Malta, Netherlands, Poland, Portugal, Romania, Slovakia, Slovenia, Spain, Sweden

```{r}
selected_countries <- c("Austria","Belgium","Bulgaria","Croatia","Cyprus","Czechia","Denmark","Estonia","Finland","France","Germany","Greece","Hungary","Ireland","Italy","Latvia","Lithuania","Luxembourg","Malta","Netherlands","Poland","Portugal","Romania","Slovakia","Slovenia","Spain","Sweden")
country_codes <- c("AT","BE","BG","HR","CY","CZ","DK","EE","FI","FR","DE","GR","HU","IE","IT","LV","LT","LU","MT","NL","PL","PT","RO","SK","SI","ES","SE")
country_mapping <- data.frame(country_name = selected_countries, country_code = country_codes)
print(country_mapping)


```










